### 2013-02-26 article/main.tex
Small fixes and comments marked as MUN1

%% bare_jrnl_compsoc.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE Computer
%% Society journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
% The Computer Society usually requires 10pt for submissions.
%
\documentclass[10pt,journal,letterpaper,compsoc]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[12pt,journal,compsoc]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

%%extra packages ------ 
\usepackage{longtable}   
\usepackage{float} %allows the use of [H] to force positioning of figures
\usepackage{subfigure}
\usepackage{hyperref} %used in references now  --- it shoud be removed
\usepackage{amssymb}
\usepackage[neveradjust]{paralist} %compact lists
\usepackage{amsthm}

\newtheorem*{theory}{Theory}

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  % \usepackage[nocompress]{cite}
\else
  % normal IEEE
  % \usepackage{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later. Note also the use of a CLASSOPTION conditional provided by
% IEEEtran.cls V1.7 and later.

% *** GRAPHICS RELATED PACKAGES ***
%
%\ifCLASSINFOpdf
   \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
   \graphicspath{{figures}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
%\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
%   \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
   %\graphicspath{{./figures/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
   %\DeclareGraphicsExtensions{.eps}
%\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%\usepackage[tight,normalsize,sf,SF]{subfigure}
%\else
%\usepackage[tight,footnotesize]{subfigure}
%\fi
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. Computer Society papers
% use a larger font and \sffamily font for their captions, hence the
% additional options needed under compsoc mode. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false]{caption}
%  \usepackage[font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false]{caption}
%  \usepackage[font=footnotesize]{subfig}
%\fi
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
% This package may be useful when used in conjunction with IEEEtran.cls'
% captionsoff option. Some IEEE journals/societies require that submissions
% have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.3.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% For subfigure.sty:
% \let\MYorigsubfigure\subfigure
% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat/subfig command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Software Development in Startup Companies: The Green-field Startup Model}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

\author{FirstName~SecondName,~\IEEEmembership{Fellow,~IEEE,}
        ..~..,~\IEEEmembership{Fellow,~IEEE,}
        and~...~....,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem Author info.\protect\\
% note need leading \protect in front of \\ to get a newline within \thanks as
% \\ is fragile and will error, could use \hfil\break instead.
E-mail: see http://something/contact.html
\IEEEcompsocthanksitem Author info.}% <-this % stops a space
\thanks{}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer 
Society Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2007  
% IEEE}%
%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer  
% Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society jorunal
% papers don't need this extra clearance.)

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
%\boldmath
Software startups are newly created companies with no operating history and 
oriented in producing cutting-edge technologies. These companies develop 
software under highly uncertain conditions, tackling fast growing markets with 
severe lack of resources. Despite their increasing importance in economy there 
are only few scientific studies attempting to address software engineering (SE) 
issues, especially for early-stage startups. This research aims to understand 
the software development strategies conducted by practitioners, in the period of 
time that goes from idea conception to the first open beta release. A 
\textit{grounded theory} approach guided the execution of a case study to 
explore the state-of-practice. As result, the Greenfield startup model (GSM)  
reveals the urgent priority of startups of releasing the product as quickly as 
possible to verify the product/market fit and to adjust the business and product 
trajectory according to early collected user feedback. Nevertheless, the need of 
shortening time-to-market, by speeding-up the development through low-precision 
and product-centric engineering activities, is counterbalanced by the need of 
restructuring the product before setting off for further growth.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway. In particular, the Computer Society does
% not want either math or citations to appear in the abstract.

% Note that keywords are not normally used for peer review papers.
\begin{keywords}
Software Development, Startups, Grounded Theory.
\end{keywords}}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEcompsoctitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynotcompsoctitleabstractindextext when compsoc mode
% is not selected <OR> if conference mode is selected - because compsoc
% conference papers position the abstract like regular (non-compsoc)
% papers do!
\IEEEdisplaynotcompsoctitleabstractindextext
% \IEEEdisplaynotcompsoctitleabstractindextext has no effect when using
% compsoc under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

%------------------------------------------------------------------------------%
\section{Introduction}
\label{intro}
% Computer Society journal papers do something a tad strange with the very
% first section heading (almost always called "Introduction"). They place it
% ABOVE the main text! IEEEtran.cls currently does not do this for you.
% However, You can achieve this effect by making LaTeX jump through some
% hoops via something like:
%
%\ifCLASSOPTIONcompsoc
%  \noindent\raisebox{2\baselineskip}[0pt][0pt]%
%  {\parbox{\columnwidth}{\section{Introduction}\label{sec:introduction}%
%  \global\everypar=\everypar}}%
%  \vspace{-1\baselineskip}\vspace{-\parskip}\par
%\else
%  \section{Introduction}\label{sec:introduction}\par
%\fi
%
% Admittedly, this is a hack and may well be fragile, but seems to do the
% trick for me. Note the need to keep any \label that may be used right
% after \section in the above as the hack puts \section within a raised box.

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps (small caps for compsoc).
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{A}{n} impressive number of new software startups are launched 
worldwide every day as a result of an increase of new markets, accessible 
technologies, and venture capital \cite{8491286}.  With the term 
\textit{software startups} we refer to those temporary organizations focused on 
the creation of high-tech and innovative products, with little or no operating 
history, aiming to grow by  aggressively scaling their business in highly 
scalable markets.

New ventures such as \textit{Facebook}, \textit{Linkedin}, \textit{Spotify}, 
\textit{Pinterest}, \textit{Instagram}, \textit{Groupon} and \textit{Dropbox}, 
to name a few, are good examples of startups that evolved into successful 
businesses. But, despite many successful stories, the great majority of startups 
fail within two years from their creation, primarily due to self-destruction 
rather than competition \cite{Crowne2002}. Operating in a chaotic, rapidly 
evolving and uncertain domain, software startups face intense time-pressure from 
the market and are exposed to tough competition \cite{Maccormack2001, 
Eisenhardt1998}. In order to succeed in this environment, it is crucial to 
choose the right features to build and be able to quickly adapt the product to 
new requests constrained by very limited amount of resources \cite{Sutton2000}.

From a software engineering perspective development in startups is challenging 
since they develop software in a context where processes can hardly follow a 
prescriptive methodology \cite{Sutton2000, Coleman2005}. Despite startups share 
some characteristics with similar domains (e.g. small and web companies), the 
combination of different factors makes the specific development context quite 
unique \cite{Blank2005, Sutton2000}. Therefore, research is needed to support 
their engineering activities \cite{Coleman2005}, guiding practitioners in taking 
decisions and avoiding choices that could easily lead the whole business to 
failure \cite{Kajko-Mattsson2008}.
However, despite the impressive size of the startup ecosystem 
\cite{ISI:000243253000007}, the state-of-the-art presents a gap. Through a 
\textit{Systematic Mapping Study} (SMS) of the literature, conducted by 
Paternoster et al. \cite{SMS}, only few publications were found related to 
engineering activities in startups. Moreover the majority of these studies do 
not possess the relevance and rigour required to form a consistent body of 
knowledge in the state-of-the-art and in the state-of-practice.

This research aims to understand how software development strategies are 
engineered by practitioners in startup companies in terms of level of: 
structure, planning and control of software projects, in the period of time  
that goes from idea conception to the first open beta release of the software 
product.

With a cross-sectional study we focused the research context in a limited 
time-frame to emphasize the nature of uncertainty in the development activities 
and time pressure from the market, as discussed in \cite{Maccormack2001, 
Eisenhardt1998}. We performed semi-structured in-depth interviews with CEOs and 
CTOs of startups, covering a wide spectrum of thematics and iteratively 
adjusting the direction of the research according to the emerging evidences.

By means of the Green-field startup model (GSM), we captured the underlying 
phenomenon of software development in early-stage startups. The framework is 
based on a systematic procedure  which is grounded into empirical concepts 
obtained by a case-study. GSM presents the most significant themes of the 
development strategies that characterise startups' context.

The results of our study suggest that the most urgent priority of software 
development in startups is to shorten time-to-market. The greater part of 
engineering activities are focused on the implementation while only little 
attention is given to more conventional activities (project management, 
requirement specifications, analysis, architecture design, automatic testing). 
The first release of the product includes only a limited set of well suitable 
functionalities.

Nevertheless, by speeding-up the development process, startups accumulate 
technical debt, ignoring processes, relying on informal communication and 
replacing documentation with low-precision artifacts. For startups it is 
essential to look ahead within short-term deadlines and determine the right 
product/market fit driven by user feedback, using flexible and reactive 
development approaches and adopting mitigation strategies to contain the 
severity of accumulated technical debt.

Practitioners can use the GSM to apply presented strategies to speed-up the 
development by taking in consideration the amount of technical debt accumulated 
during the first early-stage period. In this regard, we identified several 
commonalities between the issues related to software development in startups and 
the research focused at studying technical debt \cite{Nugroho2011,Izurieta2012}. 
On the other hand, researchers can take advantage from the GSM to 
understand how the technical debt influences the future growth of startup 
companies.

The remainder of this paper is structured as follows:  we provide relevant 
information about startups in section \ref{backg}. Section \ref{resmet} 
introduces the research questions and discusses the design and execution of the 
research process. Results are discussed in section \ref{res:gsm}, which presents 
the GSM. Section \ref{res:val}  compares results of this research to 
literature and frameworks presented in this field. Section \ref{valt} discusses 
the validity threats considered during this study. Section 
\ref{sect:theory:impl} discusses the most relevant implications of the GSM. 
Finally, this paper concludes in section \ref{conc}.

%MUN: This is maybe only a minor concern, but it may be relevant for certain 
%reviewers: The references you use in the introduction are not very recent. 
%Look at the list of papers [1-15], most of them are published before 2005.
%I understand that you use some references to strengthen your general arguments 
%(which can refer to older publications), but your introduction should also 
%show the relevance of your research right now. This is best shown by 
%referring to recent (max 5 years old) literature. Think about if you can add 
%more recent (2010-now), well published, papers in the introduction.

% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)

%CG: Not fixed. There is a huge gap on the topic, the refs in the introduction 
%are those retrieved by the SMS.

%\hfill mds

%\hfill January 11, 2007
%------------------------------------------------------------------------------%
\section{Background}
\label{backg}

Modern entrepreneurship has been boosted by the advent of the consumer internet 
markets in the middle of the 1990s and culminated with the notorious 
\textit{dot-com bubble burst} of 2000 \cite{Perkins:1999:IBI:555126, 
storey1982entrepreneurship}. Several years later, with the massification of the 
internet and mobile devices, we are now assisting to an impressive proliferation 
of software ventures - metaphorically referred as the \textit{startup bubble}. 
The easy access to vast potential markets and the low cost of services 
distribution are appealing condition for modern entrepreneurs \cite{Marmer2011}. 
Inspired by success stories, a large number of software businesses are created 
every day. However, the great majority of these companies fail within two years 
from their creation \cite{Crowne2002}.

Looking at the number of new business incubators which appeared in the last 
decade one can estimate the importance of startups \cite{Grimaldi2005}. The wave 
of disruption in new technologies has led companies to be more and more 
competitive, forcing themselves to radical organizational and innovational 
renewals,  which bring many companies to the attempt of behaving like startups 
\cite{Christensen1997}.

The implementation of methodologies to structure and control the development 
activities in startups is a major challenge for engineers \cite{Coleman2008}. 
Generally, the management of software development is achieved through the 
introduction of software processes, which define what steps the development 
organizations should take at each stage of production and provide assistance in 
making estimates, developing plans and measuring quality \cite{Chroust1996}. In 
the last decades, several models have been introduced to control software 
development activities. However, their application in startup companies doesn't 
report significant benefits \cite{Coleman2008a, Coleman2008, Sutton2000}.

In this context, software engineering (SE) faces complex and multifaceted 
obstacles in understanding how to manage development processes. Sutton defines 
startups as creative and flexible in nature and reluctant to introduce process 
or bureaucratic measures which may result ineffective \cite{Sutton2000}. Also 
Bach refers to startups as \textit{``a bunch of energetic and committed people 
without defined development processes''} \cite{Bach1998}. In fact, startups have 
very limited resources and typically wish to use them to support product 
development instead of establishing processes \cite{Coleman2008, Heitlager2007}. 
Some attempts to taylor lightweight processes to startups reported basic failure 
of their application: \textit{``Everyone is busy, and software engineering 
practices are often one of the first places developers cut corners''} 
\cite{Martin2007}. Rejecting the notion of repeatable and controlled processes, 
startups prominently take advantage of unpredictable, reactive and low-precision 
\cite{surviving-os-cockburn} engineering practices \cite{Sutton2000, 
Tanabian2005, Chorev2006, Kakati2003}.

Moreover, most startups develop packaged applications rather than software for 
a specific client \cite{genome2012}. Issues related to this domain are addressed 
in literature by the area known as \textit{market-driven software development} 
(sometimes called \textit{packaged software development} or \textit{COTS 
software development} \cite{Alves2006}). Among other results, researchers 
emphasize the importance of time-to-market as a key strategic objective 
\cite{dagMDR,sawyer99} for companies operating in this domain.  Other peculiar 
aspects which influence the software development are related to requirements, 
which are reported to be often ``invented by the software company'' 
\cite{512553}, ``rarely documented'' \cite{Karlsson02challengesin}, and can be 
validated only after the product is released to market 
\cite{dahl2003,Keil:1995}. Under these circumstances, failure of product 
launches are largely due to ``products not meeting customer needs''  
\cite{Alves2006}.

Accordingly, product-oriented practices help startups having flexible teams, 
with workflows that leave them the ability to quickly change the direction 
according to the targeted market  \cite{Heitlager2007,Sutton2000}. In this 
regard, many startups focus on team productivity, asserting more control to the 
employees instead of providing them rigid guidelines \cite{Tanabian2005, 
Chorev2006, Kakati2003}.

%------------------------------------------------------------------------------%
\subsection{General lack of research in startups}
Despite some studies tried to address the above mentioned issues, a SMS 
\cite{SMS} revealed only a few SE works in this specific area, as confirmed by 
other studies \cite{Coleman2008, Coleman2008a, Coleman2007, Sutton2000}. 
Moreover, the studies, identified in \cite{SMS}, appear to be highly fragmented 
and spread across different areas rather than constituting a consistent 
\textit{body of knowledge}.

A research published in 2000 by Sutton \cite{Sutton2000} confirmed a general 
lack of  studies in this area, claiming that \textit{``software startups 
represent a segment that has been mostly neglected in process studies''} and it 
has been further confirmed with the empirical studies of Coleman et al. 
\cite{Coleman2008,Coleman2008a,Coleman2007} eight years later.

%------------------------------------------------------------------------------%
\subsection{Software development in startups}

The word \textit{startup} appeared in the SE literature for the first time in 
1994 in an article written by Carmel \cite{Camel1994a} where he studied the 
time-to-completion in a young package firm. %MUN: If this is a quote 
%from Carmel, put it in quotation marks, like you did in the previous section 
%with Suttons quote. Generally, check that you are consistent how you use 
%quotations throughout the paper. 
%CG: fixed, btw  it was just to emphasize his the focus of his study. 
%MUN1: Ok. As you may have noticed, I have removed a lot of cases were you use 
%italics. If you emphasize *too* much, it it rather distracting than helpful 
%for the reader. Furthermore, it conflicts with you convention you use to 
%highlight high level categories of GSM.
He noticed how these companies were particularly innovative and successful, 
advocating for the need of more research investigating software development 
practices so as to replicate success and try to transfer it to other technology 
sectors.

First insights reveal how software startups are product-oriented in the first 
period of their development phase \cite{Heitlager2007}. Despite good 
achievements at the beginning, software development and organizational 
management increase in complexity \cite{1456074,Banker1998} causing  
deterioration of performance over time. Briefly, the necessity of establishing 
initial repeatable and scalable processes cannot be postponed forever.

A study of Kajko-Mattsson \cite{Kajko-Mattsson2008}, which investigated a 
Swedish software startup, reported a heavy lack of requirements gathering 
processes, minimal project management, lack of control over the change requests, 
absence of documentation to track the status and progress of the process, and 
defective releases. Accordingly, Ambler et al. report how two startups 
approaching to an upcoming IPO started to require processes to focus on scalable 
solutions, in view of the growing company's size in terms of users and employees 
\cite{Ambler2002}. In this regard, Crowne \cite{Crowne2002} specifies different 
stages through which software startups evolve. Starting without any established 
workflows, startups grow over time, creating and stabilizing processes to 
eventually improving them only when sufficiently mature.

But when startups have no time for training and orienting activities, as 
discussed in \cite{Sutton2000}, their main focus remains on team capabilities 
instead of prescriptive processes, hiring people who can \textit{``hit the 
ground running''} \cite{Yoffie1999}. Empowering the team and focus on 
methodological attributes of the processes oriented in prototyping, 
proof-of-concepts, mock-ups and demos, to test basic functionalities, have been 
the primary priority in startups \cite{Camel1994a}. Only when they grow, formal 
methodologies arise, followed by performing quality assurance and long-term 
planning processes \cite{Yoffie1999}.

The maturity of a company affects the extent to which processes are adopted. 
Tingling \cite{Tingling2007} reports how introducing Extreme Programming (XP) 
principles \cite{Beck:2004:EPE:1076267} in the development process was 
challenging because of the need of trained team-members for fully implementing 
the methodology. Da Silva and Kon \cite{Silva2005} were able to start with all 
the XP practices in place only after six months of coaching the team, trying to 
enhance maturity from day-one. Nevertheless, even then, customization of 
practices were inevitably implemented to adapt the processes to the undertaken 
startups' context \cite{Deias}.

As partially discussed above, contributions to flexibility and reactiveness of 
the development process has been conducted prominently by means of \textit{Lean} 
\cite{Gautam2008} and \textit{Agile} \cite{Abrahamsson2002} methodologies (also 
reported in \cite{Taipale2010, Kuvinka2011}), where the extreme uncertain 
conditions lead startups to learn fast from trials and errors with a strong 
customer relationship in order to avoid wasting time in building wrong 
functionalities and prevent rapidly exhaustion of resources \cite{Midler2008, 
Hilmola2003, Sutton2000}. Customer involvement in software development has also 
been discussed by Yogendra \cite{Yogendra2002} as an important factor to 
encourage an early alignment of business concerns to the technology strategies, 
because both are salient considerations to be successful \cite{Tingling2007}.

When startups take a development approach that is mainly product-oriented 
rather than process-oriented, it is essential to have a flexible team, with a 
workflow that helps them quickly to change direction according to the target 
market \cite{Sutton2000}. In this regard, many startups have been focused on 
team productivity, providing more control to the employees instead of providing 
them rigid guidelines \cite{Tanabian2005, Chorev2006, Kakati2003}.

%------------------------------------------------------------------------------%
\subsection{Software process improvement in startups}

When startups are mature enough to support software process improvement (SPI), 
the solutions considered according to the state-of-the-art are oriented towards 
light-weight processes such as a design of development process based on XP, 
proposed by Zettel \cite{Zettel2001}. The process consists of a set of 
activities and artifacts (in addition to some important roles) defined in order 
to identify responsibilities and tools to utilize. But, despite the promising 
benefits reported by Zettel, we were not able to identify any future evaluation 
in real-world settings.

Another attempt of SPI in startups has been conducted by Deakins et al. 
introducing a \textit{Helical Model for managing e-commerce development 
environment} \cite{Deakins2005}. Also in this case, the author prescribed broad 
guidelines for a rapid high-quality development process, which underwent limited 
testing only in academic settings.

The problem of \textit{one-size-fits-all}, related to the SPI  representations 
for startups, is described by Fayad \cite{Fayad1997}. The author discusses the 
problem in actuating the same \textit{best-practices} criteria for established 
companies in 10-person software startups. Thoroughly remarked in 
\cite{Sutton2000}, Sutton states that problems of SPI in software startups arise 
because of: the dynamism of the development process, which precludes 
repeatability; organizational maturity, that cannot be maintained from startups 
in view of lack of corporate direction; severe lack of resources, both human and 
technological for process definition, implementation, management, and training. 
In conclusion, the primary benefits of SPI do not address startups, which 
instead of promoting product quality, aim to minimize time-to-market.

Additionally, the role of SPI has been neglected because seen as an obstacle to 
the development team's creativity and flexibility, and to 
the need of a quick delivering product process environment \cite{Coleman2008a}.
%MUN: the second part of the sentence is just a statement without reference. If 
%it refers to a statement by Coleman, put his reference at the end OR find the 
%correct reference OR delete the statement.
%CG fixed
Product quality is left aside in favor of minimal and suitable functionalities 
to shorten the time-to-market. As reported in two studies of Mater and Mirel 
\cite{Mater2000, Mirel2000}, quality aspects, mostly taken in consideration in 
internet startups, are oriented to usability and scalability, even though the 
market and application type heavily influences the quality-demand 
\cite{Coleman2008, Kim2005}.

To maintain the development activities, oriented to limited but suitable 
functionality, studies suggest to externalize the complexity of parts of the 
project to third party solutions by means of outsourcing activities 
\cite{Hanna2010}, software reuse \cite{Jansen2008} and open-source strategies , 
\cite{Wall2001, Bean2005}. 
%MUN: Can you split up the references? I.e. write:
%"outsourcing activities \cite{Bla}, software reuse \cite{Bla}, and open-source 
%strategies \cite{Bla}.
%CG: fixed
%------------------------------------------------------------------------------%
\subsection{Technical debt}
Notwithstanding, a new area of the SE research, trying to tackle the problem of 
the \textit{technical debt},  brings and encompasses different implications in 
studying development in software startups \cite{Tom2013}. The metaphoric 
neologism was originally introduced by Cunningham in 1992 
\cite{TechnicalDebtCunn} and has recently attracted the attention of SE 
researchers\footnote{To attest the fact that \textit{technical debt} is gaining 
traction among researchers, we mention two important contributions which 
characterize the \textit{``debt landscape''} : \cite{Nugroho2011,Izurieta2012} 
and a dedicated workshop \cite{workshopDebt} organized by the Software 
Engineering Institute  and ICSE.}. The concept of \textit{technical debt} is 
well illustrated in \cite{Brown:2010:MTD:1882362.1882373} : \textit{``The idea 
is that developers sometimes accept compromises in a system in one dimension 
(e.g., modularity) to meet an urgent demand in some other dimension (e.g., a 
deadline), and that such compromises incur a ``debt'' on which ``interest'' has 
to be paid and which the ``principal'' should be repaid at some point for the 
long-term health of the project''}.  The compromise between high-speed and 
high-quality engineering is faced daily by software startups, not only in terms 
of architecture design but in multifaceted aspects (weak project management, 
testing, process control).

Since \textit{``all decisions related to product development are trade-off 
situations''} \cite{Hilmola2003}, generally startups optimize workflows to the 
dynamic context they are involved into. They typically adopt any development 
style that might work to support their first needs in what is called the 
\textit{``Just do it''} school of software startups \cite{Ries2011}. 
Additionally, as remarked by Coleman, \textit{``many managers just decide to 
apply what they know, as their experience tells them it is merely common 
sense''} \cite{Coleman2008}.

\subsection{Terminology}
In order to set a common ground and to prevent ambiguity, we use the following 
terminology throughout the paper:

\begin{compactitem}
\item Software development strategy: the overall approach adopted by the 
company to carry out the product development.
\item Engineering activities: the activities needed to bring a product from 
idea to market. Traditional engineering activities are, among others, 
requirement engineering, design, architecture, implementation, testing.
\item Engineering elements: any practice, tool and artifacts contributing and 
supporting the engineering activities.
\item Quality attributes: those overall factors that affect run-time behavior, 
system design, and user experience. They represent areas of concern that have 
the potential for applications to impact across layers and tiers. Some of these 
attributes are related to the overall system design, while others are specific 
to run time, design time, or user centric issues \cite{Microsoft2009}.
\item Growth: an increase in the company size respect to the initial conditions 
in terms of either employees or users/customers, and product complexity in terms 
of handling an increasing number of feature requests.
\item Software product: any software product and software service.
\end{compactitem}
%------------------------------------------------------------------------------%
\section{Research methodology}
\label{resmet}

\begin{figure*}[!t]
\centering
\includegraphics[width=5in]{figures/completemethodology2}
\caption{Grounded theory study process overview}
\label{fig:gt:completemethodology}
\end{figure*}
%MUN: I think you can increase the font size in the figure a bit. Try thereby
%to keep the height of the figure, expanding however the width (use horizontal
%space since it is a two column figure). Change also in figure:  - Write "Data
%Analysis" horizontally  - Change names of macro-phases to upper-case, i.e.
%Design and Execution,     Data Collection, Data Analysis, etc.  - Add section
%number to macro-phase name, i.e. 3.1 Design and Execution, 3.2     Data
%Collection  - Change "Discussion and Validation" into "Theory Validation" as
%you name it     in the text. Also, add a section (3.5) that describes shortly
%what you did in "Theory  Validation" 
%CG: fixed... fontsize cannot be increased uniformly among all the boxes, then 
%I skipped that part...
%MUN1: OK, but in the figure, the correct section number for theory validation 
%is 3.5, not 3.4

The aim of this research is to \textit{understand how software development 
strategies are engineered by practitioners in startup companies in terms of 
level of: structure, planning and control of software projects, in the period of 
time  that goes from idea conception to the first open beta release of the 
software product} \footnote{The above mentioned goal is structured according to 
the GQM paradigm \cite{Basili1992}.}.

Initially the boundaries of the research domain were set by means of 
non-systematic literature survey, which provided the initial keywords used 
further in the research process. Moreover it helped us in defining the research 
questions (RQs):

\begin{compactitem}

\item RQ-1 : What is the current state-of-practice related to software 
development strategy in early-stage startups?
\begin{compactitem}

\item RQ-1.1 : How do startups structure and execute the main engineering 
activities?
\item RQ-1.2 : How are product quality attributes considered by startups?
\end{compactitem}

\end{compactitem}

We investigated the software development approach undertaken by practitioners 
of startups. Following a \textit{Grounded Theory} (GT) methodology 
\cite{Glaser1978}, we executed 13 semi-structured interviews integrated with 
follow-up questionnaires. From this, we elaborated and extracted the Green-field 
startup model (GSM) explaining the underlying phenomenon of software development 
in startups.

\textit{Grounded Theory} methodology is described by Robson as \textit{``the 
best approach to answer the question - What is going on here?''} and defined by 
its creators \cite{ColinRobson2009} as \textit{``a set of well-developed 
categories ( e.g. themes, concepts ) that are systematically interrelated 
through statements of relations to form a theoretical framework that explains 
some relevant social, psychological, educational, nursing or other 
phenomenon''}.

Despite GT has its roots in the social sciences, it has been increasingly used 
by Information System researchers and has been almost ignored by the SE research 
\cite{Coleman2007}. In 1997 Bertelsen advocated for a need of more qualitative 
research in SE \cite{Bertelsen1997}, and in the last decade we assisted to an 
increasing number of publications that used a GT-like approach in SE 
\cite{Coleman2008a,Coleman2007, Sulayman2012}.

Following the GT principles we attempted to capture the most relevant aspects 
of software development from startup practitioners, letting a theory emerge 
from the interviews and adjusting the research hypotheses and questions as we 
proceeded through. During these interviews we collected data related to 
engineering activities undertaken by startups in the time frame between the idea 
conception and the first open beta release of the product.  Then, we proceeded 
to the analysis of the data, finding important relations among concepts with a 
formal approach.

As suggested by Coleman, in view of the different versions of GT which emerged 
in the last years, researchers should indicate which \textit{``implementation''} 
of the theory is being used \cite{Coleman2007}. After the methodology was 
initially introduced, its original authors  had divergences regarding how the 
theory should be formed by analyzing the data (Strauss and Corbin on one side, 
Glaser on the other).  Glaser advocated for a more \textit{puristic} approach, 
where the theoretical categories should \textit{``naturally emerge from the 
data''}, whilst Strauss and Corbin formulated an updated version of the 
methodology, which leaves to the researchers an higher degree of freedom. The 
latter approach empowers researchers' \textit{``theoretical sensitivity''} 
\cite{Corbin1990}, and encourages them to outline the research problem 
beforehand. Since the knowledge obtained from parallel execution of the SMS 
\cite{SMS} and our direct experience with startups companies provided a good 
initial level of knowledge, in this study we use the Corbin and Strauss 
approach \cite{Strauss1998}.

In this research we address technical aspects related to software development 
in startups, exploring their operational dynamics. In view of the lack of 
agreement on an unique definition of the word \textit{startup}, we delimited our 
initial contextual boundaries to newly created innovative and single-product 
software companies, in the time-frame that goes from the idea conception to the 
release of the first product in highly scalable markets. Moreover, since most 
startups are web-oriented companies \cite{tc-webstartups-link, Allen2003}, in 
our case study, we inquired web companies with little or no operating history.

A complete overview of the overall case study methodology and execution is 
shown in Figure \ref{fig:gt:completemethodology}, which presents how we tailored 
the general GT methodology to our specific needs.

The initial design of the case study is supported by the results of the 
\textit{Systematic Mapping Study} (SMS) \cite{SMS} which contributed to define 
the initial questions for both questionnaires and semi-structured interviews. 
The process depicted in Figure \ref{fig:gt:completemethodology} is evolutionary 
and affects the design at each new iteration. In \textit{Data Collection} the 
empirical results are integrated and stored in a single case study database and 
subsequently processed in \textit{Data Analysis} to form the theoretical 
categories.  At each iteration the new emergent theory is updated following a 
formal procedure (\textit{Theory Generation}), and after verifying that 
\textit{Theoretical Saturation}\footnote{The point at which executing more 
interviews wouldn't bring any additional value for constructing the theory.} of 
categories has been achieved, we finally proceeded to \textit{Theory 
Validation}.

The whole procedure has been executed simultaneously in pair on the same screen 
by the first two authors, handling conflicts by reviewing the rationale of 
decision with the third and fourth authors. When necessary we performed an 
in-depth review of the design of the research and data collected during the 
execution process.

The process details are described in the following subsections, structured 
according to the five macro phases depicted in bold in Figure 
\ref{fig:gt:completemethodology}.

%------------------------------------------------------------------------------%
\subsection{Design and Execution}
\label{desex}

This case study started based on a SMS of the literature regarding software 
development in startups \cite{SMS}. By means of this SMS we refined our research 
questions. The research questions have been narrowed down to outline the domain 
of our study, even though still broad enough to allow emerging new theories and 
to adjust the scope of the research \cite{Corbin1990}.

The sampling of candidate companies took place in two distinct phases. At first 
we executed an initial convenience sampling \cite{Dawson2009}, which led to the 
identification of eight companies. Then we included five additional startups 
during the theory formation process (theoretical sampling) iteratively improving 
the sample according to the emerging theory. The characteristics of the sampled 
companies are reported in Table \ref{t_interviews-stats}.

\begin{table*}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Companies overview} 
\label{t_interviews-stats}
\centering
\begin{tabular}{|c||c||c||c||c||c|}

\hline
    ID  &  Company age & Founding & Initial & Current & First product  \\
    &   (months)   & team   & developers  & employees  &  building time \\
                    &      & &   &  & (months)   \\
   \hline
C1    & 11    & 4     & 2     & 11    & 6   \\
C2    & 5     & 2     & 2     & 6     & 3      \\
C3    & 18    & 4     & 4     & 4     & 6      \\
C4    & 17    & 3     & 2     & 11    & 6     \\
C5    & 20    & 2     & 1     & 4     & 12   \\
C6    & 30    & 3     & 2     & 4     & 1      \\
C7    & 12    & 2     & 1     & 7     & 4      \\
C8    & 24    & 4     & 3     & 16    & 4    \\
C9    & 5     & 5     & 4     & 5     & 1     \\
C10   & 43    & 6     & 4     & 9     & 4      \\
C11   & 36    & 3     & 3     & 6     & 2     \\
C12   & 12    & 3     & 3     & 3     & 3      \\
C13   & 24    & 2     & 2     & 20    & 3     \\

\hline
\end{tabular}
\end{table*}

All companies, except C10, were founded within the last three years (2009-2012), 
%MUN: I've shortened the text and added absolute numbers. Is 2009-2012 correct? 
%CG: right
by an average of 3 starting members, which were in majority developers. 
Moreover, the number of current employees shows how, to different degrees, 
companies expanded the initial teams. All companies, except C5, brought their 
first product to market in less than 6 months from the idea conception. The 
products consist of web-applications, launched in different nations and market 
sectors. The growing team size and the publicly available data suggest an 
healthy status of the businesses.

We executed the case study online, supported by tools for video conferencing, 
recording each complete session. The interview subjects were the companies' CEOs 
or CTOs. We followed a step-by-step workflow, consisting of the actual 
interview, collection of additional material, preparation of the customized 
follow-up questionnaire and the iterative adjustment of the interview package 
artifacts\footnote{The interview package and follow-up questionnaire has been 
released under MIT License \cite{MITLicense} and it is available for download on 
Github \cite{GitHubInterviewPackage}. We encourage anyone interested in pushing 
this work forward, to fork the repository and contribute to it. If used in a 
research context, please inform us in order to be able to track where and how 
the interview package has been used.}.

The follow-up questionnaires were designed to capture additional data, gather 
missing information and confirm interview results by triangulation. The 
questionnaires have been tailored to each startup, partially taking advantage of 
the repertory grid principles \cite{Edwards2009}. To obtain better chances of 
quickly obtaining responses, questionnaires have been sent to respondents 
immediately after transcribing the interview results. Most of the follow-up  
questionnaires have been collected within one day after the conclusion of the 
last interview.   
%MUN: I don't understand this: you
%collected the data from questionnaires after ALL interviews have been 
%finished, but on the other hand you said the overall process was iterative 
%(data analysis and triangulation let themes emerge which in turn were used
%in subsequent interviews). 
%CG: fixed...

%------------------------------------------------------------------------------%
\subsection{Data collection}
The data collection was designed to allow triangulation, which integrates 
multiple data sources converging on the same phenomenon. After transcribing the 
interview, we extracted conceptual relations, integrating them with the 
questionnaire's results. A well structured case-study database allowed us to 
easily retrieve and seek for information, assembling the evidence from the case 
study reports, as described also in \cite{Yin1994}. The database has been stored 
and constructed using the qualitative data analysis software package 
AtlasTI\footnote{Available online at \url{http://www.atlasti.com/}.} - one of 
the most suitable tools for GT \cite{Coleman2007}. 
We overlapped interviews with questionnaire results to adjust the data 
collection and take advantage of emergent themes and reveal possible 
inconsistencies. The data was analyzed simultaneously and with flexibility in 
mind in such a way that adjustments were made according to the emerging 
findings.
%------------------------------------------------------------------------------%
\subsection{Data analysis}
Before starting the analysis, a data ordering procedure was necessary since 
interviews were spread across a multitude of topics. Therefore transcripts have 
been structured in thematic areas accordingly to different topic cards used 
during the interviews. We proceeded horizontally between same thematic areas of 
transcripts to be able to identify a better number of similar concepts, rather 
than going through an entire transcript at one time.

Once the data were ordered, we executed the process of coding interviews, 
following the steps listed below:

\begin{compactitem}

\item Labels were assigned to raw data, and a first low-level conceptualization 
was carried out using both in-vivo and open coding \cite{ColinRobson2009}.
\item Concepts were grouped together into theoretical categories and 
subcategories. By means of axial coding we first described the different 
relations between subcategories, and then relations between subcategories and 
categories.
\item Categories were refined several times in the attempt to create different 
levels of abstraction and adjusting concepts, aided by a simple knowledge 
management tool.
\item Consistency among categories was validated by exploring and analyzing 
links among subcategories by means of selective coding.
\item The core category - the one with the greatest explanatory power - was 
identified by analyzing the causal relations between high-level categories.
\end{compactitem}

During data extraction we used the technique of in-vivo coding since 
\textit{``direct quoting from the transcript gives more expressive power to the 
data''} \cite{ColinRobson2009} combined with the more descriptive procedure of 
open coding. Following the example of other grounded theories, developed in 
related areas such as Information Systems \cite{Orlikowski1993} and Software 
Process Improvement \cite{Coleman2006}, we performed the high-level 
conceptualization during creation of categories, in the process of refining 
axial and selective coding.

As we were iterating through the interviews, we analyzed new data accordingly, 
by updating codes and categories on necessity, and taking notes in the form of 
memos to adjust the new emerging theory. To improve the speed of the coding 
process, we analyzed all the transcripts one part at-the-time: transcripts 
followed the semi-structured interview format, and this approach enhanced the 
chances of finding similar codes across different transcripts, allowing us to 
quickly iterate and update categories.

Given the large number of codes, categories, properties, propositions and 
related questions that evolved from the analytical process \cite{Corbin1990}, 
an important activity that helped the coding process was memoing. It constituted 
an important component involved in the formulation and revision of theory 
during the research process. For this purpose we made use of three different 
types of memos: code memos, theoretical memos and operational memos. The first 
type of memo is related to the conceptual labeling of open codes, whilst the 
second type concerns axial and selective coding, and thus focusing on paradigm 
features. Finally, operational memos contain directions relating to the evolving 
emerging theory.

After the coding process the first representation of the experience map 
identified in GT is constructed by means of a theoretical framework. The 
theoretical framework is presented in the form of a network of categories and 
subcategories that are linked together according to cause-effect relationships 
\cite{Corbin1990}. During the formation of the theoretical framework the 
researchers operated by means of a  bottom-up approach. From the empirical data 
and coding process, the framework developed into two different levels: a 
detailed level representing the network of subcategories (identified mainly by 
the axial coding process), and a high-level representation of the main 
categories network (identified mainly by the selective coding process).

%------------------------------------------------------------------------------%
\subsection{Theory generation}

As mentioned in Section \ref{desex}, emergent theories have been tested by 
integrating additional companies into the sample, selected following the 
principle of theoretical sampling \cite{Yin1994}.

The process of theory generation took place at each iteration together with 
interview execution, where we systematically analyzed the theoretical framework 
by means of the paradigm model introduced by Corbin \cite{Corbin1990}. The 
paradigm model is composed by:

\begin{compactitem}
\item \textit{Causal conditions}: the events which lead to the occurrence of 
the phenomenon, that is our core category
\item \textit{Context}: set of conditions in which the phenomenon can be 
extrapolated
\item \textit{Intervening conditions}: the broader set of conditions from that 
the phenomenon can be generalized
\item \textit{Action/interaction strategies}: the actions and responses that 
occur as the result of the phenomenon
\item \textit{Consequences}: specification of the outcomes, both intended and 
unintended of the actions and interaction strategies
\end{compactitem}

Within the limits of the critical bounding assumptions, the role of the 
generated theory is to explain, predict and understand the underlying 
phenomenon.

%------------------------------------------------------------------------------%
\subsection{Theory Validation}
\label{rm:val}
%MUN: put here what I wrote in the comment in Section "Theory validation" (i.e.
%the section following the GSM)
%CG: fixed
%MUN1: This description is way too long and verbose. It should be MAX 1 column, 
%and focus on what you present in Section 5. Theory Validation. Until the 
%second-last paragraph you don't talk about that at all! Remove the "theory" on 
%GT validation and what you did *during* the GT process. Keep what you need 
%to motivate
% - 5.1 Comparison of frameworks
% - 5.2 Theoretical categories and existing literature
% - 5.3 Confounding factors
% - 5.4 Static assessment of the GSM
%You need to explain why you did this and how it contributes to the validation 
%of the theory.

Presenting GT is challenging for a researcher, who must pay attention to  
structure the included level of detail, and to the way data is portrayed to 
display evidence of the emergent categories.  In order to assess our research 
and determine if the GT is sufficiently grounded, we used a systematic technique 
to assess the theory, following evaluation lists and factors presented in  
\cite{Strauss1998}.

We consider the following factors:

\par\vspace {0.4cm}
\begin{compactitem}
\item \textit{Fit}, that describes the consistency of the theory with the real 
data.
\item \textit{Understanding}, that describes the degree to which participants 
agree with the theory.
\item \textit{Generality}, that describes the level of abstraction of the 
theory, and how sufficiently it can be adopted by practitioners without losing 
its relevance.
\item \textit{Control}, that describes how well the theory enables 
practitioners to fully understand the described situation.
\end{compactitem}
\par\vspace {0.4cm}

In order to achieve a high-level of fit between the theory and the raw 
empirical data, we started GT without any assumption on how the emergent 
patterns would influence the research direction. Ideas and perceptions evolved 
during the mapping study and interviews. In addition, constant comparative 
methods, overturning of some early categories as new data came to light, 
generation and testing of interim hypotheses, and constant re-evaluation of the 
interview transcripts, ensured that researchers' bias was minimized and 
theoretical fit maintained.

To represent understandable emergent theories, concepts and categories were 
carefully developed using appropriate and self-explanatory codes. The developing 
theory was not immediately presented to the participants to prevent potential 
bias in the responses. But rather, the theoretical framework has been presented 
and described to startups practitioners not involved in the data collection, 
whose reaction was positively expressed.

From generality perspective, we developed a theory from abstract categories, 
which allowed us to create a general guide to constantly changing situations. 
Nevertheless, we took attention on maintaining a reasonable level of 
abstraction, keeping focus on their sensitizing characteristics and providing 
low-level subgroup of categories.

Finally, we provided a comprehensive set of categories with detailed 
interrelations to explain how the \textit{core-category} is achieved and the 
reasons of its consequences, in order to enable practitioners to independently 
understand and analyze situations, predict changes and their consequences, and 
to be capable of revising actions or the theory itself. Using both 
methodological tools and features provided by \textit{AtlasTI}, we explored and 
tested:

\par\vspace {0.4cm}
\begin{compactitem}
\item Each category and the strength of relations between them.
\item Hypotheses, derived from and related to the emergent theory.
\item Deviant cases to ensure robustness and general applicability.
\end{compactitem}
\par\vspace {0.4cm}

Additionally, Strauss and Corbin provided a list of questions to assist in 
determining how well the findings are grounded \cite{Strauss1998}. They 
questions are as follows:

\par\vspace {0.4cm}
\begin{compactenum}
\item Are concepts generated, and are the concepts systematically related?
\item Are there many conceptual linkages and are the categories well developed?
\item Is variation\footnote{Variation refers to the variety of contexts to 
which the theory can be applied.} built into the theory and are the conditions 
under which variation can be found built into the study and explained?
\item Are the conditions under which variation can be found built into the 
study and explained?
\item Has the process been taken into account, and does the theory stand the 
test of time?
\item Do the theoretical findings seem significant, and to what extent?
\end{compactenum}
\par\vspace {0.4cm}

Answering the above questions, the concepts were generated according to the 
coding process. They were systematically related through the use of a network 
diagram (see Figure \ref{fig:gt:completemethodology}), by which we established 
the relations between the concepts. At each iteration of the 
grounded theory process, we considered and examined a concept within different 
conditions and dimensions, trying to incorporate data from a broader range of 
practitioners.

All the linkages and categories were constructed by the use of Atlas.TI and 
compared according to the data analysis process. Variations of the generated 
theory were applied according to raw grounded data with the aid of codes and 
memos. Moreover during this study extensive explanation of the concepts were 
presented in-vivo statements as reported by practitioners.

The process has been designed by the researches in different steps, explaining 
and making clear the purpose of each of them. Thus, the same processes enable 
other research to conduct the same study within similar contexts. Moreover 
\textit{comparison with the state-of-the-art} has been applied in order to 
validate the theory and enhance odds of its applicability within a wider 
time-frame. By this comparison we  highlight the areas which have been neglected 
by existing studies, providing possible directions for future studies (see 
subsection \ref{sect:theory:validation:others}). Moreover we study the 
confounding factors which could interfere with the application of the GSM model 
(see subsection \ref{sect:an:comp:cat-vs-literature:conf}).

Then, we performed an \textit{evaluation with empirical data} (see subsection 
\ref{sect:theory:validation:stats}), quantifying scores for each company 
involved in our empirical study by defining metrics based on a set of rubrics 
and evaluating startups accordingly, through an analysis of interview 
transcripts and codes. The complete statistical procedure and provided  rubrics  
to quantify the measures are illustrated in Appendix.

%------------------------------------------------------------------------------%

\section{Results: Green-field startup model}
\label{res:gsm}

By analyzing the interview transcripts with the GT process we extracted the 
information necessary to form the theoretical model. The initial coding process 
(open and in-vivo) provided 630 unique codes\footnote{Available at 
\url{https://github.com/adv0r/BTH-Interview-Package}.}, grounded to 1295 
citations in the transcripts, divided according to the thematic areas. The 
thematic areas were divided according to the development phases designed for the 
interview process. The raw codes constitute the empirical foundation of the 
theoretical categories, which has been used to form the Green-field startup 
model (GSM).

The GSM captures the underlying phenomenon of software development in 
early-stage startups. The framework is formed by 128 sub-categories, clustered 
in 35 groups, and finally in 7 categories at the highest level of abstraction. 
By the means of the GSM we provide explanations of the development strategies 
and engineering activities undertaken by startups.
%------------------------------------------------------------------------------%
\subsection{Framework overview}
\label{res:gsm:frmov}
The main concepts representing the underlying phenomenon have been grouped 
together to form high-level categories. Figure \ref{fig:gsm} shows the network 
of causal relationships (represented by arrows) between categories (represented 
by blocks). Each node is linked to subsequent nodes which are called successors, 
and preceding nodes, called predecessors. The relations between them are denoted 
by the direction of the interconnecting arrows. The network is read from left to 
right.

\begin{figure*}[!t]
\centering
\includegraphics[width=4in]{figures/high-level}
\caption{Green-field startup model}\label{fig:gsm}
\end{figure*}

In forthcoming the explanation of the GSM we emphasize labels of theoretical 
categories using the \textit{italic} font style. For the main categories shown 
in Figure \ref{fig:gsm}, when necessary we also make use of identifiers (i.e. 
CATx) to facilitate the reading of this section.
The network is centered around the core category, \textit{speed-up 
development}, which is the most interconnected node in the framework reflecting 
the fact that \textit{``it is the one (category) with the greatest explanatory 
power''}  \cite{Strauss1998}. The meaning of the labels on the three arrows that 
reach the core category (efficiency, effectiveness and performance) is discussed 
in detail in section \ref{res:val}. %MUN: this refers to the validity threats 
%section. Do you mean rather the "Theory Validation" section?
%CG:fixed
%MUN1: this should be \ref{sect:theory:validation:stats}, right? Section 5 is 
%quite long and you should point to the correct subsection

A contextual condition which characterizes, to some extent, each and every 
startup, is the \textit{severe lack of resources}. In fact, the capabilities of 
an early-stage startup to support its development activities are constrained by 
an extremely limited access to human, time and intellectual resources, 
eventually leading to a general absence of structures and processes.

The \textit{severe lack of resources} forces the company to focus on 
implementing an essential set of functionalities and it is one of the main 
reasons why the \textit{product quality has low priority} respect to other more 
urgent needs\footnote{There are some exceptions where the quality aspects 
actually matter and such cases will be discussed in section \ref{res:val}%MUN: 
%this refers to the validity threats section. Do you mean rather the "Theory 
%Validation" section?
%CG fixed
%MUN1: Same here, point to the correct subsection.
.}. At the same time, to be able to deal with such constraints, startups must 
depend on a small group of capable and motivated individuals.

As it has been unanimously expressed by respondents, from a software 
perspective, the most urgent priority is to \textit{speed-up the development} as 
much as possible by adopting an extremely flexible and effective 
\textit{evolutionary approach}. On the other hand the efficiency of teamwork is 
facilitated by the low attention initially given to architectural aspects 
related to product quality. This allows startups to have a faulty but 
functioning product, which can quickly be introduced to the market beginning 
with the in-house implementation of the prototype from day-one.

The initial employees are the ingredients which enable high levels of 
performance in software development. To support a fast-paced production 
environment, engineers are required to be highly committed, co-located, 
multi-role, and self-organized. In other words \textit{team is the catalyst of 
development}.
With an essential and flexible workflow, which relies on tacit knowledge 
instead of formal documentation, startups are able to achieve very short 
time-to-market cycles. However, each line of code, written without following 
structures and processes, contributes to grow the \textit{accumulated technical 
debt}, which is further increased by having almost non-existing specifications, 
a minimal project management and a lack of automated tests.

Despite the fact that the consequences of such debt are not clearly perceived 
in the initial stages (where finding the product/market fit as quickly as 
possible is the most important priority), startups, which survive to subsequent 
phases, will likely increment their user-base size, the product size, and the 
number of developers. This will require the company to eventually pay the 
\textit{accumulated technical debt}, confronting the fact that an 
\textit{initial growth hinders productivity}.

In the following subsections we explain in detail all the categories presented 
in Figure \ref{fig:gsm}, and conclude in subsection \ref{res:gsm:th} with the 
final theory. In the explanations we use identifiers of the companies presented 
in Table \ref{t_interviews-stats} (i.e. Cx) to highlight statements recorded 
from the interviewees.

%------------------------------------------------------------------------------%
\subsection{Severe lack of resources}
\label{res:gsm:cat7}
The concept of \textit{severe lack of resources} characterizes the uncertainty 
of the development strategies in startups and it is composed by three 
subcategories: \textit{time-shortage}, \textit{limited human resources} and 
\textit{limited access to expertise}.

Since startups want to bring the product to market as quickly as possible, the 
resource they are the most deprived of is time. Startups operate under a 
constant time pressure, mainly generated by external sources 
(\textit{investor(s) pressure}, \textit{business pressure}) and sometimes 
internal necessities such as \textit{internal deadlines} and \textit{demo 
presentations at events}. In this regard, C3 commented: \textit{``Investors 
wanted to see product features, engineers wanted to make them better. Finally 
the time-to-market was considered more important and the team interests was 
somehow sacrificed.''}

In addition, to compensate the \textit{limited human resources},  practitioners 
need to empower \textit{multi-role and full stack engineers}, as confirmed by 
C1: \textit{``Everyone was involved in any tasks, from mobile to web 
development, organizing themselves in choosing the part to implement''}.

The extent to which startups have access to specialized knowledge - both 
internal and external to the company - is extremely reduced when compared to 
traditional established software companies. Therefore, in order to partially 
mitigate the \textit{limited access to expertise}, startups rely on the external 
aid of mentors or advisors. Under these strict limitations, most of the 
decisions related to software development are fundamentally trade-off 
situations (confirming Himola's results \cite{Hilmola2003}).
%------------------------------------------------------------------------------%
\subsection{Team is the catalyst of development}
\label{res:gsm:cat4}
Among the different aspects fostering the speed of the development process, the 
startups' main focus is on the characteristics of the initial team.

In startups \textit{developers have big responsibilities}. In fact, the 
\textit{limited human resources}, discussed in CAT7, cause the team-members to 
be active in every aspect of the development process, from the definition of 
functionalities to the final deployment.

Early engineers in the founding team of a startups are typically 
\textit{multi-role and full-stack engineers}: they are full-stack for being able 
to tackle different problems at different levels of the technology stack 
(\textit{generalists developers instead of specialists}) and multi-role since 
usually \textit{engineers are responsible for marketing/sales/development}. As 
remarked by  C11: \textit{``Instead of hiring gurus in one technology, startups 
should hire young developers, generalists, that know how to quickly learn new 
technologies, and quickly move among a huge variety of tasks.''}

Moreover, having a \textit{very small and co-located development team} enables 
members to operate with high coordination, control and communication, which 
heavily rely on \textit{tacit knowledge},  replacing most of the documentation 
with informal discussions. Practitioners reported that keeping the development 
team small helps startups in being fast and flexible, as remarked by C8: 
\textit{``If you have more than 10 people, it is absolutely impossible to be 
fast''}. Then, also \textit{previous working experience} and \textit{knowing 
each other before starting the company} enforce the efficiency of activities by 
\textit{limiting the need of formalities between team-members}.

In every software company, \textit{skilled developers are essential for high 
speed} in development. Especially in startups, the ``hacking culture'' and a 
tendency to the ``just-do-it'' approach allow the team to quickly move from the 
formulation of a feature idea to its implementation. In this regard, C1 
comments: \textit{``We had a hacker culture/environment, people hacking stuff 
without formally analyzing it, but breaking it down and finding a way around. 
Ticket on trello and go.''}

A \textit{limited access to expertise} forces the team to rely mainly on their 
personal abilities, even though asking opinions to mentors is reported to be a 
viable practice to maintain feasible objectives. Furthermore \textit{teams work 
under constant pressure}, \textit{``up to 16 hours per day''}, %MUN: add Cx to 
%this statement since it is a citation? 
%CG: not fixed. It represents a subcategory, expressed by the 70% of the 
%interviewed startups
%MUN1: and how should the reader know that? See my problem with italics: you 
%use italics for both interview statements and (sub-)categories. 
mainly constrained by a tight \textit{time shortage}.

Finally, startups present founders-centric structures, and especially in the 
early-stage, it is clear that \textit{CTO/CEO background has high-impact} on the 
company's development approach. For instance, in case of an academic background, 
the CTO encourages the introduction of some architectural design before the 
development phase. Moreover, even though the CTO/CEO starts guiding initially 
the development process, most of the decision are taken consensually by all 
members of the team. Then, the CTO/CEO decides only in particular situations 
where some conflicts occur.
%------------------------------------------------------------------------------%
\subsection{Evolutionary approach}
\label{res:gsm:cat2}
Among different approaches to development, startups prefer to build an initial 
prototype and iteratively refine it over time, similarly to the concept of 
\textit{``evolutionary prototyping''} \cite{EvProt}. The reason of using this 
approach is that startups want to validate the product against the market as 
soon as possible, finding the proper product/market fit. Indeed, they can focus 
on developing only parts of the system they want to validate instead of working 
on developing a whole new system. Then, as the prototype is released, users 
detect opportunities for new functionalities and improvements, providing their 
feedback to developers.

Since \textit{flexibility and reactiveness are the main priorities}, the most 
suitable class of software development approaches are clearly highly 
evolutionary in nature. In fact, as \textit{uncertain conditions make long-term 
planning not viable}, startups cannot base their work on assumptions without 
rapidly validating them, releasing the product to market. The uncertainty is 
first of all in the team composition. Since the teams are typically very small 
and the project knowledge is not documented, even a little change in their 
composition (e.g. a developer falls ill) can have a very high impact on the 
overall product development. Furthermore, startups operate in a continuous 
evolving environment in terms of competitors and targeted market sectors. Then, 
to obtain a competitive advantage in the market, startups typically make use of 
cutting-edge solutions, characterised by an evolution that cannot be foreseen in 
the long run. However, a special role in daily decisions is covered by user 
feedback and requests, which represent the main drivers for defining the product 
features in the short-term.

To obtain fast user responses and quickly \textit{validate the product}, 
startups \textit{build a functioning prototype and iterate on it} over time. 
Paraphrasing C4, \textit{``[\ldots] you should start with something that is 
really rough and then polish it, fix it and iterate. We were under constant 
pressure. The aim was to understand as soon as possible the product market/fit 
iterating quickly, adjusting the product and releasing fast.''}

The companies focus on building a small set of functionalities to include in 
the first version, and \textit{progressively roll-out to a larger number of 
people} with very \textit{small iterations} (confirmed again by C4: \textit{``we 
deploy from 5 to 20 times a day''}).

The objective of this evolutionary approach is to avoid wasting time in 
``over-engineering the system'' and building complex functionalities that have 
not been tested on real users. By releasing a very small number of good-enough 
functionalities (see CAT3) the startup can verify the suitability of the 
features and understand how to adjust the direction of product development 
towards actual users' needs. The first version of the product is typically a 
prototype which contains the basic functionalities developed with the least 
possible effort that can validate critical features, enabling the startup's 
survival in the short term (recalling  Ries' definition of \textit{Minimum 
Viable Product} (MVP) \cite{Ries2011}).

Supported by \textit{direct contact and observation of users}, 
\textit{automated feedback collection} and \textit{analysis of product metrics}, 
startups attempt to \textit{find what is valuable for customers} with activities 
in agreement with Blank's \textit{Customer validation} stage \cite{Blank2005}.

%------------------------------------------------------------------------------%
\subsection{Product quality has low priority}
\label{res:gsm:cat3}
The interests of software startups, related to the product, are mainly 
concentrated on building a  \textit{limited number of suitable functionalities} 
rather than fulfilling restrictive non-functional requirements. This strategy 
allows them to quickly release simple products with less need of preliminary 
architectural studies.

Exploring quality aspects considered during the development process, the only 
important concerns are expressed towards the user experience\footnote{According 
to the ISO 9241-210 (Ergonomics of human-system interaction), UX is defined as 
\textit{``a person's perceptions and responses that result from the use or 
anticipated use of a product, system or service''.} Even though this definition 
leaves so much to interpretation, we can refer to it as presented in the 
glossary.}, in terms of \textit{ease of use, attractiveness of the UI} and most 
of all, \textit{smooth user-flow without interruptions}. The fact that 
\textit{UX is the only important quality} is remarked by C11: \textit{``When a 
user needs to think too much on what action should be done next, he will just 
close the application without returning''} and confirmed by C3: \textit{``If the 
product works, but it is not usable, it doesn't work''.}

The extent to which quality aspects are taken into account might heavily depend 
upon the market sector and the type of application. Nevertheless, realizing high 
level of UX  is often the most important attribute to consider for customer 
discovery of  evolutionary approaches in view of the \textit{limited human 
resources} and \textit{time shortage}, presented in CAT7. C4 confirms: 
\textit{``None of the quality aspects matter that much as the development speed 
does.''}

To achieve a good level of UX while dealing with lack of human resources and 
time shortages, startups can analyze similar products of bigger companies which 
can afford more rigorous usability studies. Then, the users' feedback and 
product metrics begin to have a central role in determining the level of UX 
achieved. Product metrics are typically web-based \textit{statistical hypothesis 
testing}, such as A/B testing \cite{AB}. 
%MUN: add reference or explain shortly what A/B testing is
%CG: fixed

Other than UX, some other factors can influence the quality concerns of 
development:

\begin{compactitem}

\item The \textit{efficiency emerges after using the product}, letting 
engineers avoid wasting time in excessive improvements of not-validated 
functionalities. The level of efficiency can be optimized after attesting the 
effectiveness of the minimal set of functionalities, obtained according to CAT2 
and to the concept of MVP.
\item The \textit{product should be reasonably ready-to-scale} to be able to 
accommodate a potential growth of the user-base\footnote{Note that startups 
tackle fast growing markets which are particularly subject to sudden user 
growths}. Thanks to modern cloud services, startups can \textit{externalize 
complexity to third party solutions} achieving a good level of scalability with 
a reasonable effort.
\item Realizing high reliability is not an urgent priority since \textit{users 
are fault-tolerant in innovative beta products}. In these cases, users have 
typically a positive attitude towards the product, even though it presents 
unreliable behaviours. In this regard, the main focus of beta testing is 
reducing frictions between the product and the users, often incorporating 
usability testing. In fact, the beta release is typically the first time that 
the software is available outside of the organization that developed it. 
\footnote{Detailed discussion of the impact of innovative products on the user 
satisfaction is presented in subsection \ref{sect:an:comp:cat-vs-literature:conf}}.
%MUN: do you use the Kano model in Theory Validation? Change text to refer to 
%the correct section. [After I've read the respective section, I'd suggest to 
%remove the Kano discussion altogether to shorten the paper. Keep the footnote, 
%reference the correct section, but remove "according to the Kano model]
%CG: fixed
\end{compactitem}

Finally, startups implement a \textit{limited number of suitable 
functionalities}, which in view of \textit{limited human resources} and 
\textit{time shortage}, represent a viable strategy to focus on shortening 
time-to-market.
%------------------------------------------------------------------------------%
\subsection{Speed-up development}
\label{res:gsm:cat1}
As remarked initially, \textit{speed-up development} represents the core 
category of the GSM. Firmly grounded as the primary objective of startups, it 
shows the most important characteristic of developing software in the early 
stage.

As mentioned before, in order to \textit{speed-up development}, startups adopt 
evolutionary approaches supported by a solid team focusing on implementing a 
minimal set of suitable functionalities. Startups \textit{keep simple and 
informal workflows} to be flexible and reactive, adapting to the fast changing 
environment. While a rigorously defined workflow generally consists of an 
established sequence of well-defined steps to follow, startups, by contrast, 
adopt informal workflows. This choice is mainly dictated by contextual factors 
characterized by the general conditions of unpredictability and uncertainty. The 
adoption of informal workflows is facilitated by the fact that teams are 
typically self-organized and \textit{developers have big responsibilities} .

Additionally, possible planning activities are refrained by the aim to shorten 
time-to-market, as reported by C8: \textit{``Speed was the essence so we didn't 
plan out too many details''}. To deal with such unpredictability, startups 
prefer to take decisions as fast as possible, mainly by means of informal and 
frequent verbal discussions.

Despite Agile principles appear to be suitable to embrace changes, especially 
in the early-stage, development practices are often perceived as time-wasters 
and ignored to accommodate the need of releasing the product to market as 
quickly as possible. In this regard, to maintain the simplest and most informal 
approach, startups aim to \textit{find the product/market fit quickly} and 
implement a \textit{limited number of suitable functionalities}. This approach 
is considered possible also in view of a lack of systematic quality assurance 
activities, since only the user experience is considered as important and other 
quality aspects, such as efficiency, can be postponed after the first release.

Another strategy, which supports startups in quickly delivering products, is 
the \textit{externalization of complexity on third party solutions} by: making 
use of third party components (COTS) and open source solutions (both for 
product components, and development tools and libraries); taking advantage of  
external services; and deploying the product on external infrastructures, for 
the sake of delivering a \textit{product reasonably ready to scale} for a 
possible future growth.

Even though \textit{the use of well-integrated and simple tools} allows 
startups to automate many activities and reduce their completion time, drawbacks 
of such approaches are the increased concerns of interoperability issues. A 
category of tools is represented by advanced version control systems, which are 
not only used to manage the code-base, but for many other purposes such as: 
assigning tasks; trace responsibilities; configuration management; issue 
management; automatic deployment; discussions and code reviews.

Startups can further improve development speed by making \textit{use of 
standards and known technologies} which are widely recognized, well tested, and 
supported by strong communities. Moreover, the use of widely adopted standards 
and frameworks reduces the need of a formal architectural design since most of 
implementation solutions are well documented and ready-to-use. This is confirmed 
by C1: \textit{``as long as you use Ruby standards with the Rails framework, the 
language is clean itself and doesn't need much documentation''.}

Other important factors that positively impact the speed of development are the 
team's desire to: \textit{create disruptive technologies}; to 
\textit{demonstrate personal abilities}; and to \textit{have the product used in 
the market}. As reported by practitioners, these factors are essential to 
enhance the morale of developers and therefore to achieve higher team 
performances. By contrast, especially in the typical sprint-based environments 
of Agile, when a developer is not \textit{able to meet deadlines} the morale 
goes down, hindering the development speed.

Finally the constant pressure under which the company regularly operates, leads 
the team to often \textit{work overtime to meet deadlines}. But as reported by 
practitioners, such way of working can be an effective strategy only in the 
short term since it can lead to poorly maintainable code and developer burnouts 
in the long run.
%------------------------------------------------------------------------------%
\subsection{Accumulated technical debt}
\label{res:gsm:cat5}
Startups achieve a very high development speed by radically ignoring aspects 
related to documentation, structures and processes. However, although these 
aspects are not considered important in the very first stages, they will become 
increasingly more important for the productivity in the long-term, as we 
illustrate in CAT6.

Instead of traditional requirement engineering activities, startups make use of 
\textit{informal specification of functionalities} by means of ticked-based 
tools to manage low-precision lists of features to implement, written in form of 
self-explanatory user stories \cite{AgilePlan}. Practitioners intensively use 
physical tools such as post-it notes and whiteboards, which help in making 
functionalities visible and prioritizing stories based on personal experiences. 
C4 comments \textit{``[\ldots] it is the only way. Too many people make the 
mistake of sitting down and write big specs and then they build it for four 
months, realizing the product is not valuable only at the end.''}

Since startups are risky businesses by nature, even less attention is given to 
the traditional phase of \textit{analysis}, which is replaced by a \textit{rough 
and quick feasibility study}   \textit{``only sometimes and however 
informally''}, as stated by C5.  First of all it is generally \textit{hard to 
analyze risks with cutting-edge technologies}. To find out the feasibility of 
such cutting-edge projects, startups attempt a first implementation with rough 
and informal specifications, assuming that project's complexity will remain 
limited to a small number of functionalities, as discussed in CAT3. Additionally 
by \textit{keeping the product as simple as possible} and learning from 
competitors' solutions and mistakes, practitioners can use their \textit{past 
experiences in similar contexts to help assessing feasibility} of the project. 
Finally, to avoid restrictions on the flexibility of the team, potential 
limiting decisions are taken only when strictly necessary and anyhow as late as 
possible.

Another important factor which contributes to \textit{accumulate the technical 
debt} is the general \textit{lack of architectural design}, substituted by 
\textit{high-level mockups and low-precision diagrams}, \textit{using 
modularization and frameworks from day 1} and \textit{describing critical 
interactions with third-party components only}. In particular, the use of 
well-known standards, frameworks and conventions limits the need of formal UML 
\cite{UML} diagrams and documentation, and provide a minimum level of 
maintainability costs. Additionally, since quality aspects are not a main 
concern (see \textit{limited number of functionalities} and \textit{efficiency 
emerges after using the product} in CAT3) having a well-structured architecture 
remains a secondary priority, and design is conducted only when strictly 
necessary.

A similar attitude towards verification and validation brings startups to a 
severe \textit{lack of automated testing}, which is mainly replaced by manual 
smoke tests executed by: trying the product internally; seeking for critical 
faults; and letting early adopters report non-critical bugs.  Paraphrasing C3, 
\textit{``Trying the product internally allows us to get rid of 50\% of bugs of 
important functionalities. Meanwhile users report bugs of secondary 
functionalities, eventually allowing us to mitigate the lack of testing. Indeed, 
staying one week in production enables us to identify 90\% of bugs''.} The lack 
of complete automated tests is partially also motivated by the fact that 
\textit{users are fault-tolerant in innovative beta products} and  by the 
limited number of functionalities, which allow the team to easily control 
critical bugs. However, in certain cases where components of the system might 
cause loss of data or severe damages to the product or users, engineers realize 
a reasonable level of automatic testing. In such cases, aided by modern 
automatic tools, they can quickly assess the status of the system integration as 
they add new functionalities to the product.

Furthermore, a rigid project management is perceived as \textit{``waste of 
time''} which hinders the development speed since the \textit{uncertainty makes 
formal scheduling pointless} (C9  reports that \textit{``initial chaos helps to 
develop faster''}). Startups' \textit{minimal project management}  is supported 
by keeping: \textit{internal milestones short and informal}, 
\textit{low-precision task assignment mechanism}  and a very low attention for 
project metrics (paraphrasing C13, \textit{``the only track of progress was made 
by looking at closed tickets''}). In this context \textit{only a final release 
milestone is viable}, which helps practitioners to remain focused on short term 
goals and put new features in production. To further \textit{speed-up 
development}, startups \textit{simplify issue management by integrating it with 
feature management}. Then, the \textit{limited number of functionalities} and 
the {use of standard/known technologies} with a \textit{simple workflow} are the 
main reasons for 
not establishing a heavy project management process since the project and 
technologies can be autonomously managed by the development team (with 
characteristics of: co-location, self-organization and very small team size). 
%MUN: do you mean small team size?
%CG: fixed

Finally, one of the categories, which most contributes in growing the 
\textit{accumulated technical debt}, is the substantial  use of informal and 
verbal communication channels on a daily basis. The high co-location and the 
fast paced development approach increase the volume of \textit{tacit knowledge} 
and severe lack of any kind of documentation.  C4 observes in this regard that:  
\textit{``[\ldots] the issue of having documentation and diagrams out of the 
source code is that you need to update them every time you change something. 
There is no time for it. Instead, there is a huge pay off in having a code that 
is understandable itself.''}
This approach is supported by the fact that startups make use of  
\textit{simple and informal workflow}, \textit{standard/known technologies}, 
\textit{very small and co-located development team} with \textit{limited need of 
formalities}.
%------------------------------------------------------------------------------%
\subsection{Initial growth hinders performance}
\label{res:gsm:cat6}
The lack of attention given in the first phases to engineering activities 
allows startups to ship code extremely quickly (see CAT5). However, if 
successful, the initial product becomes more and more complex over time, the 
number of users increases and the company size starts to grow. Under these 
circumstances the need of controlling the initial chaos forces the development 
team to return the \textit{accumulated technical debt} instead of focusing on 
new users' requests. Hence, \textit{initial growth hinders performance} in terms 
of number of new user stories implemented in a certain amount of time, which 
brings new suitable functionalities to the users.

When the number of users increases, customers become more quality demanding and 
also scalability issues start to arise. Usually, \textit{company and user size 
grow} when new business events occur, such as: a \textit{new round of funding}, 
a possible \textit{acquisition}, the release of a \textit{competing product on 
the market}, or when the project is \textit{open for the first public release}. 
As a consequence, the \textit{increasing number of users} creates a growing 
number of requests and expectations. Therefore, whereas the project lacks of 
minimal processes, suddenly, \textit{the current team is not able to manage 
increased complexity} of new functionalities to implement and maintain the 
codebase.

Subsequently, practitioners start considering the need of project management 
activities, also in view of \textit{hiring new staff members}, as discussed by 
C13: \textit{``(project management) is strictly necessary if you radically 
change the team or when the team grows.  The informal communication and lack of 
documentation slow down the process afterwards''}.  Project management becomes 
further important when the \textit{focus shifts to business concerns}. Part of 
the effort, which was initially almost entirely dedicated to product 
development, is required to move to business activities. Moreover the 
availability of project information becomes an important issue because of the 
accumulated \textit{tacit knowledge}, which hinders the ability of new hires to 
quickly start the development of project tasks.

When the company faces growth, the partial and informal engineering activities 
that have been conducted during the first phases of \textit{rush} development 
(refer to CAT5 subcategories \textit{minimal project management}, 
\textit{informal specification of functionalities}, \textit{rough and quick 
feasibility study}, \textit{lack of automated testing}, \textit{tacit 
knowledge}), force the company to \textit{pay off the accumulated technical 
debt}. Under these circumstances, startups need to be able to cope with the 
renewed needs and expectations of both the company (internal necessities) and 
customers (product).

Another factor that slows down performance is caused by the fact that 
\textit{portion of codes needs to be rewritten} and \textit{substantial 
refactoring of the codebase} is required by increasing product demands.

Practitioners realize that some decisions taken (or not taken) during the 
\textit{rough and quick feasibility study} before starting the implementation, 
have led to negative consequences on the long term performance and 
maintainability of the product. Additionally, the initial decisions of the 
product design might not be able to satisfy the increased demands of the 
product's users and developers (\textit{lack of architectural design}). The 
combination of these factors leads to the need of \textit{re-engineer the 
product}.

By re-engineering the systems, startups aim to \textit{increase the scalability 
of the product/infrastructure} and starting to \textit{standardize the codebase 
with well-known frameworks}.

C7 reports that: \textit{``To mitigate this (lack of frameworks) I had to make 
a schema for other developers when we hired them. We had to do a big 
refactoring of the codebase, moving it from custom php to Django, normalizing 
the model and making it stick with the business strategy. I had the code in 
different php servers communicating via JSON, some engineering horror. Now  that 
we are fixing it, it's really painful. We had to trash some code. However I 
don't regret that I didn't make this choice sooner, it was the only way''}.

The \textit{fear of changing a product, which is working}, arises when product 
complexity increases. The changes to the codebase, to support bug fixing, become 
highly interrelated with other functionalities and difficult to manage because 
the product is poorly engineered (see CAT5). Therefore, the fear arises from 
changing a validated product that might bring changes to the user responses.

The increasing number of feature requests causes the \textit{growing necessity 
of having a release plan}, rationalizing the rollout of new features.

Therefore, startups begin to \textit{partially replace informal communication 
with traceable systems} and \textit{introduce basic metrics for measuring 
project and team progress} in order to establish an initial structured workflow. 
In conclusion, because the increasing number of users causes issues in 
scalability and reliability, the \textit{need of more structured workflow} 
becomes of major relevance. C11 states: \textit{``Yet, it is still better to 
have a reasonable drop-down in performance when team grows than lose time in the 
beginning''}.
%------------------------------------------------------------------------------%
\subsection{Theory}
\label{res:gsm:th}
In order to explain and understand the development strategies in early-stage 
software startups we construct the theory generated and supported by the above 
presented GSM:

\begin{theory}
Focusing on limited number of suitable functionalities, and adopting partial 
and rapid evolutionary development approaches, early-stage software startups 
operate at high development speed, aided by skilled and highly co-located 
developers. Through these development strategies, early-stage software startups 
aim to find early product/market fit within extremely uncertain conditions and 
severe lack of resources. Nevertheless, by speeding-up the development process, 
they accumulate technical debt, causing an initial and temporary drop-down in 
performance before setting off for further growth.
\end{theory}

As discussed in section \ref{res:val} %MUN: this refers to the validity threats 
%section. Do you mean rather the "Theory Validation" section?
%CG:fixed
, we formed the theory by applying the paradigm model, introduced by Corbin 
\cite{Corbin1990}. The theory is composed by different elements reflecting the 
paradigm model:

\begin{compactitem}

\item \textit{Causal conditions} are represented by three main conceptual 
categories: \textit{product quality has low priority}, \textit{evolutionary 
approach} and \textit{team is the catalyst of development}.
\item \textit{Phenomenon} is represented by the core category \textit{speed-up 
development}.
\item \textit{Context} is limited to web early-stage software startups 
operating in conditions of severe lack of resources aiming to early find 
product/market fit.
\item \textit{Intervening conditions} are  summarized by the extremely 
uncertain development environment.
\item \textit{Action and interaction strategies} are represented by the 
accumulation of technical debt.
\item \textit{Consequences} lead to a temporary performance drop-down.
\end{compactitem}

%------------------------------------------------------------------------------%
\section{Theory Validation}
\label{res:val}

In this section we discuss the validity of the GSM by means of 
cross-methodological observations, as discussed in section \ref{rm:val}.
%MUN: expand this discussion a bit (motivation why you do this like you do it) 
%and move to Section 3.5. Then write here just that you perform theory 
%validation as discussed in Section 3.5 (and remove the bullet list below).
%CG:fixed
%MUN1: Ok, but fix section 3.5... you see that you don't motivate 5.1-5.4 at 
%all there and that the current content of 3.5 is not related to what you 
%actually do here as validation?

%------------------------------------------------------------------------------%
\subsection{Comparison with other frameworks}
\label{sect:theory:validation:others}

To validate the generalization of the theory, we describe conceptualizations 
defined from the GSM that are in support of previous models developed by 
Coleman \cite{Coleman2007,Coleman2008a, Coleman2008}, Baskerville 
\cite{Internet} and finally Brooks \cite{BrooksJr1987}.

Starting from the Coleman's work, we refer to him since only he has,  to the 
best of our knowledge, conducted a similar research in startups' context, even 
though with a different focus. Coleman investigated factors in software 
development that hinder initiatives of software process improvement (SPI) in a 
later stage, representing also companies in expansion phase with more than 100 
employees. %MUN: so he did not do research 
%in a startup context? You need to clarify what you mean by "more mature 
%companies"... I assume you still mean startups, but startups at a later stage, 
%i.e. after first product release.
%CG: fixed

In this subsection we illustrate the most important aspects considered by 
Coleman, examining similar and contrasting factors with respect to the GSM. We 
consider only those categories strictly related to the core category, which 
relations, as described in GT methodology, generate complete explanatory power 
of the generated theory.

Coleman's framework aims to highlight how managers consider two distinct kinds 
of processes: \textit{essentials} and \textit{non-essentials}. The 
\textit{essential} processes are the most closely linked to product development, 
such as requirements gathering, design and testing. The \textit{non-essential} 
processes are those that might be omitted, such as planning, estimating and 
staging meetings. In particular, he discusses how practices are routinely 
removed: \textit{``With most methodologies and approaches, very few stick to the 
letter of them and they are always adapted, so we adapted ours to the way we 
wanted it to work for us, for our own size and scale''} \cite{Coleman2008}. 
%MUN: add ref to Coleman's paper where he states this
%CG fixed

The GSM explores the same challenges in software startups, where the act of 
tailoring processes leads developers to adopt only minimal practices, which are 
most suitable for the startups context. Furthermore, CTOs' and CEOs' background 
has a great impact on the speed of the development process as described in CAT4. 
This is in accordance to Coleman's framework, which presents the background of 
founders and development managers as main factors that affect the management 
style of the development process.

Differently from Coleman's framework, the GSM doesn't present how 
\textit{market requirements} can affect the conduction of processes. In this 
regard, Coleman describes how the more requirements definitions are predictable 
the more well-defined workflows are established. This particular aspect is 
considered as a confounding factor\footnote{The identification of confounding 
depends on the context in which the study is conducted and on the background 
knowledge of the researchers \cite{Pearl2011}.}  in our study since we were not 
able to ground this concept in the data.

But what characterizes Coleman's network is the \textit{cost of process} (core 
category) and all the factors that in management contributed to the lack of 
software process improvements (SPI). The \textit{cost of process} represents the 
 lack of formal and prescriptive workflows in development, mainly conducted by 
means of verbal \textit{communication} limiting heavy \textit{documentation} and 
\textit{bureaucracy}.

Coleman reports the practitioners' perception that \textit{documentation} alone 
would not have ensured a complete shared understanding of project requirements. 
Moreover, defined processes are perceived by managers as additional items with a 
negative impact on the \textit{Creativity} and \textit{Flexibility} of the 
development team. This is in accordance with our generated theory, which bases 
the reasons of adopting evolutionary and low-precision engineering elements in 
the flexibility and reactivity attributes of the development process in 
startups.

In this regard we can paraphrase a comment recorded by Coleman during an 
interview: \textit{``When we set up we had more supervisory and managerial roles 
in that group than we have now and we had to scale that back which has made 
things a lot more flexible. I do think you have to be nimble, quick and capable 
being responsive in our position. That works well and I don't want to lose 
it.''} \cite{Coleman2007}.

Verbal \textit{communication}, lack of heavy \textit{documentation} and 
\textit{bureaucracy} can easily be associated to the \textit{accumulated 
technical debt} category (see CAT5). According to Coleman's framework, they 
describe how startups experience the lack of main engineering activities and 
documentation. Additionally, the \textit{speed-up development} category 
expresses the same concept of \textit{flexibility} and \textit{process erosion} 
since they have direct relation to the subcategory of \textit{keep a simple and 
informal workflow}, as discussed in CAT1.

As also reported in the GSM, the definition of a \textit{``minimum process''} 
is not a matter of poor knowledge and training, but rather it is the necessity 
of operating with solutions that let the company move faster. 
\textit{``One-size-fits-all''} solutions have always found difficulty in 
penetrating small software organizations \cite{Staples2007}. When startups began 
establishing any SPI process, they experienced \textit{process erosions}  
\cite{Coleman2008} %MUN: 
%add ref (Coleman?). You should add, for each category you use from Coleman, 
%add the reference to the paper. This makes it clear that the category comes 
%from him (and is not one of our categories)
%CG: fixed
, which resolved to a barely sufficient workflow to satisfy the organizational 
business needs.

Then, software startups favor the use of agile principles in support of 
\textit{creativity} and \textit{flexibility} instead of SPI methods, whereby 
processes need to be predictable and repeatable. Nimble and ad-hoc solutions 
prevent the use of heavy bureaucracy and formal communication strategies %MUN: 
%seems to contradict each other: "ad-hoc solutions prevent the use of heavy 
%bureaucracy AND fast communication strategies" I agree that ad-hoc solutions 
%do not produce heavy bureaucracy, but they allow fast communication 
%strategies... I may be wrong, but then you need to explain it better
%CG:fixed, you are right :) ...  
, even though the accumulated tacit knowledge is hard to manage and to transfer 
to new hires as also discussed in category \textit{CAT6}.

Coleman describes how the \textit{``management approach''} is oriented towards 
\textit{``embrace and empower''} solutions in contrast with \textit{``command 
and control''} style, where there is an evidence for trust in the  development 
staff to carry out tasks with less direct supervision, greater delegation of 
responsibilities and a more generally consensual environment \cite{Coleman2008}. 
Nevertheless, software development managers and founders have still impact on 
management style and indirectly on the software development process. In the case 
of early-stage startups, founders are mainly software development managers as 
CEOs/CTOs and  technical practitioners at the same time. As Coleman identified 
the influence of the founders' and managers' background on the software 
development process, the GSM similarly identifies that the CEOs and CTOs 
background shapes the high-level strategies adopted in developing the initial 
product. Notwithstanding, team members remain self-organized, able to intervene 
in all the aspects of the development process without any direct supervision, as 
discussed in CAT4.

In conclusion, by the comparison of the GSM with Coleman's framework, we have
found that despite Coleman's study contributes to define which factors hinder
the introduction of SPI in startups 
%MUN: add: "of SPI in startups"? CG: fixed
his results are coherent with the theory generated within our research, focusing
at explaining the overall strategies within software development in startups.
%MUN: add "our research, focused at explaining software development in
%general in startups." ? CG:fixed
In this regard, a wide variability of processes is a key factor for startups.  
In contrast with some SPI processes, such as \textit{Six Sigma} which objective 
is minimizing unpredictability in the definition of workflows \cite{Sixsigma}, 
startups seem to follow the \textit{Fischer's Fundamental Theorem of natural  
selection} \cite{Fisher}.
Moving towards flexible and variable processes (easy  to adapt to the uncertain
conditions), increase the odds of ``natural  selection'' only when some
``restrictive conditions'' are met. In other words, even though ignoring SPI
practices is advisable, under circumstances of high uncertainty and severe lack
of resources, %MUN: under which  %circumstances? This statement sounds too
%general, i.e. don't do SPI ever... %CG: fixed 
%And the following part,I did not fully understand. Try to clarify. CG:fixed  
not following SPI does not predict
major adaptiveness of startups development towards the customer's needs.
Nevertheless, to imply a major evolutionary product/market fit, certain
conditions explained in GSM need to be considered.

%MUN: As I wrote already in the previous draft, you need to clarify at the 
%beginning of this subsection which frameworks you are going to compare GSM to. 
%Now, you say that Coleman is the only author who did research, comparable to 
%yours, into startups. Now there comes Baskerville and Brooks. Write at the 
%beginning of this section, shortly, which frameworks/authors you compare GSM 
%to and why, then create subsections for each.

%CG: partially fixed. Agreed on the beginning of the section, not agreed on  the 
%division in subsections. Let's discuss it by a skype call.

Also Baskerville \cite{Internet} refers to SPI as a development approach
typically effective only in large-scale, long-term development efforts with
stable and disciplined processes. Internet-speed software development (i.e. an
agile method oriented using daily builds aimed at developing a product with high
speed) differs from traditional software development. Baskerville, by applying a
grounded theory analysis in 10 companies, found that major causal factors that
influence development are:

\begin{compactenum}

\item A desperate rush-to-market.
\item A new and unique software market environment.
\item A lack of experience developing software under the conditions this 
environment imposed.

\end{compactenum}

As discussed in CAT7, 1) can be mapped to \textit{time shortage} and 3) to 
\textit{limited access to expertise}, whilst  2) can refer to \textit{uncertain 
conditions make long-term planning not viable}, explained in CAT2. Even though 
with different research focus and context of study % - understand how and why 
%Internet-speed software development (i.e. an agile method oriented using daily 
%builds aimed at developing a product with high speed) differs from traditional 
%software development - MUN: put this sentence where you introduce 
%Baskerville's work. %CG: fixed
, Baskerville revealed similar causal factors as the GSM.

Baskerville argues that the dawn of the Internet era has intensified software 
development problems by emphasizing shorter cycle times, stating to:

\begin{compactitem}
\item Speed-up development by releasing more often the software and 
``implanting'' customers in the development environment (similar to the 
\textit{find the product/market fit} concept presented in CAT2). 
\item Make heavy use of simple tools and existing components (similar to the 
concepts expressed in CAT1).
\item Invest time in facilitating development of scalable systems by the use of 
simple but stable architectural solutions (see subcategory \textit{using 
modularization and frameworks from day 1} discussed in CAT5). 
\item Tailor the development process daily according to the intense demands for 
speed. In this regard, he argues that the trend is to skip phases or tasks that 
might impede the ability to deliver software quickly even though producing lower 
quality software.

\end{compactitem}

With a wider research focus, Brooks \cite{BrooksJr1987} discussed what 
challenges are involved in constructing software products. Brooks divides 
difficulties in development into essence (inherent to the nature of the 
software), and accidents (difficulties attending software production but which 
are not inherent). In other words essence, concerns the hard part of building a 
software through activities such as specification, design, testing. Accidents 
refer to the labor of representing the software or testing its representation.

Brooks claims that the major effort applied by engineers in the last decades 
was dedicated towards accidents problems, trying to exploit new strategies to 
enhance software performance, reliability and simplicity of development, such as 
the introduction of high-level languages for programming.  Despite the great 
achievements in improving development performance, the \textit{essence} property 
of the software remained unaltered.

Essence difficulties are inherent to: complexity, conformity, changeability and 
invisibility. Complexity  mainly refers to interaction of software modules and 
elements between each others in some non-linear fashion, which increases 
accordingly to the project size. Conformity refers to the duty of software to 
adapt to human institutions and systems. Changeability  refers to the constant 
pressure of software products to accommodate culture, market, laws and hardware 
transitions. Invisibility  refers to the difficulty of representing software in 
space.

Here we present the basic attacks (or mitigation strategies) presented in 1986 
by Brooks on the conceptual essence ascertained within the GSM model 
description:

\begin{compactitem}

\item Buy versus build-  the most radical possible solution for constructing 
software is not to construct it at all, taking advantage of what others have 
already implemented. It is the main strategy, which enables startups to 
externalize complexity to third party solutions explained within CAT1,  
\textit{speed-up development}.
\item Requirements refinement and rapid prototyping- avoid deciding precisely 
what to build but rather iteratively extract and refine the product requirements 
with customers and users. It represents the evolutionary development approach to 
maintain a fast release-cycles during development (CAT2).
\item Incremental development- starting from simple solutions allows 
organizations to early prototype and control complexity overtime. It represents 
the main purpose of focusing on \textit{limited number of suitable 
functionalities} examined in \textit{product quality has low priority}  category 
(CAT3).
\item Great team- people are the center of a software project and it is 
profoundly important to empower and liberate their creative mind. It is the main 
objective of empowering developers' capabilities described in \textit{team is 
the catalyst of development} (CAT4).
\end{compactitem}

Brooks' forecast about software development strategies is revealed to be 
extremely accurate, according to the state-of-practice in modern startups.
%------------------------------------------------------------------------------%
\subsection{Theoretical categories and existing literature}
\label{sect:theory:validation:sms}

In this subsection we report the evaluation of our theory in contrast with the 
main contributions of the studies retrieved and discussed within the SMS 
\cite{SMS}. For each retrieved article we present the main results, mapping them 
to categories in accordance to the GSM model. Table \ref{tab:an:literature_comp} 
shows the main contributions of the 37 identified studies, where an 'X' is 
placed in correspondence to the theoretical category they considered. The table 
is sorted by the number of GSM categories covered by the studies.

\begin{table*}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{SMS \cite{SMS} and GSM} 
\label{tab:an:literature_comp}
\centering
\begin{tabular}{|l||c||c||c||c||c||c||c||c||c|}

\hline
Author (year) & CAT1 & CAT2 & CAT3 & CAT4 & CAT5 & CAT6 & CAT7 & Count & Ref. \\
\hline
Sutton (2000) & X & X & X & X & X & X & X & 7 & \cite{Sutton2000} \\
Kajko-Mattson (2008) & X & X & X & X & X & X & X & 7 & 
\cite{Kajko-Mattsson2008} \\
Crowne (2002) & X & X & X & X & X & X & X & 7 & \cite{Crowne2002} \\
Coleman (2008) & X & X & X & X & X & X & X & 7 & \cite{Coleman2008a} \\
Coleman (2008) & X & X & X & X & X & X & X & 7 & \cite{Coleman2008} \\
Coleman (2007) & X & X & X & X & X & X & X & 7 & \cite{Coleman2007} \\
Carmel (1994) & X & X & X & X & X & X & X & 7 & \cite{Camel1994a} \\
Yoffie (1999) & X & X & X & X & X & X & & 6 & \cite{Yoffie1999} \\
Zettel (2001) & X & X & X & X & & & X & 5 & \cite{Zettel2001} \\
Jansen (2008) & X & & X & & X & X & X & 5 & \cite{Jansen2008} \\
Heitlager (2007) &  &  & X & X & X & X & X & 5 & \cite{Heitlager2007} \\
Deias (2002) & X & X & & X & X & X & & 5 & \cite{Deias} \\
Ambler (2002) & X & X &  &  & X & X & X & 5 & \cite{Ambler2002} \\
Wood (2005) & X &  & X  & X &  &  & X & 4 & \cite{Wood2005} \\
Tingling (2007) &  & X &  & X & X & X &  & 4 & \cite{Tingling2007} \\
Taipale (2010) & X & X &  & X & X &  &  & 4 & \cite{Taipale2010} \\
da Silva (2005) & X & X &  & X & X &  &  & 4 & \cite{Silva2005} \\
Mirel (2000) & X &  & X & X & X &  &  & 4 & \cite{Mirel2000} \\
Midler (2008) & X & X &  & X &  &  & X & 4 & \cite{Midler2008} \\
Tanabian (2005) & X &  &  & X &  &  & X  & 3 & \cite{Tanabian2005} \\
Stanfill (2007) & X &  &  &  & X &  & X & 3 & \cite{Stanfill2007} \\
Mater (2000) & X & X &  &  &  & X &  & 3 & \cite{Mater2000} \\
Kuvinka (2011) & X & X &  & X &  &  &  & 3 & \cite{Kuvinka2011} \\
Deakins (2005) & X & X &  & X &  &  &  & 3 & \cite{Deakins2005} \\
Yogendra (2002) & X & X &  &  &  &  &  & 2 & \cite{Yogendra2002} \\
Wall (2001) & X &  &  &  &  &  & X & 2 & \cite{Wall2001} \\
Su-Chuang (2007) & X &  &  & X &  &  &  & 2 & \cite{Su-Chan2007} \\
Steenhuis (2008) &  &  &  & X &  &  & X & 2 & \cite{Steenhuis2008} \\
Sau-ling Lai (2010) &  & X &  &  &  &  & X & 2 & \cite{Lai2010} \\
Kakati (2003) &  & X &  & X &  &  &  & 2 & \cite{Kakati2003} \\
Himola (2003) & X &  &  &  &  &  & X & 2 & \cite{Hilmola2003} \\
H\"{a}sel (2010) & X &  &  & X &  &  &  & 2 & \cite{Hasel2010} \\
Hanna (2010) & X &  &  &  &  &  & X  & 2 & \cite{Hanna2010} \\
Bean (2005) & X &  &  &  & X &  &  & 2 & \cite{Bean2005} \\
Kim (2005) &  & X &  &  &  &  &  & 1 & \cite{Kim2005} \\
Fayad (1997) &  &  &  & X &  &  &  & 1 & \cite{Fayad1997} \\
Chorev (2006) &  &  &  & X &  &  &  & 1 & \cite{Chorev2006} \\
\hline
Count & 29 & 22 & 13 & 26 & 18 & 14 & 20 & &\\
%MUN: add a row which shows the sum of each column. This shows how much 
%categories are covered overall %CG: fixed
\hline
\end{tabular}
\end{table*}

Only 7 studies out of 37, discuss aspects related to the overall software 
development process in early stage startups. The remaining part of the studies 
is only partially covering the phenomenon described by the GSM. Looking at the 
papers' main contribution%MUN: which results? The papers' main contribution 
%or...? %CG: fixed
, we were able to extrapolate information which confirmed the GSM categories.

We can see how the majority of the retrieved studies (29) mention issues 
related to \textit{speed-up development} (CAT1) which confirms the importance of 
the core category. On the other hand, we can observe that only a limited number 
of studies mentions results related to \textit{product quality has low priority} 
(CAT3), \textit{accumulated technical debt} (CAT5), and \textit{initial grow 
hinders performance} (CAT6), suggesting directions of possible future primary 
studies.

These results confirm the relevance of development teams as widely discussed in 
the SE literature, here mapping 26 studies to \textit{team is the catalyst of 
development} (CAT4). The importance of people has been widely discussed in other 
studies in SE (Cooper \cite{Cooper1986}, DeMarco \cite{peopleware-demarco1999}, 
Coleman \cite{Coleman2004}, Valtanen \cite{Valtanen2008},  Adolph 
\cite{Adolph2011} and Cockburn \cite{people-first-order}), advocating for the 
need of empowering people, which is the element with the greatest impact on 
software development. Humans, being non-linear variables \cite{Human}
%MUN: add ref %CG:fixed
, are unlikely to follow repeatable prescriptive methodologies. As stated by 
Highsmith \cite{Highsmith2000}, \textit{``A drawback of each and every 
methodology is to expect people to behave consistently over time, when is 
clearly not like that''}. Especially in early-stage startups, where the whole 
company overlaps with the development team, the GSM confirms the importance of 
factors related to people.

%MUN: Think of structuring this long subsection. Have a subsubsection for each 
%CAT, i.e. 5.2.1 Severe lack of resources (CAT7)
%I've added to each category you mention the corresponding CATx since that  
%identifier you use in table 2. I did not check if you discuss each and every 
%category. Do you? If yes, ok. If not, why not? 
%CG: partially fixed, the categories CAT3 and CAT6 are not well covered as 
%discussed above.
Starting from the category \textit{severe lack of resources} (CAT7), many 
authors describe startups facing with inescapable constraints defined by 
engineering/business concerns and constant time-pressure \cite{Camel1994a, 
Cugola98softwareprocesses:}. Because of their young and inexperienced working 
conditions, startups present limited resources in terms of management and 
strategic alliances \cite{Sutton2000}.

Crowne \cite{Crowne2002} identifies four main stages of a startup's lifecycle: 
startup, stabilization, growth and maturity. Inexperienced practitioners 
operating in a chaotic environment characterizes the first phase, where there is 
no strategic plan for developing the product. But, moving towards the 
stabilization phase, the product starts to be unreliable and requirements 
unmanageable. Consequently, when startups move to the growth phase, they 
initiate structuring and controlling processes, which are incrementally 
integrated in the development environment. The maturity phase represents the 
moment when the company is ready to introduce process improvement initiatives. 
Crowne's study profoundly corresponds to different conceptualizations in the 
GSM. Starting from the minimal and low-precision engineering elements, startups 
start to grow the accumulated technical debt which requires fast integration of 
more structured and standard processes overtime.
Issues related to low-precision engineering practices during companies' growth 
are discussed in an experience paper by Ambler \cite{Ambler2002}. He reports on 
two case studies of growing Internet startups approaching an initial public  
offering (IPO). %MUN:explain %acronym %CG: fixed
Despite the two companies were facing business and team growth, attitudes 
against the introduction of processes to structure and model the system were 
evident. The need to re-engineer the codebase lead the researchers to suggest a 
tailored version of the Rational Unified Process (RUP) 
%MUN: add reference
%CG: not fixed, not sure if needed since Ambler has the detailed 
%explaination of his process, similar to RUP... anyway a dedicated resource 
%would be an IBM website
development framework. Nevertheless, even with the aid of CASE 
tools\footnote{CASE tools are a class of software that automates many of the 
activities involved in various engineering activities. See 
\url{http://www.unl.csi.cuny.edu/faqs/software-enginering/tools.html} .}, 
startups failed in practicing engineering activities in view of excessive lack of 
flexibility. Developers fast moved to simpler tools such as using a white-board, 
where they were able to provide input and insights, question assumptions, share 
past experiences, and even talk about potential implementation strategies. 
%MUN: Candidate for removal since details can be found in Ambler anyway: "In 
%general the development remained highly iterative, where project teams started 
%to have a little modeling and testing only when necessary. They didn't take 
%the serial approach of completely developing and accepting the requirements 
%model, analysis model and so on. This let them react swiftly and adapt 
%rapidly according to changes into their highly competitive markets. The 
%researchers report \textit{``modeling was streamlined''}, that is modeling 
%what was needed and go straightforward to coding, avoiding analysis 
%paralysis." %CG: ok

The use of whiteboards in working environments supports \textit{use of 
well-integrated and simple tools}. In both the startups studied by Ambler, 
documentation was minimal, finding the ``sweet spot'' to have just barely enough 
tools and artifacts to meet their needs. Generally, the use of \textit{keep a 
simple and informal workflow} endorses the use of low-precision engineering 
practices which lead to speed-up development initially, but with negative 
effects on the accumulated technical debt of a minimal project management and 
consequently on the final company growth.

Kajko-Mattsson et al. \cite{Kajko-Mattsson2008} investigated a Swedish software 
startup company involved in mobile applications, reporting the following issues: 
a lack of requirements gathering process; release cycles and their length were 
not defined; the releases and their scope were not planned; lack of control over 
the change requests; lack of process control in terms of  absence of any 
documentation to track the status and progress of the process; defective 
releases in terms of lack of testing; poor communication since the communication 
inside the company was informal and not documented at all. In the same study 
they designed a process improvement model, based on three main phases: Evaluate, 
Plan and Change.

%MUN: the whole paragraph needs to be summarized to its essence. Details can be 
%read in the referred paper. 
%CG: partially fixed, removed the detailed process description
In the evaluation phase, the status of the development workflow within the 
company  is analyzed. The researchers documented a state of chaos where any 
initiatives of quantitative methodology assessment would have been a waste of 
time. Therefore, they applied a qualitative methodology to define a model of the 
release management process. 
%They moved towards the plan phase where the 
%researchers decided to develop a model to manage releases, introducing a process 
%composed by: release scope preparation, release planning, release development, 
%system testing, acceptance testing and release deployment. 
Despite the effort spent by the researchers to establish a development workflow, 
practices were not fully conducted even though some marginal benefits have been 
reported: planned major release from four to six months; planned minor releases 
from two to four weeks; encompassing urgent corrective releases from one to five 
days. 
Regarding the previously mentioned problems, only 
partial improvements have been achieved since the major obstacle was to 
motivate employees to change their habits. With perfect hindsight the 
researchers claim that understanding and adapting solutions by solving one 
problem at a time would have increased the benefits. Then, despite the reported 
results, starting the establishment of a simple and informal workflow aided the 
organization to move from the stage of knowing nothing to knowing at least 
something about their development process.

The integration of minimal project management during the development activities 
can foster the control over development productivity and control the 
\textit{accumulated technical debt} (CAT5), but still with particular attention 
to the development workflow status within the company, designing specific and 
tailored solutions.

Prescribed development models, oriented towards SPI, are basically ignored by 
startups, as discussed in \cite{Zettel2001}. 
%MUN: <-- rephrase, not clear to me %CG:fixed
To overcome the need of structure and to accelerate the development process, 
Zettel developed a \textit{``LIghtweight Process for E-business software 
development''} (LIPE). This lightweight development method integrates the XP 
approach with ideas from areas of software measurement for project control and 
process improvement as a compromise between an ad-hoc and a more rigorous 
approach. Zettel argues that LIPE provides scalability of the development 
process with a certain degree of flexibility, omitting parts of prescribed 
practices. Nevertheless, we couldn't identify any empirical evaluation of LIPE 
that supports its suitability in a startup context.

Deakins \cite{Deakins2005} designed a new spiral methodology, constructed  
during a fieldwork with a developer team that had limited resources managing an 
innovative product in volatile e-commerce environment. This methodology 
represents a fast evolutionary approach with flexible processes. 
It aims to obtain customer's feedback as soon as possible for continuous 
product improvement.%MUN: %<-- rephrase, sentence not parsable CG:fixed
Even in this case, the spiral methodology%MUN: which model? CG:fixed
is compliant with the nature of a random, ad-hoc and  \textit{evolutionary 
development approach} (CAT2). Evolutionary development approaches are 
typically more suitable for web applications, as confirmed in 
\cite{Deshpande2001}.
Moreover, they provide practices oriented towards rapid and iterative design 
with complete freedom to the developer team, enabled to operate anywhere within 
the organization.
%MUN: <-- the whole paragraph is not that clear to me. Try to start 
%with the category (team is catalyst), then write about Deakins spiral model
%CG: partially fixed, the focus here is on the evolutionary development approach

Carmel \cite{Camel1994a} reports on the need of flexible and rapid development 
solutions to shorten time-to-market, but even more important, the need of 
high-skilled team developers, which has an impact on \textit{speed-up 
development} (CAT1). He suggests that entrepreneurs need to look for a well 
formed, skilled core development team and not just a set of product ideas and 
features. Moreover the team must be empowered with full-stack and 
self-organization settings, paired with the flexibility of minimal bureaucracy. 
A studied company stated: \textit{``we didn't need weeks or months of detailed 
modeling and documentation, but rather modeling the architecture a little and 
then either exploring strategies by providing users with code or simply starting 
work on the actual software itself''}.

%MUN: To which category does the following paragraph belong? 
%CG: they represent subcategories of speed-up development 
Eventually to externalize complexity of the developing product
%MUN: solutions on what? CG: fixed
, two other studies \cite{Wall2001, Bean2005} suggest to \textit{``reuse the 
wheel, instead of reinventing it''}, taking advantage of open source software 
whenever it is possible, building systems quickly and effectively as a result. 
Even though the principles in leading-edge development processes are the same 
principles of yesteryear, the way in which the fundamentals are applied has 
changed from prescriptive to agile solutions, with less concern on quality 
aspects, also confirmed by Mater \cite{Mater2000}.

%MUN: When you subdivide this sections, as proposed above, finish with a  
%subsubsection "Summary", summarizing the supported you have found in 
%literature for your theoretical categories. Nothing extensive, 3-4 sentences 
%to conclude this part of the validation, similar as below but a bit more 
%explanatory (what does "compliant with the behavioural model of the 
%phenomenon" mean?). 
%CG: not fixed, when I wrote it I tried to have an overview of the main 
%contributions and make it fluently by papers description. Creating subsections 
%would mean to divide papers from different focuses. This is a trade-off where 
%we took a decision in the past. Take it in consideration, but we can discuss 
%it during the next meeting.
%MUN1: Why would creating subsubsections like 5.2.1 Severe lack of resources 
%(CAT7) remove the flow? This subdivision would give the reader clues what the 
%subsubsection is about, he can relate it to Table 2, and provides some 
%navigational support. Now there are nearly 2p of flowing text with now clue in 
%what order/priority the literature is covered. I think you don't even need to 
%rewrite much, just add the subsubsections and a proper summary.

Thus, by summarizing the comparison, we have found that most of the 
contributions of the studies are compliant with the behavioral model of the 
phenomenon we provided.

%------------------------------------------------------------------------------%
\subsection{Confounding factors}
\label{sect:an:comp:cat-vs-literature:conf}

The purpose of this subsection is to identify which confounding factors need to 
be taken into consideration when evaluating the GSM. Confounding factors are 
those variables which and have not been considered in the model and might 
interfere with the theoretical framework positively or negatively 
\cite{ColinRobson2009}. We report those factors that have been explicitly 
identified by the SMS \cite{SMS} %MUN: do you mean the SMS? Then refer to it.
%CG: fixed 
%I remember you mentioned CF in the results (GSM), maybe also refer to the 
%respective sections CG: not parsed.
%MUN1: me neither... forget what I wrote :-)
, related to: creativity and innovation, market requirements and application 
type, and developer experience.

%MUN: Again, structure this into subsubsections according to the above factors. 
%I did it, but check if you agree...
\subsubsection{Creativity and innovation} 
Heitlager et al. \cite{Heitlager2007} contributed to the investigation of 
creativity and innovation in startups. 
%MUN: the rest of this paragraph is too detailed. I'd remove this "The study 
%reports how development approaches oriented to the product in the early-stage 
%life-cycle, called \textit{fluid} have main impact to innovative solutions 
%initially. After a while the need of structuring and controlling the growing 
%company size (along with \textit{transition} and \textit{specific} phases) 
%arises the need of project management activities and software processes for 
%long-term scalability issues."  CG: ok.
%MUN: Please reformulate the first sentence, I cannot parse it. CG: fixed
The proposed model identifies the three lifecycle stages of a startup: 
\textit{fluid}, \textit{transition} and \textit{specific}. At each stage, the 
model presents different development approaches (product-oriented and 
process-oriented), according to the level of innovation achieved. The study 
reports how product-oriented development, in contrast with forced and mature 
process imposition, provides a degree of freedom to the development team, that 
enhances the creativity of developers and augments the innovation capability of 
the company in the early-stage.

In the interviews we conducted, and on which the GSM is grounded, the theme of 
innovation only seldom emerged. However, if innovative and creative factors were 
considered in this model, the integration of results would have been 
straightforward, since it is compliant with the description of how the 
\textit{accumulated technical debt} (CAT5) causes the need of establishing a 
process in the long run (see how the \textit{initial growth hinders performance} 
(CAT6)). %MUN: not clear to me what you mean by "evolution 
%of this model" and what "compliant with our assumption" means. Also, don't use 
%only CATx in this context, but cite the category name AND CATx. CG: fixed
From the \textit{fluid} phase to the \textit{specific} phase, the innovation 
confounding factor seems to behave initially the same as the \textit{speed-up 
development}, both highly focused towards a product-oriented development. 
%MUN: <-- please provide some motivation... not obvious. CG: fixed
Nevertheless, we were unable to ground these evidences in the GT data even 
though those are factors that might be positively correlated 
%MUN: "present strict correlation" %What does that mean, I don't understand. 
%CG: fixed.
to the choice of an early product-development orientation and the need of 
low-precision engineering elements.

\subsubsection{Market requirements and application type} 
Market requirements \cite{Coleman2007} and application type \cite{Sutton2000, 
Coleman2008, Coleman2008a} are two further potential confounding factors to the 
GSM. Their main impact is related to the adoption of flexible and reactive 
solutions for the development process. In particular, they refer to the 
necessity of fulfillment of quality concerns that goes beyond scalability and UX 
defined by the GSM. Especially when requirements are rigidly imposed or the 
application domain is well-known, providing low-quality products to final users 
might determine the failure of a startup.

For example, safety-critical systems possess quality concerns that must be 
considered. Even in e-commerce systems factors related to quality have 
high-impact effects on obtaining the first customer reference, 
%MUN: <-- cannot parse it. Remove the first part and write: "Even e-commerce 
%systems...". Then rephrase the second part of the sentence to make it clearer. 
%CG: fixed
as discussed in \cite{Deakins2005,Kim2005,Silva2005}. Quality attributes  
especially %MUN: what is "this"? CG: fixed
characterize those applications, whose part of functionalities are already 
successfully implemented in other software systems or whose failure might 
inflict serious financial or social damages \cite{Bass2003}. Yet, limiting our 
context to early-stage software startups, which typically develop first 
cutting-edge products in beta versions, partially mitigates the high quality 
demands of users.

%MUN: The following 4 paragraphs explain how you use the Kano model to position 
%startups w.r.t. customer expectations and quality concerns. It's ok, but it 
%adds little value to the validation of GSM. I'd remove it.
%To understand how to obtain processes in accordance to a set of customer 
%needs, correlating the proper levels of customers' satisfaction, we 
%introduce the Kano model, described in \cite{Ullah2011} and extensively 
%studied in product development.  As shown in the Kano model, users 
%preferences can be classified into four categories: \textit{basic}, 
%\textit{excitement}, \textit{performance} % and \textit{indifference} areas.
% 
%The \textit{basic} area represents the space where customers are dissatisfied 
%if those \textit{must-be} functionalities are not met. This is the case of 
%similar functionalities already implemented in the current market with 
%well-known solutions. The \textit{performance} area represents the space where 
%customers' satisfaction is proportional to how many of functionalities are 
%implemented. In the \textit{indifference} area the customer does not really 
%care about the functionalities: the satisfaction remains invariant. This 
%represents the case where startups have not (yet) achieved the right 
%product/market fit. Finally, the \textit{excitement} area represents the space 
%where customers will be more satisfied if functionalities are implemented, but 
%not disappointed if they are not implemented. This is the case of extremely 
%innovative product development.
% 
%Well-known domains such as e-commerce applications or other \textit{``quality 
%critical products''} belong to the \textit{basic} area, where if a quality 
%concern such as security in bank transfer services is missing satisfaction of 
%customers will be profoundly affected. But as discussed, in early-stage 
%startups that typically develop first beta products, developers usually deal 
%with cutting-edge technologies. Those innovative solutions, if in the right 
%market, are likely to belong to  the \textit{excitement} area, where 
%satisfaction can hardly be negatively affected by faulty products. In fact, in 
%the GSM model, minimal functionalities and less concern about quality aspects 
%have low-impact on customers' satisfaction, supporting startup's ability to 
%mainly focus on \textit{speed-up development}. Moreover operating in the 
%\textit{excitement area} is fully compliant with an evolutionary approach and 
%minimal functionalities since from the first product released to the market, 
%it is unlikely to lose users because 
% of high-quality demand.
% 
%Nevertheless, this aspect profoundly undermines the generalizability of the 
%theoretical model according to the theory, in particular \textit{CAT3}. 
%therefore, future studies within wider startups' context are advisable to 
%extrapolate insights from those situations where quality concerns cannot be 
%postponed or externalized. CG: ok

\subsubsection{Developer experience} 
%MUN: Ups, a confounding factor you did not mention initially. Add it... 
%CG: fixed
Finally, also the experience of team-members is considered as an important 
confounding factor. Since only little evidence has been gathered during the 
interviews, ``lack of experience'' was not considered as a sub-category in 
\textit{severe lack of resources}. Nevertheless, startups rely at the beginning 
on clever, but inexperienced developers \cite{Crowne2002}. Having team-members 
with deep experience is a ``double-edge sword'' \cite{Yoffie1999}. On one hand, 
experience might quickly provide structure and maturity to the development 
process, on the other it might cause challenges in managing self-confident 
overachievers that almost inevitably clash. Consequently, team management 
requires higher control and coordination activities that inevitably hinder the 
flexibility of the development environment, that is essential in early-stage 
startups. Accordingly, also Coleman \cite{Coleman2007,
Coleman2008a,Coleman2008} states that the operating strategies towards 
\textit{``minimum processes''} in startups is not a matter of poor knowledge and 
training, but rather it is the necessity of operating with solutions that let 
the company move faster.

%MUN: removed below paragraph since it does not add much
%Nevertheless, since the amount of the results did not allow ``the lack of 
%experience'' to be in the GSM model (in view of the fact that not enough data 
%were identified \footnote{A category is formed if the codes are representative 
%of more than the 70\% of the analyzed transcripts.}), this confounding factor 
%requires further investigation. CG: ok

%------------------------------------------------------------------------------%
\subsection{Static assessment of the GSM}
\label{sect:theory:validation:stats}

%MUN: explain in the research method section (new section on how you performed 
%validation), exactly how you did static validation and what statistical 
%methods you used on what data, then refer to that section from here (no need 
%to repeat again here). CG: fixed
In this subsection we validate the most important causal relations presented in 
our model by attesting their correctness  with a quantification of empirical 
trends observed in the interview data. 
%MUN1: The following is part of the method and should be in 3.5: "We identify 
%trends and recurrent patterns by thoroughly analyzing and synthesizing the 
%complete set of data obtained from the case study on which the GSM is built 
%upon. We illustrate our observations aided by simple statistical methods, 
%providing researchers an additional tool, supported by further companies' 
%data, to validate our model."

We argued in section \ref{res:gsm:frmov}, that the most essential feature 
startups aim to achieve is high speed. As illustrated in Figure \ref{fig:gsm}, 
the core category (\textit{speed-up development}) is influenced by: the degree 
of quality concerns (contributes to the \textit{efficiency}); the undertaken 
development approach (provides \textit{effectiveness}); and finally the 
intrinsic characteristics of the team (guarantee \textit{performance}).

%MUN: for the paragraphs below, you need to explain what and why you do this. 
%CG: fixed, explained in methodology 
%What do you mean by "we grounded in the interview transcripts the fact that 
%those relations were characterized in terms of performances since 
%time-to-market was the main concern." Not clear to me. 
%CG: fixed, performance for us is efficiency AND effectiveness
Before starting the evaluation of the impact of the three categories on 
\textit{speed-up development}, we grounded in the interview transcripts the fact 
that those relations were characterized in terms of "effectiveness" and 
"efficiency" since time-to-market was the main concern (as confirmed in 
\cite{SMS}):

\begin{compactitem}
\item The team enhances the overall \textit{performance} of software 
development. Recalling that startups deal with severe lack of resources, the 
team members are the company's main assets. They develop software with no need 
of formalities and structures, left to the team's own capabilities to 
efficiently develop the product. Moreover, the effectiveness is granted by 
continuous advice given by mentors and by a management style oriented towards 
self-organization, full-stack, and multi-role settings. The importance of a 
skilled team for development performance is expressed in many studies such as 
\cite{surviving-os-cockburn, Highsmith2000}, described in CAT4 and confirmed by 
Sutton \cite{Sutton2000}.
\item The evolutionary approach enhances the \textit{effectiveness} in 
implementing the correct functionalities. This approach improves the company's 
capabilities to adjust the trajectory of product development (in agreement to 
the concept of pivoting discussed by Ries \cite{Ries2011}).
\item Since product quality has low priority, the development is focused only on 
implementing a limited number of suitable functionalities, enhancing its 
\textit{efficiency}. An essential prototype (MVP) does not require to comply 
with heavy quality constraints, enabling the team to quickly implement 
functionalities, ready to be validated by final users.
\end{compactitem}

An optimal combination of these three categories enables startups to ship code 
extremely quickly. However, the price that startups have to pay for achieving 
such high speed is related to the extent of the \textit{accumulated technical 
debt}, which will finally contribute to hinder performance in later stages.

Accordingly, by analyzing the data of the 13 companies, we formulated two 
hypotheses which further describe the above illustrated relations:
%MUN: You need to be more precise with terminology. Above you state that the 
%two items below are "trends which further empower relations", but then, 
%further below, you write "to assess these hypotheses". Observed trends != 
%Hypotheses. This is also why you need to be extremely explicit and precise 
%when explaining what the a) aim, b) method, and c) outcome/result of this 
%section is. I'm aware that it is difficult to boil down what you wrote in the 
%thesis to this limited space, but I think it is doable (and relevant for 
%supporting validity of GSM). CG: fixed

\begin{compactenum}
\item The more a capable \footnote{The team capability is evaluated according 
to the characteristics presented in CAT4.} team, with few quality constraints, 
uses an evolutionary approach, the faster the company releases the product.
\item The faster the company releases the product, the larger the degree of 
\textit{accumulated technical debt}\footnote{The debt here is considered 
regardless of context-specific debt mitigation factors and tactics.}.
\end{compactenum}

To assess these hypotheses we defined three measures reflecting the theoretical 
model:

\begin{compactitem}

\item \textit{Execution speed}: a metric that represents the development speed 
of the startup during different phases of the first release, computed by means 
of a weighted average speed for each phase. It was obtained by analysing 
interview transcripts looking at subcategories of \textit{Speed-up development}.

\item \textit{Technical debt}: a metric that represents the extent to which 
processes are controlled, structured, planned and documented by means of 
engineering artifacts and practices. It has been computed by means of a weighted 
average of the debt accumulated in each development phase, observing 
subcategories of  \textit{accumulated technical debt} with consequences on the 
startups growth.

\item \textit{Potential capability}: a metric that represents the degree to 
which each company reflected the capability of reaction and flexibility to the 
dynamic environment during the development process, given by the three 
categories that mostly affect \textit{speed-up development}.
\end{compactitem}

We quantified these measures for each company involved in our empirical study 
by defining score metrics based on a set of rubrics and evaluating startups 
accordingly, through an analysis of interview transcripts, codes and 
questionnaire results. The complete statistical procedure and provided rubrics 
to quantify the measures are illustrated and discussed in Appendix.
%MUN: make sure it is clear that the PROCEDURE material is available in the 
%ONLINE appendix and refer to it. Still, you need to report the actual RESULTS 
%of you statistical analysis here... 
%CG: partially fixed. we always referred to appendix without references until 
%now.

%MUN1: I don't really understand why you add the second clause in H01 
%"AND with less quality constraints" --> just makes the hypothesis more 
%difficult to understand. 
We conducted statistical tests  using the analysis of variance to assess the 
existence of relations between the dimensions. First we defined two null 
hypotheses:

\begin{itemize}
\item[$H0_1$:] \textit{Startups do not release the product faster 
when a capable team adopts a more evolutionary approach AND with less quality 
constraints}
\item[$H0_2$:] \textit{The execution speed does not increase the amount 
of accumulated technical debt}. 
\end{itemize}

Then we tested $H0_1$ and $H0_2$  with an 
one-tailed test using Pearson's product moment correlation 
coefficient\footnote{To perform the Pearson's correlation we verified two 
assumptions: a) data is normally distributed; b) data is on interval scale.}, 
with positive association analysis, fixing the level of confidence to 95\%.
We reject both $H0_1$ and $H0_2$, and conclude that in our sample:
\begin{compactenum}
\item Higher values of \textit{Execution speed} are strongly associated 
(p-value: 0.002) with higher values for \textit{Technical debt}.

\item Higher values of \textit{Execution speed} are 
strongly associated (p-value: 0.005) with higher values for  
\textit{Potential capability}.
\end{compactenum}

The validation has been performed on the most critical  causal relationships of 
the model. According to the paradigm model those are the causal condition which 
are linked to the core-category (see Figure \ref{fig:gsm}). There are four 
high-level relationships  which have not been validated by quantitative 
methods. However their correctness has been verified by a cross-analysis with 
existing literature and by attesting the causal relations among codes and 
categories, through axial and selective coding processes.

%------------------------------------------------------------------------------%
\section{Discussion} %MUN:Implications of the GSM
\label{sect:theory:impl}

%MUN: There are two issues with this section: it is too long and it basically 
%reiterates (in a concise manner) what you have said in the preceding 12 pages. 
%The real value of this section (suggested by my proposed name above), would be 
%if you can state in 0.5 (MAX 1 page), what are the implications of your 
%results on THEORY (or state-of-art, research) and PRACTICE (state-of-practice, 
%people who want to embark a startup). I'd propose you have a short intro here, 
%then have two subsections: State-of-art, State-of-practice where you state the 
%implications of GSM (or use a better subdivision if you come up with one, or 
%use none if you don't exceed 0.5 pages). These implications need to logically 
%follow from what you have said before (and the data you collected in the 
%interviews and, when necessary, by reference to the SMS). Don't have a long 
%bullet list, but structure your arguments in a logical flow. This is the 
%section in which you have to show, based on your previous rigorous and well 
%done analysis, the essence / gained knowledge, indicating what are the possible 
%next steps for research in this area. You can be creative/visionary here as 
%long as your arguments follow from the previous analysis.
%CG: not fixed, I have shortened the content of one page. But still we cannot 
%create the subsections you proposed since the implications are all derived 
%from the model. Then, we can still shorten it if you want. Let's discuss it 
%during the next skype call.

%MUN1: Ok, the subdivision is better than the plain bullet list and I'm fine 
%with the subsection names. Still, I would rename the section: "Discussion" --> 
%"Implications of the GSM"

In this section we present relevant implications which 
emerge from the behavior of early-stage startups, formally expressed in the GSM, 
answering to our research question (\textit{What is the current 
state-of-practice related to software development strategy in early-stage 
startups?}) through its sub-questions. Although the startups we inquired in the 
case study were spread across different nations and market sectors, we have 
found several patterns that emerged very clearly.

%------------------------------------------------------------------------------%
\subsection{Shortening time-to-market}

The most urgent priority of software development in startups is to shorten 
time-to-market. When developing a new product, which has not been attempted 
before, it is crucial for the company to release the product soon and observe 
the users reactions. The company's chances to survive, are strictly related to 
the ability to find the right product/market fit as quickly as possible. 
Developing software without releasing it for a long period of time, coming up 
with very complex solutions often is not a viable option. Business demands, 
investors interests, market pressures and the general uncertain conditions force 
startups to quickly iterate on the product by exploring and experimenting new 
functionalities progressively rolled-out to early-adopters.
%MUN1: Ok, I'm now "inventing" possible implications in the sections below... 
%this is your job and you are much more qualified to do that than me :-)

%MUN1:Do we (research) know enough about how to shorten time-to-market without 
%accumulating technical debt? What are possible venues for research?

%------------------------------------------------------------------------------%
\subsection{Light-weight methodology}
Startups do not apply any standard development methodology. When focused 
on building the first version of the product, startups do not observe any 
specific and standard development methodologies or processes.  From an accurate 
and in-depth investigation into daily activities, it emerged that the number of  
the observed development \textit{practices} is extremely small and never 
systematic (only little \textit{pair programming} sessions and rarely 
\textit{test driven development}). Despite the wide number of publications, 
which discuss the adoption of lightweight methodologies, we have found that 
especially in this first stage not only methodologies but even individual 
practices are basically neglected.

The closest development approach undertaken by early-stage startups is 
the \textit{Lean startup methodology}.  A highly evolutionary development 
approach, centered around the quick production of a functioning prototype and 
guided by customer feedback is a central aspect of the \textit{Lean startup 
methodology} pioneered by Ries \cite{Ries2011}. However, startups do not 
formally follow the cycle \textit{build-measure-learn} as it is proposed by the 
methodology, where they should set-up an experiment to \textit{test the riskiest 
element of the business idea}, nor make explicit use of the suggested good 
practices.
%MUN1: Why don't follow startups the build-measure-learn cycle proposed by 
%Ries? Well, naively I'd say that since he published in 2011, his ideas do not 
%have yet transferred to industry... Any other reason you can think of? Is it 
%still too heavyweight and in conflict with the highest priority of shorting 
%time-to-market? Is Lean still to Heavy for startups (in their early phases)?
For example, when it comes to user feedback, startups do 
not explicitly follow the step-by-step process of \textit{customer development}
formally defined by Blank \cite{Blank2005}. Instead, startups absorb and 
implement the high-level principles that outlines the methodology%MUN1:<-- 
%which methodology?
, reflected in the GSM by the theoretical category \textit{find the 
product/market fit quickly}.

%MUN1: paragraph below is candidate for removal
%The greatest part of engineering activities of startups are focused on 
%the implementation while only little attention is given to more conventional 
%activities (project management, requirement specifications, analysis, 
%architecture design, automatic testing, \ldots). Especially when bringing the 
%first product to the market, the attitude of startups towards traditional 
%engineering activities is strongly characterized by a general feeling of 
%unnecessary. Every document, artifact or process, which does not produce a 
%short-term tangible benefit to the software production is perceived as a waste 
%of time.

%------------------------------------------------------------------------------%
\subsection{Focus on minimal set of functionalities}
The first release of the product includes only a limited set of well 
suitable functionalities focused on user experience. The beta version is usually 
focused on a minimal and essential set of functionalities providing a reasonable 
user experience (UX). 
%MUN1: This rises the question, how do startups prioritize requirements. 
%There's abundant amount of literature on how to select/prioritize reqs., but 
%do they work in startups? Are they also too heavyweight or could they be 
%adopted by startups in a modified form in order to increase the probabilities 
%of success in the market?
As confirmed by practitioners, the user must be able to 
accomplish a task on the product without major interruptions in the flow, which 
would cause frustration before being able to capture his attention. To realize 
UX in a short time frame, developers focus on a limited number of suitable 
functionalities, especially in the first version(s) of the product. Other 
product qualities are regarded as less important: for instance, obtaining a 
high-reliability is not a priority since early adopters of innovative product 
are keen to accept minor bugs in exchange of new functionalities\footnote{These 
results further highlight differences between software development in startups 
and traditional web engineering, where quality aspects are rated as very 
important, even more than time-to-market \cite{Offutt2002}.}. From a business 
perspective, one might argue that established companies might take advantage of 
the new idea and develop a better version of the product internally.
Despite this might represent a risk, established companies usually cannot 
afford to operate in a market of \textit{innovators} and \textit{early adopters} 
group of users - described by the Roger's bell curve in \cite{Rogers} - because 
of the little return of investment they would gain, as discussed by Christensen 
in \cite{Christensen2003}. Instead, established companies have more interests 
in acquiring a startup and all the knowledge it has gained during the 
\textit{customer development} process while focusing effort on acquiring larger 
shares of a ``mass market'' sector. 
%MUN: candidate for shorting
%Furthermore, thanks to cloud computing and 
%more in general high availability of third party components, aspects such as 
%performance, security and scalability can be achieved with a reasonable 
%effort. 
%Thus, given the general lack of quality concerns, and supported by the fact 
%that the optimal strategy to improve the UX is by having the product used by 
%real customers, startups can effectively focus on achieving short 
%time-to-market.

%------------------------------------------------------------------------------%
\subsection{Empowering the team members}
The founding team is the most important determinant of speed. More than 
techniques and methodologies, the features of the team are the essential factors 
to achieve high-development speed. Discussed in the theoretical category the 
\textit{team is the catalyst of development}, this result has been confirmed in 
multiple occasions both in the interviews and in the questionnaires. What is 
truly important for the startup, among other factors, is the extremely 
high-colocation which enables continuous informal and colloquial coordination 
and discussions. 
%$MUN1: candidate for removal
%The initial (small) development team usually spends most of the 
%hours of the week working closely to implement the product, often extending 
%the conventional office hours.
The large amount of communication, spontaneously held in informal places, has 
been proven to be an extremely effective element of 
several successful software projects \cite{Highsmith2000}. Hence, the vastly 
\textit{informal} development environment of startups acts as a catalyst of 
speed.
%MUN1: candidate for removal
%Other important elements, which contribute in achieving such an initial 
%high-speed, are: minimal set of functionalities with low concerns on product 
%quality; evolutionary prototyping approach; smart use of advanced tools and 
%frameworks; low precision artifacts; and the close colocation and capability 
%of the team.

%------------------------------------------------------------------------------%
\subsection{Paying back the technical debt}
The initial lack of structures and processes negatively affects the 
performance when the company starts to grow. One of the primary objectives of 
startups is to expand their business into fast growing markets. If 
successful, the company will face a growth in terms of number of customers, 
employees and product capabilities. Then, the increased complexity arises the 
necessity of controlling the initially chaotic software development environment. 
Under delicate circumstances, where an increased market demand meets a poorly 
engineered product and process, the development team need to start returning 
the accumulated technical debt. As reported by practitioners, introducing more 
structures, documentation and standard procedures, cause an initial drop-down in 
performance which severity depends on the speed of growth, the amount of 
technical debt and the capacity of the team to absorb it. 
%MUN1: what does literature say about technical debt? Are there strategies that 
%maybe applicable to startups and if not why? Is there more research needed? 
%Obviously, shortening time-to-market and preventing technical debt are two 
%conflicting goals, but is there a chance to find a sweet spot where 
%development speed can be balanced to achieve market success and still reducing 
%the performance bump that hampers growth?

\subsection{Summary}
All of the above observations are extracted from the analysis of the model, 
which is systematically derived from empirical data of the case study. Indeed, 
with slightly different levels of adherence, the implications we presented are 
reflected in the behavior of the great majority of companies in the sample. The 
results of this analysis, which has been focused on the software development 
strategy from the idea conception to the first release, indicate that startups 
are far from adopting standard or ad-hoc development methodologies, especially 
in this phase. The typical tendency is to focus on the team capability to 
implement and quickly iterate on a prototype, which is released very fast. Thus, 
in a context where even for the most \textit{lightweight} agile methodologies is 
hard to penetrate, we believe that it is premature to discuss issues related to 
SPI. Researchers should rather focus on the trade-off between development speed 
and accumulated technical debt \cite{Brown:2010:MTD:1882362.1882373}, which 
appears to be the most important determinant for the future of 
software development in the company.

%------------------------------------------------------------------------------%
%\subsection{Future work}
%label{futwork}
%MUN: Don't have future work as a subsection in threats to validity. Either 
%merge it in "Implications of the GSM" OR into "Conclusions". CG: fixed

%MUN1: The paragraph below does say little about future work but rather 
%presents the contributions of GSM. In "Future work" you should rather state 
%what you consider needed to work on in relation with the research you have 
%done. Granted, replicating your case study and validating/extending GSM is 
%future work too, but it is not what *you* would do. Hence I'd suggest to 
%remove this section. Check that you state the contributions of GSM in the 
%conclusions.

%The GSM model enables practitioners in startup companies to speed-up 
%development empowering the role of the team, focusing on minimal 
%functionalities and adopting evolutionary development approaches. Researchers 
%can use the GSM model to reproduce the study with other software startups, and 
%focus on mitigation strategies to minimize the impact of the accumulated 
%technical debt over the need of shortening time-to-market. Moreover, the  
%mutual collaborative dialog between the research in the development strategies 
%in software startups and technical debt can benefits in understanding how both 
%influence the growth of a company after the release of the first product to the 
%market.

%------------------------------------------------------------------------------%
\section{Threats to Validity}
\label{valt}
%MUN: Also here, introduce some structure: first, explain shortly what types of 
%validity threats exist for the research methodology you applied. Then, discuss 
%each type individually. Think about which threat type is the most prominent 
%one for your research methodology and start with that one. I did not review 
%this section in detail now since it will change considerably after your 
%restructuring.
%CG: fixed

%MUN1: After reading the discussion on validity threats, I think you have
%chosen the wrong framework (Wohlin) to discuss threats. His framework is 
%targeted at experiments and even though it has been applied in other contexts, 
%other taxonomies are a better fit (see Yin and also Runeson). You tried to 
%squeeze your threats and strategies into Wohlin's taxonomy which seems in some 
%cases quite artificial and forced. I think it is better if you adopt Yin's 
%taxonomy of validity threats. If you read Yin (p 33-38) you'll see that his 
%taxonomy is a much better fit and that you can easier place and motivate your 
%mitigation strategies.
%------------------------------------------------------------------------------%
In this section we discuss the validity of the overall results. %MUN1: threats 
%are discussed w.r.t. the research methodology, not the results... 
We structure the discussion according to Wohlin's taxonomy \cite{Wohlin2000}.

%MUN1: Ok, what I meant was, not to explain the 4 categories in *general*, but 
%to explain (shortly) their relevance w.r.t. GT. I guess my comment above was 
%misleading...
%\begin{compactitem}

%\item External validity - which requires establishing the domain to which the 
%study's findings can be generalized (partially discussed above).

%\item Internal validity - which is enhanced by establishing causal 
%relationships whereby we attest that certain conditions lead to other 
%conditions, as distinguished from spurious relationships\footnote{A 
%\textit{spurious} relationship is a relationship in which two events have no 
%direct causal connection, but rather due to either a coincidence or the 
%presence of confounding factors.}.
%\item Construct validity - which is enforced by clearly specifing operational 
%procedures, and providing an adequate description of 
%constructs\footnote{Constructs are complex ideas that we as humans form in 
%order to summarize observations about things that we cannot see directly. For 
%example a construct is \textit{``job satisfaction''}, which to be defined, 
%requires a higher level of abstraction than a concrete object(e.g. a book).} 
%that are used to generate the theory.
%\item Conclusion validity - which represent the degree to which conclusions 
%are reasonable, according to the defined relationships in the data. It 
%involves ensuring adequate sampling procedures, appropriate analytical methods 
%and reliable measurement procedures. Reliability requires to demonstrate that 
%the operations of a study - such as data collection procedures - can be 
%repeated by other researchers, obtaining the same results.
%\end{compactitem}

%------------------------------------------------------------------------------%
\subsection{External validity}
%External
One threat to external validity is the selection of subjects interviewed 
for the study. This threat profoundly affects GT, since it is a qualitative 
research method, which makes the use of semi-structured interviews, and centers 
the research on respondent's opinions. 
In order to mitigate this threat we selected interviewees which 
covered positions of CTOs' and CEOs'. Their perspectives on what is taking 
place within the organization, were the only meaningful data taken in 
consideration in the study.  In many cases there was no supporting evidence to 
verify the opinion expressed. The questionnaire dedicated to ``data 
triangulation'' was not of primary support because of the limited time 
practitioners were able to grant to it. However, researchers have to accept the 
veracity of what respondents reported during the interviews \cite{Coleman2008a}. 
%MUN1: candidate for removal
%Notwithstanding, the issues surrounding semi-structured interviews are vital. 
%Within our study, although the reality might be different from what has been 
%described, the reported data are CTOs' and CEOs' perception, which is used to 
%base their decisions.

The comparison of the GSM with similar frameworks helped in establishing the 
domain to which the study's findings can be generalized. In particular the 
previous framework developed by Coleman \cite{Coleman2008} has allowed 
us to find similarities, differences and broader reasoning related to 
factors that hinder maturing processes in startups. Moreover, we analyzed 
Brook's framework describing the \textit{basic attacks} to the challenges of 
developing software systems and other literature covered by the SMS on startups.
\cite{SMS}\footnote{Note that the generalizability of results is applied only to 
those startups that operate in software-intensive activities.}. 
%MUN1: candidate for removal: detail on RM / seems out of place
%In conclusion, regarding the validation of 
%the subcategories, a continuous re-evaluation of the interview transcripts has 
%been performed.

%------------------------------------------------------------------------------%
\subsection{Internal validity}
%Internal
To enhance the internal validity, we created a three-dimensional research 
framework. By means of a grounded theory approach, formed by a 
\textit{systematic mapping study}, \textit{interviews} and \textit{follow-up 
questionnaires}, we searched for convergence among different sources of 
information to confirm the generated theory.

To validate the GSM we conducted a comparison of the emergent theory with 
existing literature and previously developed models. Furthermore we 
quantitatively evaluated three important relationships in the model. With the 
theory validation we highlighted and examined similarities, contrasts and 
explanations as discussed in \cite{Eisenhardt2007}. In this regard Eisenhardt 
stated: ``Tying the emergent theory to existing literature enhances the internal 
validity, generalizability, and theoretical level of the theory building from 
a case study research [\ldots] because the findings often rest on a very limited 
number of cases.''

%MUN1: candidate for shortening
%Little ambiguities about direction of causal influences have been reduced by 
%introducing arrows to indicate the direction of the causal relation in the GSM 
%model. Furthermore, by means of comparison between the GSM model and extant 
%studies in the broader fields, we revealed that the theory resembles in wider 
%startup development context. 
Nevertheless, we identified important confounding factors, related to 
innovation, market requirements and developer experience, which might 
impact the internal validity of the GSM.

%------------------------------------------------------------------------------%
\subsection{Construct validity}
%Construct
One threat to this study is a possible inadequate description of constructs. In 
order to diminish this risk, the entire study constructs have been adapted to 
the terminology utilized by practitioners and defined at an adequate level for 
each theoretical conceptualization. For instance, we defined \textit{Time 
shortage} in terms of \textit{Investor pressure, CEO/business 
pressure, Demo presentations at events and internal final deadline} as used by 
most of the interviewees in the study. 
Moreover, during the coding of interview transcripts, we adopted explanatory 
descriptive labels for theoretical categories, to capture the underlying 
phenomenon without losing relevant details.

%MUN1: I get the threat, but I don't understand its mitigation: let the 
%interviewees drive the study?! Try to reformulate...
The second important threat is caused by the fact that interviewees might 
already be aware of the possible emergent theories analyzed by researchers. To 
reduce the risk that some hypothesis could have been involved in the design of 
the research, we let mainly participants drive the direction of the case study.

Another threat associated with construct validity is evaluation 
apprehension, which describes people's rejection attitude towards the 
assessment of their behavior. 
To address this risk we didn't attained to first answers from practitioners, but 
rather went towards details of engineering elements and activities. %MUN1: <-- 
%I don't understand this...
Moreover, we developed a rigorous data collection protocol to create case study 
databases and employ integration of multiple data collection methods.%MUN1: <-- 
%Since you write "Moreover, ..." the reader gets the impression that this is a 
%further mitigation strategy to evaluation apprehension. I don't see however 
%how this mitigates this risk. Fix...
Reporting bias was mitigated by packaging all the needed 
material for conducting new researches in other contexts in order to criticize 
our findings (interview package with instructions has been made available 
online\footnote{Available at 
\url{https://github.com/adv0r/BTH-Interview-Package}.}. %MUN1: --> What exactly 
%is reporting bias? Let's assume that it means that you only report results, 
%i.e. interviewee statements, that are favorable to your theory. How does 
%publishing the auxiliary study material help to mitigate this risk? Publishing 
%the raw data (what you not do) would only allow researchers to check the 
%conclusions you made. The published auxiliary material helps to replicate the 
%study, i.e. makes is possible to test your theory in a different context. 
Moreover, two researchers not involved in the execution of the study conducted 
a peer-review analysis of the model's constructs.

To control distortion during analysis we made extensive use of memos. To 
demonstrate that we were aware of personal biases, memos provided us another 
important function in controlling the quality of data analysis. Through the use 
of memos and comparative analysis we were able to check if data fitted into the 
emerging theory and also countering subjectivity that ultimately enhanced the 
likelihood of producing accurate research findings.

%------------------------------------------------------------------------------%
\subsection{Conclusion validity}

Grounded theory has been already applied by other researches in similar 
contexts to attest relationships among conceptualizations of an examined 
phenomenon (see \cite{Coleman2007, Basri, Coleman2008a}). Those relationships 
between sampling and analysis should be verified in such a way that emerging 
findings remain consistent as further data are collected. To attest this, we  
explored and tested:

\begin{compactitem}
\item Each category and the strength of relations between them.
\item Hypotheses, derived from and related to the emergent theory.
\item Deviant cases to ensure robustness and general applicability.
\end{compactitem}

In particular we were prepared to modify generated categories so that the new 
data could be adapted into the emerging theory according to the concepts of 
theoretical sampling and theoretical saturation.

According to the theoretical sampling concept, we adjusted our 
research design and the emergent theory until only marginal results were 
generated. Moreover to enhance reliability of the outcome conceptualizations and 
relations, the process of coding interviews has been systematically conducted 
following a detailed process. Then, we asked to other practitioners a 
confirmation to further validate the generated framework and enhance the 
reliability of the relations among the identified categories.

An important issue is related to the fact that the limited number of interviews 
might not represent the complete scenarios in our context of study. 
Nevertheless, this issue has been partially mitigated as result of the 
theoretical saturation concept. This experience is attested by Martin's 
statement: \textit{``By the time three or four sets of data have been analysed, 
the majority of useful concepts will have been discovered.''} 
\cite{giardinoEtAl}. To obtain major certainty of the obtained results, what we 
applied was the final validation.

Despite we couldn't apply any power analysis procedures, in grounded theory the 
sample size depended on when data saturation occurred. Ramer in \cite{Ramer}, 
comparing quantitative to qualitative researches, states: \textit{``reaching 
data saturation, which involves obtaining data until no new information emerges, 
is critical for obtaining applicability in qualitative research''}. To avoid 
interrupting the research prematurely, after attesting that no more relevant 
information could be gathered from executing additional interviews, we iterated 
grounded theory study one more time, verifying that the explanatory power of the 
core category was fulfilled.

The  approach of generating theory allowed us to check emerging categories 
and their properties by gathering new evidence within an iterative and 
evolutionary approach, in view of the wide context of the research area that was 
under investigation.

%------------------------------------------------------------------------------%
%\subsection{Subsection Heading Here}
%Subsection text here.

% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

%\subsubsection{Subsubsection Heading Here}
%Subsubsection text here.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%

% Note that IEEE CS typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.
% However, the Computer Society has been known to put floats at the bottom.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE CS papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals use top floats exclusively.
% However, Computer Society journals sometimes do use bottom floats - bear
% this in mind when choosing appropriate optional arguments for the
% figure/table environments.
% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
% floats. This can be corrected via the \fnbelowfloat command of the
% stfloats package.

\section{Conclusion}
\label{conc}
Startups are able to produce cutting-edge software products with a wide impact 
on the market, significantly contributing to the global economy. Software 
development, especially in the early-stage, is at the core of the company's 
daily activities. Despite their severely high failure-rate, the SMS \cite{SMS} 
found that the quick proliferation of startups is not supported by an adequate 
and scientific body of knowledge. %MUN: <-- last sentence: isn't that a 
%conclusion from the SMS? CG: fixed
To be able to intervene on the software development strategy with scientific and 
engineering approaches, the first step is to understand startups' behavior. 
Hence, in this research work, we provided an initial explanation of the 
underlying phenomenon by means of a grounded theory case study, focusing on 
early engineering activities, from the idea conception to the first open beta 
release of the product.

Through an exploratory research, conducted by a systematic approach, we 
produced a theoretical model grounded into the hindsight knowledge collected 
among practitioners with the aim of explaining how development strategies are 
engineered startups. The explanatory capability and correctness of the model 
have been validated by means of systematic comparisons with the state-of-the-art 
results and empirical data. The systematic mapping study \cite{SMS}, 
with 37 studies extracted, revealed a 
multi-faceted state-of-the-art inadequate to support software development 
activities in startup companies. %MUN: <-- you mix again SMS and GT here... fix 
%it CG: not fixed, I would keep it since important part of the initial GT 
%methodology
%MUN1: Ok, I understand that. But this paper is about the GT, ONLY the GT. The 
%SMS is published elsewhere. Maybe it is ok how you use the SMS now... let's 
%see what Tony says...
On the other hand, the case study conducted in 13 early-stage startups, 
provided a broad set of empirical evidences obtained by combining different 
research methodologies in a grounded theory approach.

The overall results of our research confirm that startups possess %MUN: 
%why unique? CG: fixed.
characteristics of uncertainty, lack of resources and time-pressure. These 
factors influence the software development to an extent that transforms every 
decision related to the development strategies into a difficult trade-off for 
the company. Moreover, although startups share different characteristics with 
other similar SE domains (e.g. market-driven development, small companies and 
web engineering), the unique combination of coexisting factors poses a whole new 
set of challenges which need to be addressed by primary studies. Especially when 
bringing the first product to market, startups' most urgent priority is 
releasing the product as quickly as possible to verify the product/market fit 
and to adjust the business and product trajectory according to early feedback 
and collected metrics. In this stage, startups often discard any formal project 
management, documentation, analysis, planning, testing and other traditional 
process activities. Practitioners take advantage of an evolutionary prototyping 
approach, using well-integrated tools and externalizing complexity to third 
party solutions.

However, the initial gain obtained in terms of flexibility and speed is 
counterbalanced by the need of restructuring the product and controlling the 
engineering activities when the company starts to grow. If successful, the 
startup will face a growth in terms of number of customers, employees and 
product functionalities that leads to the necessity to control the initially 
chaotic software development environment. In this context the most significant 
challenge for early-stage startups is finding a sweet spot between being fast 
enough to enter the market early while controlling the amount of accumulated 
technical debt.  %MUN: Not sure whether you have discussed/mentioned the meaning 
%of chaotic/complex domain in the paper, I think not. Hence you need to 
%rephrase that (but you can not introduce new concepts here in the conclusion). %CG: fixed. 
Therefore, from a software development perspective, working with low-precision 
engineering activities represents the only viable strategy that startups can 
follow in the early stage of their lifecycle.

In this research we discussed a number of novel challenges for both 
practitioners and researchers, while presenting a first set of concepts, terms 
and activities which set the software engineering scene for the rapidly 
increasing startup phenomenon. By making a comparison with Berry's definition of 
SE \cite{Berry1992}, we would like to see the rise of a new discipline - 
\textit{startup engineering} - which can be defined as  \textit{the use of 
scientific, engineering, managerial and systematic approaches with the aim of 
successfully developing software systems in startup companies}. %MUN: Nice ;-)

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

% use section* for acknowledgement
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi

The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{tse}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{John Doe}
Biography text here.
\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here.
\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}