### 2013-03-04 article/main.tex
Compiles again...

%% bare_jrnl_compsoc.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE Computer
%% Society journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%% bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do ***
% *** not appear when using other class files. ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
% The Computer Society usually requires 10pt for submissions.
%
\documentclass[10pt,journal,letterpaper,compsoc]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[12pt,journal,compsoc]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
% % pdf code
% \else
% % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

%%extra packages ------
\usepackage{longtable}
\usepackage{float} %allows the use of [H] to force positioning of figures
\usepackage{subfigure}
\usepackage{hyperref} %used in references now --- it shoud be removed
\usepackage{amssymb}
\usepackage[neveradjust]{paralist} %compact lists
\usepackage{amsthm}

\newtheorem*{theory}{Theory}

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  % \usepackage[nocompress]{cite}
\else
  % normal IEEE
  % \usepackage{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later. Note also the use of a CLASSOPTION conditional provided by
% IEEEtran.cls V1.7 and later.

% *** GRAPHICS RELATED PACKAGES ***
%
%\ifCLASSINFOpdf
   \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
   \graphicspath{{figures}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
%\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
% \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
   %\graphicspath{{./figures/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
   %\DeclareGraphicsExtensions{.eps}
%\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%\usepackage[tight,normalsize,sf,SF]{subfigure}
%\else
%\usepackage[tight,footnotesize]{subfigure}
%\fi
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. Computer Society papers
% use a larger font and \sffamily font for their captions, hence the
% additional options needed under compsoc mode. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\ifCLASSOPTIONcompsoc
% \usepackage[caption=false]{caption}
% \usepackage[font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
% \usepackage[caption=false]{caption}
% \usepackage[font=footnotesize]{subfig}
%\fi
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\ifCLASSOPTIONcompsoc
% \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
% \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

%\ifCLASSOPTIONcaptionsoff
% \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
% This package may be useful when used in conjunction with IEEEtran.cls'
% captionsoff option. Some IEEE journals/societies require that submissions
% have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.3.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% For subfigure.sty:
% \let\MYorigsubfigure\subfigure
% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat/subfig command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex). ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Software Development in Startup Companies: The Green-field Startup Model}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc
% journal papers have the author affiliations above the "Manuscript
% received ..." text while in non-compsoc journals this is reversed. Sigh.

\author{FirstName~SecondName,~\IEEEmembership{Fellow,~IEEE,}
        ..~..,~\IEEEmembership{Fellow,~IEEE,}
        and~...~....,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem Author info.\protect\\
% note need leading \protect in front of \\ to get a newline within \thanks as
% \\ is fragile and will error, could use \hfil\break instead.
E-mail: see http://something/contact.html
\IEEEcompsocthanksitem Author info.}% <-this % stops a space
\thanks{}}

% note the % following the last \IEEEmembership and also \thanks -
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
%
% \author{....lastname \thanks{...} \thanks{...} }
% ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer
Society Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
%
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers. ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2007
% IEEE}%
%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer
% Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society jorunal
% papers don't need this extra clearance.)

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
%\boldmath
Software startups are newly created companies with no operating history and
oriented in producing cutting-edge technologies. These companies develop
software under highly uncertain conditions, tackling fast growing markets with
severe lack of resources. Despite their increasing importance in economy there
are only few scientific studies attempting to address software engineering (SE)
issues, especially for early-stage startups. This research aims to understand
the software development strategies conducted by practitioners, in the period of
time that goes from idea conception to the first open beta release. A
\textit{grounded theory} approach guided the execution of a case study to
explore the state-of-practice. As result, the Greenfield startup model (GSM)
reveals the urgent priority of startups of releasing the product as quickly as
possible to verify the product/market fit and to adjust the business and product
trajectory according to early collected user feedback. Nevertheless, the need of
shortening time-to-market, by speeding-up the development through low-precision
and product-centric engineering activities, is counterbalanced by the need of
restructuring the product before setting off for further growth.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway. In particular, the Computer Society does
% not want either math or citations to appear in the abstract.

% Note that keywords are not normally used for peer review papers.
\begin{keywords}
Software Development, Startups, Grounded Theory.
\end{keywords}}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEcompsoctitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynotcompsoctitleabstractindextext when compsoc mode
% is not selected <OR> if conference mode is selected - because compsoc
% conference papers position the abstract like regular (non-compsoc)
% papers do!
\IEEEdisplaynotcompsoctitleabstractindextext
% \IEEEdisplaynotcompsoctitleabstractindextext has no effect when using
% compsoc under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

%------------------------------------------------------------------------------%
\section{Introduction}
\label{intro}
% Computer Society journal papers do something a tad strange with the very
% first section heading (almost always called "Introduction"). They place it
% ABOVE the main text! IEEEtran.cls currently does not do this for you.
% However, You can achieve this effect by making LaTeX jump through some
% hoops via something like:
%
%\ifCLASSOPTIONcompsoc
% \noindent\raisebox{2\baselineskip}[0pt][0pt]%
% {\parbox{\columnwidth}{\section{Introduction}\label{sec:introduction}%
% \global\everypar=\everypar}}%
% \vspace{-1\baselineskip}\vspace{-\parskip}\par
%\else
% \section{Introduction}\label{sec:introduction}\par
%\fi
%
% Admittedly, this is a hack and may well be fragile, but seems to do the
% trick for me. Note the need to keep any \label that may be used right
% after \section in the above as the hack puts \section within a raised box.

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps (small caps for compsoc).
%
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
%
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
%
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
%
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{A}{n} impressive number of new software startups are launched
worldwide every day as a result of an increase of new markets, accessible
technologies, and venture capital \cite{8491286}. With the term
\textit{software startups} we refer to those temporary organizations focused on
the creation of high-tech and innovative products, with little or no operating
history, aiming to grow by aggressively scaling their business in highly
scalable markets.

New ventures such as \textit{Facebook}, \textit{Linkedin}, \textit{Spotify},
\textit{Pinterest}, \textit{Instagram}, \textit{Groupon} and \textit{Dropbox},
to name a few, are good examples of startups that evolved into successful
businesses. But, despite many successful stories, the great majority of startups
fail within two years from their creation, primarily due to self-destruction
rather than competition \cite{Crowne2002}. Operating in a chaotic, rapidly
evolving and uncertain domain, software startups face intense time-pressure from
the market and are exposed to tough competition \cite{Maccormack2001,
Eisenhardt1998}. In order to succeed in this environment, it is crucial to
choose the right features to build and be able to quickly adapt the product to
new requests constrained by very limited amount of resources \cite{Sutton2000}.

From a software engineering perspective development in startups is challenging
since they develop software in a context where processes can hardly follow a
prescriptive methodology \cite{Sutton2000, Coleman2005}. Despite startups share
some characteristics with similar domains (e.g. small and web companies), the
combination of different factors makes the specific development context quite
unique \cite{Blank2005, Sutton2000}. Therefore, research is needed to support
their engineering activities \cite{Coleman2005}, guiding practitioners in taking
decisions and avoiding choices that could easily lead the whole business to
failure \cite{Kajko-Mattsson2008}. However, despite the impressive size of the
startup ecosystem  \cite{ISI:000243253000007}, the state-of-the-art presents a
gap. Through a  \textit{Systematic Mapping Study} (SMS) of the literature,
conducted by  Paternoster et al. \cite{SMS}, only few publications were found
related to  engineering activities in startups. Moreover the majority of these
studies do  not possess the relevance and rigour required to form a consistent
body of  knowledge in the state-of-the-art and in the state-of-practice.

This research aims to understand how software development strategies are
engineered by practitioners in startup companies in terms of level of:
structure, planning and control of software projects, in the period of time
that goes from idea conception to the first open beta release of the software
product.

With a cross-sectional study we focused the research context in a limited  time-
frame to emphasize the nature of uncertainty in the development activities  and
time pressure from the market, as discussed in \cite{Maccormack2001,
Eisenhardt1998}. We performed semi-structured in-depth interviews with CEOs and
CTOs of startups, covering a wide spectrum of thematics and iteratively
adjusting the direction of the research according to the emerging evidences.

By means of the Green-field startup model (GSM), we captured the underlying
phenomenon of software development in early-stage startups. The framework is
based on a systematic procedure  which is grounded into empirical concepts
obtained by a case-study. GSM presents the most significant themes of the
development strategies that characterise startups' context.

The results of our study suggest that the most urgent priority of software
development in startups is to shorten time-to-market. The greater part of
engineering activities are focused on the implementation while only little
attention is given to more conventional activities (project management,
requirement specifications, analysis, architecture design, automatic testing).
The first release of the product includes only a limited set of well suitable
functionalities.

Nevertheless, by speeding-up the development process, startups accumulate
technical debt, ignoring processes, relying on informal communication and
replacing documentation with low-precision artifacts. For startups it is
essential to look ahead within short-term deadlines and determine the right
product/market fit driven by user feedback, using flexible and reactive
development approaches and adopting mitigation strategies to contain the
severity of accumulated technical debt.

Practitioners can use the GSM to apply presented strategies to speed-up the
development by taking in consideration the amount of technical debt accumulated
during the first early-stage period. In this regard, we identified several
commonalities between the issues related to software development in startups and
the research focused at studying technical debt \cite{Nugroho2011,Izurieta2012}.
On the other hand, researchers can take advantage from the GSM to  understand
how the technical debt influences the future growth of startup companies.

The remainder of this paper is structured as follows:  we provide relevant
information about startups in section \ref{backg}. Section \ref{resmet}
introduces the research questions and discusses the design and execution of the
research process. Results are discussed in section \ref{res:gsm}, which presents
the GSM. Section \ref{res:val}  compares results of this research to  literature
and frameworks presented in this field. Section \ref{valt} discusses  the
validity threats considered during this study. Section  \ref{sect:theory:impl}
discusses the most relevant implications of the GSM.  Finally, this paper
concludes in section \ref{conc}.

%MUN: This is maybe only a minor concern, but it may be relevant for certain
%reviewers: The references you use in the introduction are not very recent.
%Look at the list of papers [1-15], most of them are published before 2005. I
%understand that you use some references to strengthen your general arguments
%(which can refer to older publications), but your introduction should also
%show the relevance of your research right now. This is best shown by  referring
%to recent (max 5 years old) literature. Think about if you can add  more recent
%(2010-now), well published, papers in the introduction.

% You must have at least 2 lines in the paragraph with the drop letter (should
% never be an issue)

%CG: Not fixed. There is a huge gap on the topic, the refs in the introduction
%are those retrieved by the SMS.

%\hfill mds

%\hfill January 11, 2007
%------------------------------------------------------------------------------%
\section{Background} \label{backg}

Modern entrepreneurship has been boosted by the advent of the consumer internet
markets in the middle of the 1990s and culminated with the notorious  \textit
{dot-com bubble burst} of 2000 \cite{Perkins:1999:IBI:555126,
storey1982entrepreneurship}. Several years later, with the massification of the
internet and mobile devices, we are now assisting to an impressive proliferation
of software ventures - metaphorically referred as the \textit{startup bubble}.
The easy access to vast potential markets and the low cost of services
distribution are appealing condition for modern entrepreneurs \cite{Marmer2011}.
Inspired by success stories, a large number of software businesses are created
every day. However, the great majority of these companies fail within two years
from their creation \cite{Crowne2002}.

Looking at the number of new business incubators which appeared in the last
decade one can estimate the importance of startups \cite{Grimaldi2005}. The wave
of disruption in new technologies has led companies to be more and more
competitive, forcing themselves to radical organizational and innovational
renewals,  which bring many companies to the attempt of behaving like startups
\cite{Christensen1997}.

The implementation of methodologies to structure and control the development
activities in startups is a major challenge for engineers \cite{Coleman2008}.
Generally, the management of software development is achieved through the
introduction of software processes, which define what steps the development
organizations should take at each stage of production and provide assistance in
making estimates, developing plans and measuring quality \cite{Chroust1996}. In
the last decades, several models have been introduced to control software
development activities. However, their application in startup companies doesn't
report significant benefits \cite{Coleman2008a, Coleman2008, Sutton2000}.

In this context, software engineering (SE) faces complex and multifaceted
obstacles in understanding how to manage development processes. Sutton defines
startups as creative and flexible in nature and reluctant to introduce process
or bureaucratic measures which may result ineffective \cite{Sutton2000}. Also
Bach refers to startups as \textit{``a bunch of energetic and committed people
without defined development processes''} \cite{Bach1998}. In fact, startups have
very limited resources and typically wish to use them to support product
development instead of establishing processes \cite{Coleman2008, Heitlager2007}.
Some attempts to taylor lightweight processes to startups reported basic failure
of their application: \textit{``Everyone is busy, and software engineering
practices are often one of the first places developers cut corners''}
\cite{Martin2007}. Rejecting the notion of repeatable and controlled processes,
startups prominently take advantage of unpredictable, reactive and low-precision
\cite{surviving-os-cockburn} engineering practices \cite{Sutton2000,
Tanabian2005, Chorev2006, Kakati2003}.

Moreover, most startups develop packaged applications rather than software for
a specific client \cite{genome2012}. Issues related to this domain are addressed
in literature by the area known as \textit{market-driven software development}
(sometimes called \textit{packaged software development} or \textit{COTS
software development} \cite{Alves2006}). Among other results, researchers
emphasize the importance of time-to-market as a key strategic objective
\cite{dagMDR,sawyer99} for companies operating in this domain.  Other peculiar
aspects which influence the software development are related to requirements,
which are reported to be often ``invented by the software company''
\cite{512553}, ``rarely documented'' \cite{Karlsson02challengesin}, and can be
validated only after the product is released to market
\cite{dahl2003,Keil:1995}. Under these circumstances, failure of product
launches are largely due to ``products not meeting customer needs''
\cite{Alves2006}.

Accordingly, product-oriented practices help startups having flexible teams,
with workflows that leave them the ability to quickly change the direction
according to the targeted market  \cite{Heitlager2007,Sutton2000}. In this
regard, many startups focus on team productivity, asserting more control to the
employees instead of providing them rigid guidelines \cite{Tanabian2005,
Chorev2006, Kakati2003}.

%------------------------------------------------------------------------------%
\subsection{General lack of research in startups} Despite some studies tried to
address the above mentioned issues, a SMS  \cite{SMS} revealed only a few SE
works in this specific area, as confirmed by  other studies \cite{Coleman2008,
Coleman2008a, Coleman2007, Sutton2000}.  Moreover, the studies, identified in
\cite{SMS}, appear to be highly fragmented  and spread across different areas
rather than constituting a consistent  \textit{body of knowledge}.

A research published in 2000 by Sutton \cite{Sutton2000} confirmed a general
lack of  studies in this area, claiming that \textit{``software startups
represent a segment that has been mostly neglected in process studies''} and it
has been further confirmed with the empirical studies of Coleman et al.
\cite{Coleman2008,Coleman2008a,Coleman2007} eight years later.

%------------------------------------------------------------------------------%
\subsection{Software development in startups}

The word \textit{startup} appeared in the SE literature for the first time in
1994 in an article written by Carmel \cite{Camel1994a} where he studied the
time-to-completion in a young package firm. %MUN: If this is a quote  
%from Carmel, put it in quotation marks, like you did in the previous section  
%with Suttons quote. Generally, check that you are consistent how you use  
%quotations throughout the paper.  
%CG: fixed, btw  it was just to emphasize his the focus of his study.  
%MUN1: Ok. As you may have noticed, I have removed a lot of cases were you use  
%italics. If you emphasize *too* much, it it rather distracting than helpful  
%for the reader. Furthermore, it conflicts with you convention you use to  
%highlight high level categories of GSM. CG: ok. 
He noticed how these companies were particularly innovative and successful,  
advocating for the need of more research investigating software development  
practices so as to replicate success and try to transfer it to other 
technology sectors.

First insights reveal how software startups are product-oriented in the first
period of their development phase \cite{Heitlager2007}. Despite good
achievements at the beginning, software development and organizational
management increase in complexity \cite{1456074,Banker1998} causing
deterioration of performance over time. Briefly, the necessity of establishing
initial repeatable and scalable processes cannot be postponed forever.

A study of Kajko-Mattsson \cite{Kajko-Mattsson2008}, which investigated a
Swedish software startup, reported a heavy lack of requirements gathering
processes, minimal project management, lack of control over the change requests,
absence of documentation to track the status and progress of the process, and
defective releases. Accordingly, Ambler et al. report how two startups
approaching to an upcoming IPO started to require processes to focus on scalable
solutions, in view of the growing company's size in terms of users and employees
\cite{Ambler2002}. In this regard, Crowne \cite{Crowne2002} specifies different
stages through which software startups evolve. Starting without any established
workflows, startups grow over time, creating and stabilizing processes to
eventually improving them only when sufficiently mature.

But when startups have no time for training and orienting activities, as
discussed in \cite{Sutton2000}, their main focus remains on team capabilities
instead of prescriptive processes, hiring people who can \textit{``hit the
ground running''} \cite{Yoffie1999}. Empowering the team and focus on
methodological attributes of the processes oriented in prototyping,  proof-of-
concepts, mock-ups and demos, to test basic functionalities, have been  the
primary priority in startups \cite{Camel1994a}. Only when they grow, formal
methodologies arise, followed by performing quality assurance and long-term
planning processes \cite{Yoffie1999}.

The maturity of a company affects the extent to which processes are adopted.
Tingling \cite{Tingling2007} reports how introducing Extreme Programming (XP)
principles \cite{Beck:2004:EPE:1076267} in the development process was
challenging because of the need of trained team-members for fully implementing
the methodology. Da Silva and Kon \cite{Silva2005} were able to start with all
the XP practices in place only after six months of coaching the team, trying to
enhance maturity from day-one. Nevertheless, even then, customization of
practices were inevitably implemented to adapt the processes to the undertaken
startups' context \cite{Deias}.

As partially discussed above, contributions to flexibility and reactiveness of
the development process has been conducted prominently by means of \textit{Lean}
\cite{Gautam2008} and \textit{Agile} \cite{Abrahamsson2002} methodologies (also
reported in \cite{Taipale2010, Kuvinka2011}), where the extreme uncertain
conditions lead startups to learn fast from trials and errors with a strong
customer relationship in order to avoid wasting time in building wrong
functionalities and prevent rapidly exhaustion of resources \cite{Midler2008,
Hilmola2003, Sutton2000}. Customer involvement in software development has also
been discussed by Yogendra \cite{Yogendra2002} as an important factor to
encourage an early alignment of business concerns to the technology strategies,
because both are salient considerations to be successful \cite{Tingling2007}.

When startups take a development approach that is mainly product-oriented
rather than process-oriented, it is essential to have a flexible team, with a
workflow that helps them quickly to change direction according to the target
market \cite{Sutton2000}. In this regard, many startups have been focused on
team productivity, providing more control to the employees instead of providing
them rigid guidelines \cite{Tanabian2005, Chorev2006, Kakati2003}.

%------------------------------------------------------------------------------%
\subsection{Software process improvement in startups}

When startups are mature enough to support software process improvement (SPI),
the solutions considered according to the state-of-the-art are oriented towards
light-weight processes such as a design of development process based on XP,
proposed by Zettel \cite{Zettel2001}. The process consists of a set of
activities and artifacts (in addition to some important roles) defined in order
to identify responsibilities and tools to utilize. But, despite the promising
benefits reported by Zettel, we were not able to identify any future evaluation
in real-world settings.

Another attempt of SPI in startups has been conducted by Deakins et al.
introducing a \textit{Helical Model for managing e-commerce development
environment} \cite{Deakins2005}. Also in this case, the author prescribed broad
guidelines for a rapid high-quality development process, which underwent limited
testing only in academic settings.

The problem of \textit{one-size-fits-all}, related to the SPI  representations
for startups, is described by Fayad \cite{Fayad1997}. The author discusses the
problem in actuating the same \textit{best-practices} criteria for established
companies in 10-person software startups. Thoroughly remarked in
\cite{Sutton2000}, Sutton states that problems of SPI in software startups arise
because of: the dynamism of the development process, which precludes
repeatability; organizational maturity, that cannot be maintained from startups
in view of lack of corporate direction; severe lack of resources, both human and
technological for process definition, implementation, management, and training.
In conclusion, the primary benefits of SPI do not address startups, which
instead of promoting product quality, aim to minimize time-to-market.

Additionally, the role of SPI has been neglected because seen as an obstacle to
the development team's creativity and flexibility, and to  the need of a quick
delivering product process environment \cite{Coleman2008a}. 
%MUN: the second part of the sentence is just a statement without reference. If 
%it refers to a statement by Coleman, put his reference at the end OR find the  
%correct reference OR delete the statement. %CG fixed 
Product quality is left aside in favor of minimal and suitable functionalities  
to shorten the time-to-market. As reported in two studies of Mater and Mirel  
\cite{Mater2000, Mirel2000}, quality aspects, mostly taken in consideration in  
internet startups, are oriented to usability and scalability, even though the  
market and application type heavily influences the quality-demand  
\cite{Coleman2008, Kim2005}.

To maintain the development activities, oriented to limited but suitable
functionality, studies suggest to externalize the complexity of parts of the
project to third party solutions by means of outsourcing activities
\cite{Hanna2010}, software reuse \cite{Jansen2008} and open-source strategies ,
\cite{Wall2001, Bean2005}.  %MUN: Can you split up the references? I.e. write:
%"outsourcing activities \cite{Bla}, software reuse \cite{Bla}, and open-source
%strategies \cite{Bla}. %CG: fixed

%------------------------------------------------------------------------------%
\subsection{Technical debt} Notwithstanding, a new area of SE research,
trying to tackle the problem of the \textit{technical debt}, brings and
encompasses different implications in studying development in software startups
\cite{Tom2013}. The metaphoric neologism was originally introduced by
Cunningham in 1992 \cite{TechnicalDebtCunn} and has recently attracted the
attention of SE researchers\footnote{To attest the fact that \textit{technical
debt} is gaining  traction among researchers, we mention two important
contributions which characterize the \textit{``debt landscape''}:
\cite{Nugroho2011,Izurieta2012} and a dedicated workshop \cite{workshopDebt}
organized by the Software  Engineering Institute  and ICSE.}. The concept of
\textit{technical debt} is  well illustrated in
\cite{Brown:2010:MTD:1882362.1882373}: \textit{``The idea  is that developers
sometimes accept compromises in a system in one dimension  (e.g., modularity) to
meet an urgent demand in some other dimension (e.g., a  deadline), and that such
compromises incur a ``debt'' on which ``interest'' has  to be paid and which the
``principal'' should be repaid at some point for the long-term health of the
project''}.  The compromise between high-speed and high-quality engineering is
faced daily by software startups, not only in terms  of architecture design but
in multifaceted aspects (weak project management, testing, process control).

Since \textit{``all decisions related to product development are trade-off
situations''} \cite{Hilmola2003}, generally startups optimize workflows to the
dynamic context they are involved into. They typically adopt any development
style that might work to support their first needs in what is called the
\textit{``Just do it''} school of software startups \cite{Ries2011}.
Additionally, as remarked by Coleman, \textit{``many managers just decide to
apply what they know, as their experience tells them it is merely common
sense''} \cite{Coleman2008}.

\subsection{Terminology} In order to set a common ground and to prevent
ambiguity, we use the following  terminology throughout the paper:

\begin{compactitem} 
\item Software development strategy: the overall approach adopted by the  
company to carry out the product development. 
\item Engineering activities: the activities needed to bring a product from  
idea to market. Traditional engineering activities are, among others,  
requirement engineering, design, architecture, implementation, testing. 
\item Engineering elements: any practice, tool and artifacts contributing and  
supporting the engineering activities.
\item Quality attributes: those overall factors that affect run-time
behavior, system design, and user experience. They represent areas of concern
that have  the potential for applications to impact across layers and tiers.
Some of these  attributes are related to the overall system design, while others
are specific  to run time, design time, or user centric issues
\cite{Microsoft2009}. 
\item Growth: an increase in the company size respect to the initial 
conditions in terms of either employees or users/customers, and
product complexity in terms  of handling an increasing number of feature
requests. 
\item Software product: any software product and software service.
\end{compactitem}

%------------------------------------------------------------------------------%
\section{Research methodology} \label{resmet}

\begin{figure*}[!t] \centering
\includegraphics[width=5in]{figures/completemethodology2} \caption{Grounded
theory study process overview} \label{fig:gt:completemethodology} \end{figure*}
%MUN: I think you can increase the font size in the figure a bit. Try thereby
%to keep the height of the figure, expanding however the width (use horizontal
%space since it is a two column figure). Change also in figure:  - Write "Data
%Analysis" horizontally  - Change names of macro-phases to upper-case, i.e.
%Design and Execution,     Data Collection, Data Analysis, etc.  - Add section
%number to macro-phase name, i.e. 3.1 Design and Execution, 3.2     Data
%Collection  - Change "Discussion and Validation" into "Theory Validation" as
%you name it     in the text. Also, add a section (3.5) that describes shortly
%what you did in "Theory  Validation"  %CG: fixed... fontsize cannot be
%increased uniformly among all the boxes, then  I skipped that part... 
%MUN1: OK, but in the figure, the correct section number for theory validation  
%is 3.5, not 3.4. CG:fixed

The aim of this research is to \textit{understand how software development
strategies are engineered by practitioners in startup companies in terms of
level of: structure, planning and control of software projects, in the period of
time  that goes from idea conception to the first open beta release of the
software product} \footnote{The above mentioned goal is structured according to
the GQM paradigm \cite{Basili1992}.}.

Initially the boundaries of the research domain were set by means of  non-
systematic literature survey, which provided the initial keywords used  further
in the research process. Moreover it helped us in defining the research
questions (RQs):

\begin{compactitem}

\item RQ-1 : What is the current state-of-practice related to software
development strategy in early-stage startups?

  \begin{compactitem}
  \item RQ-1.1 : How do startups structure and execute the main engineering
    activities? 
  \item RQ-1.2 : How are product quality attributes considered by
    startups? 
  \end{compactitem}

\end{compactitem}

We investigated the software development approach undertaken by practitioners
of startups. Following a \textit{Grounded Theory} (GT) methodology
\cite{Glaser1978}, we executed 13 semi-structured interviews integrated with
follow-up questionnaires. From this, we elaborated and extracted the Green-field
startup model (GSM) explaining the underlying phenomenon of software development
in startups.

\textit{Grounded Theory} methodology is described by Robson as \textit{``the
best approach to answer the question - What is going on here?''} and defined by
its creators \cite{ColinRobson2009} as \textit{``a set of well-developed
categories ( e.g. themes, concepts ) that are systematically interrelated
through statements of relations to form a theoretical framework that explains
some relevant social, psychological, educational, nursing or other
phenomenon''}.

Despite GT has its roots in the social sciences, it has been increasingly used
by Information System researchers and has been almost ignored by the SE research
\cite{Coleman2007}. In 1997 Bertelsen advocated for a need of more qualitative
research in SE \cite{Bertelsen1997}, and in the last decade we assisted to an
increasing number of publications that used a GT-like approach in SE
\cite{Coleman2008a,Coleman2007, Sulayman2012}.

Following the GT principles we attempted to capture the most relevant aspects
of software development from startup practitioners, letting a theory emerge
from the interviews and adjusting the research hypotheses and questions as we
proceeded through. During these interviews we collected data related to
engineering activities undertaken by startups in the time frame between the idea
conception and the first open beta release of the product.  Then, we proceeded
to the analysis of the data, finding important relations among concepts with a
formal approach.

As suggested by Coleman, in view of the different versions of GT which emerged
in the last years, researchers should indicate which \textit{``implementation''}
of the theory is being used \cite{Coleman2007}. After the methodology was
initially introduced, its original authors  had divergences regarding how the
theory should be formed by analyzing the data (Strauss and Corbin on one side,
Glaser on the other).  Glaser advocated for a more \textit{puristic} approach,
where the theoretical categories should \textit{``naturally emerge from the
data''}, whilst Strauss and Corbin formulated an updated version of the
methodology, which leaves to the researchers an higher degree of freedom. The
latter approach empowers researchers' \textit{``theoretical sensitivity''}
\cite{Corbin1990}, and encourages them to outline the research problem
beforehand. Since the knowledge obtained from parallel execution of the SMS
\cite{SMS} and our direct experience with startups companies provided a good
initial level of knowledge, in this study we use the Corbin and Strauss
approach \cite{Strauss1998}.

In this research we address technical aspects related to software development
in startups, exploring their operational dynamics. In view of the lack of
agreement on an unique definition of the word \textit{startup}, we delimited our
initial contextual boundaries to newly created innovative and single-product
software companies, in the time-frame that goes from the idea conception to the
release of the first product in highly scalable markets. Moreover, since most
startups are web-oriented companies \cite{tc-webstartups-link, Allen2003}, in
our case study, we inquired web companies with little or no operating history.

A complete overview of the overall case study methodology and execution is
shown in Figure \ref{fig:gt:completemethodology}, which presents how we tailored
the general GT methodology to our specific needs.

The initial design of the case study is supported by the results of the
\textit{Systematic Mapping Study} (SMS) \cite{SMS} which contributed to define
the initial questions for both questionnaires and semi-structured interviews.
The process depicted in Figure \ref{fig:gt:completemethodology} is evolutionary
and affects the design at each new iteration. In \textit{Data Collection} the
empirical results are integrated and stored in a single case study database and
subsequently processed in \textit{Data Analysis} to form the theoretical
categories.  At each iteration the new emergent theory is updated following a
formal procedure (\textit{Theory Generation}), and after verifying that
\textit{Theoretical Saturation}\footnote{The point at which executing more
interviews wouldn't bring any additional value for constructing the theory.} of
categories has been achieved, we finally proceeded to \textit{Theory
Validation}.

The whole procedure has been executed simultaneously in pair on the same screen
by the first two authors, handling conflicts by reviewing the rationale of
decision with the third and fourth authors. When necessary we performed an  in-
depth review of the design of the research and data collected during the
execution process.

The process details are described in the following subsections, structured
according to the five macro phases depicted in bold in Figure
\ref{fig:gt:completemethodology}.

%------------------------------------------------------------------------------%
\subsection{Design and Execution} \label{desex}

This case study started based on a SMS of the literature regarding software
development in startups \cite{SMS}. By means of this SMS we refined our research
questions. The research questions have been narrowed down to outline the domain
of our study, even though still broad enough to allow emerging new theories and
to adjust the scope of the research \cite{Corbin1990}.

The sampling of candidate companies took place in two distinct phases. At first
we executed an initial convenience sampling \cite{Dawson2009}, which led to the
identification of eight companies. Then we included five additional startups
during the theory formation process (theoretical sampling) iteratively improving
the sample according to the emerging theory. The characteristics of the sampled
companies are reported in Table \ref{t_interviews-stats}.

\begin{table*}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Companies overview}
\label{t_interviews-stats}
\centering
\begin{tabular}{|c||c||c||c||c||c|}

\hline
    ID & Company age & Founding & Initial & Current & First product \\
    & (months) & team & developers & employees & building time \\
                    & & & & & (months) \\
   \hline
C1 & 11 & 4 & 2 & 11 & 6 \\
C2 & 5 & 2 & 2 & 6 & 3 \\
C3 & 18 & 4 & 4 & 4 & 6 \\
C4 & 17 & 3 & 2 & 11 & 6 \\
C5 & 20 & 2 & 1 & 4 & 12 \\
C6 & 30 & 3 & 2 & 4 & 1 \\
C7 & 12 & 2 & 1 & 7 & 4 \\
C8 & 24 & 4 & 3 & 16 & 4 \\
C9 & 5 & 5 & 4 & 5 & 1 \\
C10 & 43 & 6 & 4 & 9 & 4 \\
C11 & 36 & 3 & 3 & 6 & 2 \\
C12 & 12 & 3 & 3 & 3 & 3 \\
C13 & 24 & 2 & 2 & 20 & 3 \\

\hline
\end{tabular}
\end{table*}

All companies, except C10, were founded within the last three years (2009-2012),
%MUN: I've shortened the text and added absolute numbers. Is 2009-2012 correct?
%CG: right by an average of 3 starting members, which were in majority
developers.  Moreover, the number of current employees shows how, to different
degrees,  companies expanded the initial teams. All companies, except C5,
brought their  first product to market in less than 6 months from the idea
conception. The  products consist of web-applications, launched in different
nations and market  sectors. The growing team size and the publicly available
data suggest an  healthy status of the businesses.

We executed the case study online, supported by tools for video conferencing,
recording each complete session. The interview subjects were the companies' CEOs
or CTOs. We followed a step-by-step workflow, consisting of the actual
interview, collection of additional material, preparation of the customized
follow-up questionnaire and the iterative adjustment of the interview package
artifacts\footnote{The interview package and follow-up questionnaire has been
released under MIT License \cite{MITLicense} and it is available for download on
Github \cite{GitHubInterviewPackage}. We encourage anyone interested in pushing
this work forward, to fork the repository and contribute to it. If used in a
research context, please inform us in order to be able to track where and how
the interview package has been used.}.

The follow-up questionnaires were designed to capture additional data, gather
missing information and confirm interview results by triangulation. The
questionnaires have been tailored to each startup, partially taking advantage of
the repertory grid principles \cite{Edwards2009}. To obtain better chances of
quickly obtaining responses, questionnaires have been sent to respondents
immediately after transcribing the interview results. Most of the follow-up
questionnaires have been collected within one day after the conclusion of the
last interview.    
%MUN: I don't understand this: you %collected the data from questionnaires 
%after ALL interviews have been finished, but on the other hand
%you said the overall process was iterative  (data analysis and triangulation
%let themes emerge which in turn were used %in subsequent interviews).  
%CG: fixed...

%------------------------------------------------------------------------------%
\subsection{Data collection} The data collection was designed to allow
triangulation, which integrates  multiple data sources converging on the same
phenomenon. After transcribing the  interview, we extracted conceptual
relations, integrating them with the  questionnaire's results. A well structured
case-study database allowed us to  easily retrieve and seek for information,
assembling the evidence from the case  study reports, as described also in
\cite{Yin1994}. The database has been stored  and constructed using the
qualitative data analysis software package  AtlasTI\footnote{Available online at
\url{http://www.atlasti.com/}.} - one of  the most suitable tools for GT
\cite{Coleman2007}.  We overlapped interviews with questionnaire results to
adjust the data  collection and take advantage of emergent themes and reveal
possible  inconsistencies. The data was analyzed simultaneously and with
flexibility in  mind in such a way that adjustments were made according to the
emerging  findings.
%------------------------------------------------------------------------------%
\subsection{Data analysis} Before starting the analysis, a data ordering
procedure was necessary since  interviews were spread across a multitude of
topics. Therefore transcripts have  been structured in thematic areas
accordingly to different topic cards used  during the interviews. We proceeded
horizontally between same thematic areas of  transcripts to be able to identify
a better number of similar concepts, rather  than going through an entire
transcript at one time.

Once the data were ordered, we executed the process of coding interviews,
following the steps listed below:

\begin{compactitem}

\item Labels were assigned to raw data, and a first low-level conceptualization
was carried out using both in-vivo and open coding \cite{ColinRobson2009}. 
\item Concepts were grouped together into theoretical categories and  
subcategories. By means of axial coding we first described the different  
relations between subcategories, and then relations between subcategories and  
categories. 
\item Categories were refined several times in the attempt to create different  
levels of abstraction and adjusting concepts, aided by a simple knowledge  
management tool. 
\item Consistency among categories was validated by exploring and
analyzing  links among subcategories by means of selective coding. 
\item The core category - the one with the greatest explanatory power - was  
identified by analyzing the causal relations between high-level categories. 
\end{compactitem}

During data extraction we used the technique of in-vivo coding since
\textit{``direct quoting from the transcript gives more expressive power to the
data''} \cite{ColinRobson2009} combined with the more descriptive procedure of
open coding. Following the example of other grounded theories, developed in
related areas such as Information Systems \cite{Orlikowski1993} and Software
Process Improvement \cite{Coleman2006}, we performed the high-level
conceptualization during creation of categories, in the process of refining
axial and selective coding.

As we were iterating through the interviews, we analyzed new data accordingly,
by updating codes and categories on necessity, and taking notes in the form of
memos to adjust the new emerging theory. To improve the speed of the coding
process, we analyzed all the transcripts one part at-the-time: transcripts
followed the semi-structured interview format, and this approach enhanced the
chances of finding similar codes across different transcripts, allowing us to
quickly iterate and update categories.

Given the large number of codes, categories, properties, propositions and
related questions that evolved from the analytical process \cite{Corbin1990},
an important activity that helped the coding process was memoing. It constituted
an important component involved in the formulation and revision of theory
during the research process. For this purpose we made use of three different
types of memos: code memos, theoretical memos and operational memos. The first
type of memo is related to the conceptual labeling of open codes, whilst the
second type concerns axial and selective coding, and thus focusing on paradigm
features. Finally, operational memos contain directions relating to the evolving
emerging theory.

After the coding process the first representation of the experience map
identified in GT is constructed by means of a theoretical framework. The
theoretical framework is presented in the form of a network of categories and
subcategories that are linked together according to cause-effect relationships
\cite{Corbin1990}. During the formation of the theoretical framework the
researchers operated by means of a  bottom-up approach. From the empirical data
and coding process, the framework developed into two different levels: a
detailed level representing the network of subcategories (identified mainly by
the axial coding process), and a high-level representation of the main
categories network (identified mainly by the selective coding process).

%------------------------------------------------------------------------------%
\subsection{Theory generation}

As mentioned in Section \ref{desex}, emergent theories have been tested by
integrating additional companies into the sample, selected following the
principle of theoretical sampling \cite{Yin1994}.

The process of theory generation took place at each iteration together with
interview execution, where we systematically analyzed the theoretical framework
by means of the paradigm model introduced by Corbin \cite{Corbin1990}. The
paradigm model is composed by:

\begin{compactitem} 
\item \textit{Causal conditions}: the events which lead to the occurrence of  
the phenomenon, that is our core category 
\item \textit{Context}: set of conditions in which the phenomenon can be 
extrapolated
\item \textit{Intervening conditions}: the broader set of conditions from that
the phenomenon can be generalized 
\item \textit{Action/interaction strategies}: the actions and responses that  
occur as the result of the phenomenon 
\item \textit{Consequences}: specification of the outcomes, both intended and
unintended of the actions and interaction strategies 
\end{compactitem}

Within the limits of the critical bounding assumptions, the role of the
generated theory is to explain, predict and understand the underlying
phenomenon.

%------------------------------------------------------------------------------%
\subsection{Theory Validation} \label{rm:val} 
%MUN: put here what I wrote in the comment in Section "Theory validation" (i.e. 
%the section following the GSM)
%CG: fixed 
%MUN1: This description is way too long and verbose. It should be MAX 1 column, 
%and focus on what you present in Section 5. Theory Validation. Until the  
%second-last paragraph you don't talk about that at all! Remove the "theory" on 
%GT validation and what you did *during* the GT process. Keep what you need  
%to motivate 
% - 5.1 Comparison of frameworks 
% - 5.2 Theoretical categories and existing literature 
% - 5.3 Confounding factors 
% - 5.4 Static assessment of the GSM 
%You need to explain why you did this and how it contributes to the validation  
%of the theory. CG: fixed, removing the initial verbose part.

Presenting GT is challenging for a researcher, who must pay attention to
structure the included level of detail, and to the way data is portrayed to
display evidence of the emergent categories.  In order to assess our research
and determine if the GT is sufficiently grounded, we used a systematic technique
to assess the theory.

Strauss and Corbin provided a list of questions to assist in  determining how
well the findings are grounded \cite{Strauss1998}. They  questions are as
follows:

\par\vspace {0.4cm} 
\begin{compactenum} 
\item Are concepts generated, and are the concepts systematically related? 
\item Are there many conceptual linkages and are the categories well developed? 
\item Is variation\footnote{Variation refers to the variety of contexts to  
which the theory can be applied.} built into the theory and are the conditions  
under which variation can be found built into the study and explained? 
\item Are the conditions under which variation can be found built into the  
study and explained? 
\item Has the process been taken into account, and does the theory stand the  
test of time? 
\item Do the theoretical findings seem significant, and to what extent? 
\end{compactenum}
\par\vspace {0.4cm}

Answering the above questions, the concepts were generated according to the
coding process. They were systematically related through the use of a network
diagram (see Figure \ref{fig:gt:completemethodology}), by which we established
the relations between the concepts. At each iteration of the  grounded theory
process, we considered and examined a concept within different  conditions and
dimensions, trying to incorporate data from a broader range of  practitioners.

All the linkages and categories were constructed by the use of Atlas.TI and
compared according to the data analysis process. Variations of the generated
theory were applied according to raw grounded data with the aid of codes and
memos. Moreover during this study extensive explanation of the concepts were
presented in-vivo statements as reported by practitioners.

The process has been designed by the researches in different steps, explaining
and making clear the purpose of each of them. Thus, the same processes enable
other research to conduct the same study within similar contexts. Moreover
\textit{comparison with the state-of-the-art} has been applied in order to
validate the theory and enhance odds of its applicability within a wider  time-
frame. By this comparison we  highlight the areas which have been neglected  by
existing studies, providing possible directions for future studies (see
subsection \ref{sect:theory:validation:others}). Moreover we study the
confounding factors which could interfere with the application of the GSM model
(see subsection \ref{sect:an:comp:cat-vs-literature:conf}).

Then, we performed an \textit{evaluation with empirical data} (see subsection
\ref{sect:theory:validation:stats}), quantifying scores for each company
involved in our empirical study by defining metrics based on a set of rubrics
and evaluating startups accordingly, through an analysis of interview
transcripts and codes. The complete statistical procedure and provided  rubrics
to quantify the measures are illustrated in Appendix.

\section{Results: Green-field startup model}\label{res:gsm}

By analyzing the interview transcripts with the GT process we extracted the
information necessary to form the theoretical model. The initial coding process
(open and in-vivo) provided 630 unique codes\footnote{Available at
\url{https://github.com/adv0r/BTH-Interview-Package}.}, grounded to 1295
citations in the transcripts, divided according to the thematic areas. The
thematic areas were divided according to the development phases designed for the
interview process. The raw codes constitute the empirical foundation of the
theoretical categories, which has been used to form the Green-field startup
model (GSM).

The GSM captures the underlying phenomenon of software development in  early-
stage startups. The framework is formed by 128 sub-categories, clustered  in 35
groups, and finally in 7 categories at the highest level of abstraction.  By the
means of the GSM we provide explanations of the development strategies  and
engineering activities undertaken by startups.
%------------------------------------------------------------------------------%
\subsection{Framework overview} \label{res:gsm:frmov} The main concepts
representing the underlying phenomenon have been grouped  together to form high-
level categories. Figure \ref{fig:gsm} shows the network  of causal
relationships (represented by arrows) between categories (represented  by
blocks). Each node is linked to subsequent nodes which are called successors,
and preceding nodes, called predecessors. The relations between them are denoted
by the direction of the interconnecting arrows. The network is read from left to
right.

\begin{figure*}[!t] 
\centering 
  \includegraphics[width=4in]{figures/high-level}
  \caption{Green-field startup model}\label{fig:gsm}
\end{figure*}

In forthcoming the explanation of the GSM we emphasize labels of theoretical
categories using the \textit{italic} font style. For the main categories shown
in Figure \ref{fig:gsm}, when necessary we also make use of identifiers (i.e.
CATx) to facilitate the reading of this section. The network is centered around
the core category, \textit{speed-up  development}, which is the most
interconnected node in the framework reflecting  the fact that \textit{``it is
the one (category) with the greatest explanatory  power''}  \cite{Strauss1998}.
The meaning of the labels on the three arrows that  reach the core category
(efficiency, effectiveness and performance) is discussed  in detail in section
\ref{sect:theory:validation:stats}. %MUN: this refers to the validity threats
%section. Do you mean rather the "Theory Validation" section? %CG:fixed 
%MUN1: this should be \ref{sect:theory:validation:stats}, right? Section 5 is  
%quite long and you should point to the correct subsection. CG: fixed.

A contextual condition which characterizes, to some extent, each and every
startup, is the \textit{severe lack of resources}. In fact, the capabilities of
an early-stage startup to support its development activities are constrained by
an extremely limited access to human, time and intellectual resources,
eventually leading to a general absence of structures and processes.

The \textit{severe lack of resources} forces the company to focus on
implementing an essential set of functionalities and it is one of the main
reasons why the \textit{product quality has low priority} respect to other more
urgent needs\footnote{There are some exceptions where the quality aspects
actually matter and such cases will be discussed in section \ref{sect:an:comp
:cat-vs-literature:conf}
%MUN: this refers to the validity threats section. Do you mean rather the 
%"Theory Validation" section? %CG fixed 
%MUN1: Same here, point to the correct subsection. CG:fixed 
.}. At the same time, to be able to deal with such constraints, startups must  
depend on a small group of capable and motivated individuals.

As it has been unanimously expressed by respondents, from a software
perspective, the most urgent priority is to \textit{speed-up the development} as
much as possible by adopting an extremely flexible and effective
\textit{evolutionary approach}. On the other hand the efficiency of teamwork is
facilitated by the low attention initially given to architectural aspects
related to product quality. This allows startups to have a faulty but
functioning product, which can quickly be introduced to the market beginning
with the in-house implementation of the prototype from day-one.

The initial employees are the ingredients which enable high levels of
performance in software development. To support a fast-paced production
environment, engineers are required to be highly committed, co-located,  multi-
role, and self-organized. In other words \textit{team is the catalyst of
development}. With an essential and flexible workflow, which relies on tacit
knowledge  instead of formal documentation, startups are able to achieve very
short  time-to-market cycles. However, each line of code, written without
following  structures and processes, contributes to grow the \textit{accumulated
technical  debt}, which is further increased by having almost non-existing
specifications,  a minimal project management and a lack of automated tests.

Despite the fact that the consequences of such debt are not clearly perceived
in the initial stages (where finding the product/market fit as quickly as
possible is the most important priority), startups, which survive to subsequent
phases, will likely increment their user-base size, the product size, and the
number of developers. This will require the company to eventually pay the
\textit{accumulated technical debt}, confronting the fact that an
\textit{initial growth hinders productivity}.

In the following subsections we explain in detail all the categories presented
in Figure \ref{fig:gsm}, and conclude in subsection \ref{res:gsm:th} with the
final theory. In the explanations we use identifiers of the companies presented
in Table \ref{t_interviews-stats} (i.e. Cx) to highlight statements recorded
from the interviewees.

%------------------------------------------------------------------------------%
\subsection{Severe lack of resources}\label{res:gsm:cat7} 
The concept of \textit{severe lack of resources} characterizes the uncertainty  
of the development strategies in startups and it is composed by 
three subcategories: \textit{time-shortage}, \textit{limited human resources} 
and \textit{limited access to expertise}.

Since startups want to bring the product to market as quickly as possible, the
resource they are the most deprived of is time. Startups operate under a
constant time pressure, mainly generated by external sources
(\textit{investor(s) pressure}, \textit{business pressure}) and sometimes
internal necessities such as \textit{internal deadlines} and \textit{demo
presentations at events}. In this regard, C3 commented: \textit{``Investors
wanted to see product features, engineers wanted to make them better. Finally
the time-to-market was considered more important and the team interests was
somehow sacrificed.''}

In addition, to compensate the \textit{limited human resources},  practitioners
need to empower \textit{multi-role and full stack engineers}, as confirmed by
C1: \textit{``Everyone was involved in any tasks, from mobile to web
development, organizing themselves in choosing the part to implement''}.

The extent to which startups have access to specialized knowledge - both
internal and external to the company - is extremely reduced when compared to
traditional established software companies. Therefore, in order to partially
mitigate the \textit{limited access to expertise}, startups rely on the external
aid of mentors or advisors. Under these strict limitations, most of the
decisions related to software development are fundamentally trade-off
situations (confirming Himola's results \cite{Hilmola2003}).
%------------------------------------------------------------------------------%
\subsection{Team is the catalyst of development} \label{res:gsm:cat4} 
Among the different aspects fostering the speed of the development process, 
the startups' main focus is on the characteristics of the initial team.

In startups \textit{developers have big responsibilities}. In fact, the
\textit{limited human resources}, discussed in CAT7, cause the team-members to
be active in every aspect of the development process, from the definition of
functionalities to the final deployment.

Early engineers in the founding team of a startups are typically  \textit{multi-
role and full-stack engineers}: they are full-stack for being able  to tackle
different problems at different levels of the technology stack
(\textit{generalists developers instead of specialists}) and multi-role since
usually \textit{engineers are responsible for marketing/sales/development}. As
remarked by  C11: \textit{``Instead of hiring gurus in one technology, startups
should hire young developers, generalists, that know how to quickly learn new
technologies, and quickly move among a huge variety of tasks.''}

Moreover, having a \textit{very small and co-located development team} enables
members to operate with high coordination, control and communication, which
heavily rely on \textit{tacit knowledge},  replacing most of the documentation
with informal discussions. Practitioners reported that keeping the development
team small helps startups in being fast and flexible, as remarked by C8:
\textit{``If you have more than 10 people, it is absolutely impossible to be
fast''}. Then, also \textit{previous working experience} and \textit{knowing
each other before starting the company} enforce the efficiency of activities by
\textit{limiting the need of formalities between team-members}.

In every software company, \textit{skilled developers are essential for high
speed} in development. Especially in startups, the ``hacking culture'' and a
tendency to the ``just-do-it'' approach allow the team to quickly move from the
formulation of a feature idea to its implementation. In this regard, C1
comments: \textit{``We had a hacker culture/environment, people hacking stuff
without formally analyzing it, but breaking it down and finding a way around.
Ticket on trello and go.''}

A \textit{limited access to expertise} forces the team to rely mainly on their
personal abilities, even though asking opinions to mentors is reported to be a
viable practice to maintain feasible objectives. Furthermore \textit{teams work
under constant pressure}, \textit{``up to 16 hours per day''}, %MUN: add Cx to
%this statement since it is a citation?  
%CG: not fixed. It represents a subcategory, expressed by the 70% of the  
%interviewed startups 
%MUN1: and how should the reader know that? See my problem with italics: you  
%use italics for both interview statements and (sub-)categories. 
%CG: skip as acked 
mainly constrained by a tight \textit{time shortage}.

Finally, startups present founders-centric structures, and especially in the
early-stage, it is clear that \textit{CTO/CEO background has high-impact} on the
company's development approach. For instance, in case of an academic background,
the CTO encourages the introduction of some architectural design before the
development phase. Moreover, even though the CTO/CEO starts guiding initially
the development process, most of the decision are taken consensually by all
members of the team. Then, the CTO/CEO decides only in particular situations
where some conflicts occur.
%------------------------------------------------------------------------------%
\subsection{Evolutionary approach}\label{res:gsm:cat2} 
Among different approaches to development, startups prefer to build an initial  
prototype and iteratively refine it over time, similarly to the concept of
\textit{``evolutionary prototyping''} \cite{EvProt}. The reason of using this
approach is that startups want to validate the product against the market as
soon as possible, finding the proper product/market fit. Indeed, they can focus
on developing only parts of the system they want to validate instead of working
on developing a whole new system. Then, as the prototype is released, users
detect opportunities for new functionalities and improvements, providing their
feedback to developers.

Since \textit{flexibility and reactiveness are the main priorities}, the most
suitable class of software development approaches are clearly highly
evolutionary in nature. In fact, as \textit{uncertain conditions make long-term
planning not viable}, startups cannot base their work on assumptions without
rapidly validating them, releasing the product to market. The uncertainty is
first of all in the team composition. Since the teams are typically very small
and the project knowledge is not documented, even a little change in their
composition (e.g. a developer falls ill) can have a very high impact on the
overall product development. Furthermore, startups operate in a continuous
evolving environment in terms of competitors and targeted market sectors. Then,
to obtain a competitive advantage in the market, startups typically make use of
cutting-edge solutions, characterised by an evolution that cannot be foreseen in
the long run. However, a special role in daily decisions is covered by user
feedback and requests, which represent the main drivers for defining the product
features in the short-term.

To obtain fast user responses and quickly \textit{validate the product},
startups \textit{build a functioning prototype and iterate on it} over time.
Paraphrasing C4, \textit{``[\ldots] you should start with something that is
really rough and then polish it, fix it and iterate. We were under constant
pressure. The aim was to understand as soon as possible the product market/fit
iterating quickly, adjusting the product and releasing fast.''}

The companies focus on building a small set of functionalities to include in
the first version, and \textit{progressively roll-out to a larger number of
people} with very \textit{small iterations} (confirmed again by C4: \textit{``we
deploy from 5 to 20 times a day''}).

The objective of this evolutionary approach is to avoid wasting time in  ``over-
engineering the system'' and building complex functionalities that have  not
been tested on real users. By releasing a very small number of good-enough
functionalities (see CAT3) the startup can verify the suitability of the
features and understand how to adjust the direction of product development
towards actual users' needs. The first version of the product is typically a
prototype which contains the basic functionalities developed with the least
possible effort that can validate critical features, enabling the startup's
survival in the short term (recalling  Ries' definition of \textit{Minimum
Viable Product} (MVP) \cite{Ries2011}).

Supported by \textit{direct contact and observation of users},
\textit{automated feedback collection} and \textit{analysis of product metrics},
startups attempt to \textit{find what is valuable for customers} with activities
in agreement with Blank's \textit{Customer validation} stage \cite{Blank2005}.

%------------------------------------------------------------------------------%
\subsection{Product quality has low priority}\label{res:gsm:cat3}
The interests of software startups, related to the product, are mainly  
concentrated on building a  \textit{limited number of suitable functionalities} 
rather than fulfilling restrictive non-functional requirements. This strategy  
allows them to quickly release simple products with less need of preliminary  
architectural studies.

Exploring quality aspects considered during the development process, the only
important concerns are expressed towards the user experience\footnote{According
to the ISO 9241-210 (Ergonomics of human-system interaction), UX is defined as
\textit{``a person's perceptions and responses that result from the use or
anticipated use of a product, system or service''.} Even though this definition
leaves so much to interpretation, we can refer to it as presented in the
glossary.}, in terms of \textit{ease of use, attractiveness of the UI} and most
of all, \textit{smooth user-flow without interruptions}. The fact that
\textit{UX is the only important quality} is remarked by C11: \textit{``When a
user needs to think too much on what action should be done next, he will just
close the application without returning''} and confirmed by C3: \textit{``If the
product works, but it is not usable, it doesn't work''.}

The extent to which quality aspects are taken into account might heavily depend
upon the market sector and the type of application. Nevertheless, realizing high
level of UX  is often the most important attribute to consider for customer
discovery of  evolutionary approaches in view of the \textit{limited human
resources} and \textit{time shortage}, presented in CAT7. C4 confirms:
\textit{``None of the quality aspects matter that much as the development speed
does.''}

To achieve a good level of UX while dealing with lack of human resources and
time shortages, startups can analyze similar products of bigger companies which
can afford more rigorous usability studies. Then, the users' feedback and
product metrics begin to have a central role in determining the level of UX
achieved. Product metrics are typically web-based \textit{statistical hypothesis
testing}, such as A/B testing \cite{AB}.
%MUN: add reference or explain shortly what A/B testing is %CG: fixed

Other than UX, some other factors can influence the quality concerns of
development:

\begin{compactitem}

\item The \textit{efficiency emerges after using the product}, letting
engineers avoid wasting time in excessive improvements of not-validated
functionalities. The level of efficiency can be optimized after attesting the
effectiveness of the minimal set of functionalities, obtained according to CAT2
and to the concept of MVP. 
\item The \textit{product should be reasonably ready-to-scale} to be able to  
accommodate a potential growth of the user-base\footnote{Note that startups  
tackle fast growing markets which are particularly subject to sudden user  
growths}. Thanks to modern cloud services, startups can \textit{externalize  
complexity to third party solutions} achieving a good level of scalability with 
a reasonable effort. 
\item Realizing high reliability is not an urgent priority since \textit{users  
are fault-tolerant in innovative beta products}. In these cases, users have  
typically a positive attitude towards the product, even though it presents  
unreliable behaviours. In this regard, the main focus of beta testing is 
reducing frictions between the product and the users, often incorporating  
usability testing. In fact, the beta release is typically the first time that  
the software is available outside of the organization that developed it.  
\footnote{Detailed discussion of the impact of innovative products on the user  
satisfaction is presented in subsection 
\ref{sect:an:comp:cat-vs-literature:conf}}. 
%MUN: do you use the Kano model in Theory Validation? Change text to refer to  
%the correct section. [After I've read the respective section, I'd suggest to  
%remove the Kano discussion altogether to shorten the paper. Keep the footnote, 
%reference the correct section, but remove "according to the Kano model] 
%CG: fixed 
\end{compactitem}

Finally, startups implement a \textit{limited number of suitable
functionalities}, which in view of \textit{limited human resources} and
\textit{time shortage}, represent a viable strategy to focus on shortening
time-to-market.
%------------------------------------------------------------------------------%
\subsection{Speed-up development}\label{res:gsm:cat1} 
As remarked initially, \textit{speed-up development} represents the core  
category of the GSM. Firmly grounded as the primary objective of startups, it  
shows the most important characteristic of developing software in the early  
stage.

As mentioned before, in order to \textit{speed-up development}, startups adopt
evolutionary approaches supported by a solid team focusing on implementing a
minimal set of suitable functionalities. Startups \textit{keep simple and
informal workflows} to be flexible and reactive, adapting to the fast changing
environment. While a rigorously defined workflow generally consists of an
established sequence of well-defined steps to follow, startups, by contrast,
adopt informal workflows. This choice is mainly dictated by contextual factors
characterized by the general conditions of unpredictability and uncertainty. The
adoption of informal workflows is facilitated by the fact that teams are
typically self-organized and \textit{developers have big responsibilities} .

Additionally, possible planning activities are refrained by the aim to shorten
time-to-market, as reported by C8: \textit{``Speed was the essence so we didn't
plan out too many details''}. To deal with such unpredictability, startups
prefer to take decisions as fast as possible, mainly by means of informal and
frequent verbal discussions.

Despite Agile principles appear to be suitable to embrace changes, especially
in the early-stage, development practices are often perceived as time-wasters
and ignored to accommodate the need of releasing the product to market as
quickly as possible. In this regard, to maintain the simplest and most informal
approach, startups aim to \textit{find the product/market fit quickly} and
implement a \textit{limited number of suitable functionalities}. This approach
is considered possible also in view of a lack of systematic quality assurance
activities, since only the user experience is considered as important and other
quality aspects, such as efficiency, can be postponed after the first release.

Another strategy, which supports startups in quickly delivering products, is
the \textit{externalization of complexity on third party solutions} by: making
use of third party components (COTS) and open source solutions (both for
product components, and development tools and libraries); taking advantage of
external services; and deploying the product on external infrastructures, for
the sake of delivering a \textit{product reasonably ready to scale} for a
possible future growth.

Even though \textit{the use of well-integrated and simple tools} allows
startups to automate many activities and reduce their completion time, drawbacks
of such approaches are the increased concerns of interoperability issues. A
category of tools is represented by advanced version control systems, which are
not only used to manage the code-base, but for many other purposes such as:
assigning tasks; trace responsibilities; configuration management; issue
management; automatic deployment; discussions and code reviews.

Startups can further improve development speed by making \textit{use of
standards and known technologies} which are widely recognized, well tested, and
supported by strong communities. Moreover, the use of widely adopted standards
and frameworks reduces the need of a formal architectural design since most of
implementation solutions are well documented and ready-to-use. This is confirmed
by C1: \textit{``as long as you use Ruby standards with the Rails framework, the
language is clean itself and doesn't need much documentation''.}

Other important factors that positively impact the speed of development are the
team's desire to: \textit{create disruptive technologies}; to
\textit{demonstrate personal abilities}; and to \textit{have the product used in
the market}. As reported by practitioners, these factors are essential to
enhance the morale of developers and therefore to achieve higher team
performances. By contrast, especially in the typical sprint-based environments
of Agile, when a developer is not \textit{able to meet deadlines} the morale
goes down, hindering the development speed.

Finally the constant pressure under which the company regularly operates, leads
the team to often \textit{work overtime to meet deadlines}. But as reported by
practitioners, such way of working can be an effective strategy only in the
short term since it can lead to poorly maintainable code and developer burnouts
in the long run.
%------------------------------------------------------------------------------%
\subsection{Accumulated technical debt}\label{res:gsm:cat5} 
Startups achieve a very high development speed by radically ignoring aspects  
related to documentation, structures and processes. However, although these  
aspects are not considered important in the very first stages, they will become
increasingly more important for the productivity in the long-term, as we
illustrate in CAT6.

Instead of traditional requirement engineering activities, startups make use of
\textit{informal specification of functionalities} by means of ticked-based
tools to manage low-precision lists of features to implement, written in form of
self-explanatory user stories \cite{AgilePlan}. Practitioners intensively use
physical tools such as post-it notes and whiteboards, which help in making
functionalities visible and prioritizing stories based on personal experiences.
C4 comments \textit{``[\ldots] it is the only way. Too many people make the
mistake of sitting down and write big specs and then they build it for four
months, realizing the product is not valuable only at the end.''}

Since startups are risky businesses by nature, even less attention is given to
the traditional phase of \textit{analysis}, which is replaced by a \textit{rough
and quick feasibility study}   \textit{``only sometimes and however
informally''}, as stated by C5.  First of all it is generally \textit{hard to
analyze risks with cutting-edge technologies}. To find out the feasibility of
such cutting-edge projects, startups attempt a first implementation with rough
and informal specifications, assuming that project's complexity will remain
limited to a small number of functionalities, as discussed in CAT3. Additionally
by \textit{keeping the product as simple as possible} and learning from
competitors' solutions and mistakes, practitioners can use their \textit{past
experiences in similar contexts to help assessing feasibility} of the project.
Finally, to avoid restrictions on the flexibility of the team, potential
limiting decisions are taken only when strictly necessary and anyhow as late as
possible.

Another important factor which contributes to \textit{accumulate the technical
debt} is the general \textit{lack of architectural design}, substituted by
\textit{high-level mockups and low-precision diagrams}, \textit{using
modularization and frameworks from day 1} and \textit{describing critical
interactions with third-party components only}. In particular, the use of  well-
known standards, frameworks and conventions limits the need of formal UML
\cite{UML} diagrams and documentation, and provide a minimum level of
maintainability costs. Additionally, since quality aspects are not a main
concern (see \textit{limited number of functionalities} and \textit{efficiency
emerges after using the product} in CAT3) having a well-structured architecture
remains a secondary priority, and design is conducted only when strictly
necessary.

A similar attitude towards verification and validation brings startups to a
severe \textit{lack of automated testing}, which is mainly replaced by manual
smoke tests executed by: trying the product internally; seeking for critical
faults; and letting early adopters report non-critical bugs.  Paraphrasing C3,
\textit{``Trying the product internally allows us to get rid of 50\% of bugs of
important functionalities. Meanwhile users report bugs of secondary
functionalities, eventually allowing us to mitigate the lack of testing. Indeed,
staying one week in production enables us to identify 90\% of bugs''.} The lack
of complete automated tests is partially also motivated by the fact that
\textit{users are fault-tolerant in innovative beta products} and  by the
limited number of functionalities, which allow the team to easily control
critical bugs. However, in certain cases where components of the system might
cause loss of data or severe damages to the product or users, engineers realize
a reasonable level of automatic testing. In such cases, aided by modern
automatic tools, they can quickly assess the status of the system integration as
they add new functionalities to the product.

Furthermore, a rigid project management is perceived as \textit{``waste of
time''} which hinders the development speed since the \textit{uncertainty makes
formal scheduling pointless} (C9  reports that \textit{``initial chaos helps to
develop faster''}). Startups' \textit{minimal project management}  is supported
by keeping: \textit{internal milestones short and informal},  \textit{low-
precision task assignment mechanism}  and a very low attention for  project
metrics (paraphrasing C13, \textit{``the only track of progress was made  by
looking at closed tickets''}). In this context \textit{only a final release
milestone is viable}, which helps practitioners to remain focused on short term
goals and put new features in production. To further \textit{speed-up
development}, startups \textit{simplify issue management by integrating it with
feature management}. Then, the \textit{limited number of functionalities} and
the {use of standard/known technologies} with a \textit{simple workflow} are the
main reasons for  not establishing a heavy project management process since the
project and  technologies can be autonomously managed by the development team
(with  characteristics of: co-location, self-organization and very small team
size).  %MUN: do you mean small team size? %CG: fixed

Finally, one of the categories, which most contributes in growing the
\textit{accumulated technical debt}, is the substantial  use of informal and
verbal communication channels on a daily basis. The high co-location and the
fast paced development approach increase the volume of \textit{tacit knowledge}
and severe lack of any kind of documentation.  C4 observes in this regard that:
\textit{``[\ldots] the issue of having documentation and diagrams out of the
source code is that you need to update them every time you change something.
There is no time for it. Instead, there is a huge pay off in having a code that
is understandable itself.''} This approach is supported by the fact that
startups make use of   \textit{simple and informal workflow},
\textit{standard/known technologies},  \textit{very small and co-located
development team} with \textit{limited need of  formalities}.
%------------------------------------------------------------------------------%
\subsection{Initial growth hinders performance}\label{res:gsm:cat6} 
The lack of attention given in the first phases to engineering activities  
allows startups to ship code extremely quickly (see CAT5). However, if  
successful, the initial product becomes more and more complex over time, the  
number of users increases and the company size starts to grow. Under these  
circumstances the need of controlling the initial chaos forces the development  
team to return the \textit{accumulated technical debt} instead of focusing on  
new users' requests. Hence, \textit{initial growth hinders performance} in terms 
 of number of new user stories implemented in a certain amount of time, which  
brings new suitable functionalities to the users.

When the number of users increases, customers become more quality demanding and
also scalability issues start to arise. Usually, \textit{company and user size
grow} when new business events occur, such as: a \textit{new round of funding},
a possible \textit{acquisition}, the release of a \textit{competing product on
the market}, or when the project is \textit{open for the first public release}.
As a consequence, the \textit{increasing number of users} creates a growing
number of requests and expectations. Therefore, whereas the project lacks of
minimal processes, suddenly, \textit{the current team is not able to manage
increased complexity} of new functionalities to implement and maintain the
codebase.

Subsequently, practitioners start considering the need of project management
activities, also in view of \textit{hiring new staff members}, as discussed by
C13: \textit{``(project management) is strictly necessary if you radically
change the team or when the team grows.  The informal communication and lack of
documentation slow down the process afterwards''}.  Project management becomes
further important when the \textit{focus shifts to business concerns}. Part of
the effort, which was initially almost entirely dedicated to product
development, is required to move to business activities. Moreover the
availability of project information becomes an important issue because of the
accumulated \textit{tacit knowledge}, which hinders the ability of new hires to
quickly start the development of project tasks.

When the company faces growth, the partial and informal engineering activities
that have been conducted during the first phases of \textit{rush} development
(refer to CAT5 subcategories \textit{minimal project management},
\textit{informal specification of functionalities}, \textit{rough and quick
feasibility study}, \textit{lack of automated testing}, \textit{tacit
knowledge}), force the company to \textit{pay off the accumulated technical
debt}. Under these circumstances, startups need to be able to cope with the
renewed needs and expectations of both the company (internal necessities) and
customers (product).

Another factor that slows down performance is caused by the fact that
\textit{portion of codes needs to be rewritten} and \textit{substantial
refactoring of the codebase} is required by increasing product demands.

Practitioners realize that some decisions taken (or not taken) during the
\textit{rough and quick feasibility study} before starting the implementation,
have led to negative consequences on the long term performance and
maintainability of the product. Additionally, the initial decisions of the
product design might not be able to satisfy the increased demands of the
product's users and developers (\textit{lack of architectural design}). The
combination of these factors leads to the need of \textit{re-engineer the
product}.

By re-engineering the systems, startups aim to \textit{increase the scalability
of the product/infrastructure} and starting to \textit{standardize the codebase
with well-known frameworks}.

C7 reports that: \textit{``To mitigate this (lack of frameworks) I had to make
a schema for other developers when we hired them. We had to do a big
refactoring of the codebase, moving it from custom php to Django, normalizing
the model and making it stick with the business strategy. I had the code in
different php servers communicating via JSON, some engineering horror. Now  that
we are fixing it, it's really painful. We had to trash some code. However I
don't regret that I didn't make this choice sooner, it was the only way''}.

The \textit{fear of changing a product, which is working}, arises when product
complexity increases. The changes to the codebase, to support bug fixing, become
highly interrelated with other functionalities and difficult to manage because
the product is poorly engineered (see CAT5). Therefore, the fear arises from
changing a validated product that might bring changes to the user responses.

The increasing number of feature requests causes the \textit{growing necessity
of having a release plan}, rationalizing the rollout of new features.

Therefore, startups begin to \textit{partially replace informal communication
with traceable systems} and \textit{introduce basic metrics for measuring
project and team progress} in order to establish an initial structured workflow.
In conclusion, because the increasing number of users causes issues in
scalability and reliability, the \textit{need of more structured workflow}
becomes of major relevance. C11 states: \textit{``Yet, it is still better to
have a reasonable drop-down in performance when team grows than lose time in the
beginning''}.
%------------------------------------------------------------------------------%
\subsection{Theory}\label{res:gsm:th}
In order to explain and understand the development strategies in early-stage  
software startups we construct the theory generated and supported by the above  
presented GSM:

\begin{theory} Focusing on limited number of suitable functionalities, and
adopting partial  and rapid evolutionary development approaches, early-stage
software startups  operate at high development speed, aided by skilled and
highly co-located  developers. Through these development strategies, early-stage
software startups  aim to find early product/market fit within extremely
uncertain conditions and  severe lack of resources. Nevertheless, by speeding-up
the development process,  they accumulate technical debt, causing an initial and
temporary drop-down in  performance before setting off for further growth.
\end{theory}

As discussed in section \ref{res:val} %MUN: this refers to the validity threats
%section. Do you mean rather the "Theory Validation" section? %CG:fixed
, we formed the theory by applying the paradigm model, introduced by Corbin
\cite{Corbin1990}. The theory is composed by different elements reflecting the
paradigm model:

\begin{compactitem}

\item \textit{Causal conditions} are represented by three main conceptual
categories: \textit{product quality has low priority}, \textit{evolutionary
approach} and \textit{team is the catalyst of development}. 
\item \textit{Phenomenon} is represented by the core category \textit{speed-up
development}. 
\item \textit{Context} is limited to web early-stage software startups  
operating in conditions of severe lack of resources aiming to early
find  product/market fit. 
\item \textit{Intervening conditions} are  summarized by the extremely  
uncertain development environment. 
\item \textit{Action and interaction strategies} are represented by the  
accumulation of technical debt.
\item \textit{Consequences} lead to a temporary performance drop-down.
\end{compactitem}

%------------------------------------------------------------------------------%
\section{Theory Validation} \label{res:val}

In this section we discuss the validity of the GSM by means of 
cross-methodological observations, as discussed in section \ref{rm:val}. 
%MUN: expand this discussion a bit (motivation why you do this like you do it)  
%and move to Section 3.5. Then write here just that you perform theory  
%validation as discussed in Section 3.5 (and remove the bullet list below). 
%CG:fixed 
%MUN1: Ok, but fix section 3.5... you see that you don't motivate 5.1-5.4 at  
%all there and that the current content of 3.5 is not related to what you  
%actually do here as validation? CG: skip as acked

%------------------------------------------------------------------------------%
\subsection{Comparison with other frameworks}
\label{sect:theory:validation:others}

To validate the generalization of the theory, we describe conceptualizations
defined from the GSM that are in support of previous models developed by
Coleman \cite{Coleman2007,Coleman2008a, Coleman2008}, Baskerville
\cite{Internet} and finally Brooks \cite{BrooksJr1987}.

Starting from the Coleman's work, we refer to him since only he has,  to the
best of our knowledge, conducted a similar research in startups' context, even
though with a different focus. Coleman investigated factors in software
development that hinder initiatives of software process improvement (SPI) in a
later stage, representing also companies in expansion phase with more than 100
employees. %MUN: so he did not do research  in a startup context? You need to
%clarify what you mean by "more mature companies"... I assume you still mean
%startups, but startups at a later stage, i.e. after first product release.
%CG: fixed

In this subsection we illustrate the most important aspects considered by
Coleman, examining similar and contrasting factors with respect to the GSM. We
consider only those categories strictly related to the core category, which
relations, as described in GT methodology, generate complete explanatory power
of the generated theory.

Coleman's framework aims to highlight how managers consider two distinct kinds
of processes: \textit{essentials} and \textit{non-essentials}. The
\textit{essential} processes are the most closely linked to product development,
such as requirements gathering, design and testing. The \textit{non-essential}
processes are those that might be omitted, such as planning, estimating and
staging meetings. In particular, he discusses how practices are routinely
removed: \textit{``With most methodologies and approaches, very few stick to the
letter of them and they are always adapted, so we adapted ours to the way we
wanted it to work for us, for our own size and scale''} \cite{Coleman2008}.
%MUN: add ref to Coleman's paper where he states this %CG fixed

The GSM explores the same challenges in software startups, where the act of
tailoring processes leads developers to adopt only minimal practices, which are
most suitable for the startups context. Furthermore, CTOs' and CEOs' background
has a great impact on the speed of the development process as described in CAT4.
This is in accordance to Coleman's framework, which presents the background of
founders and development managers as main factors that affect the management
style of the development process.

Differently from Coleman's framework, the GSM doesn't present how
\textit{market requirements} can affect the conduction of processes. In this
regard, Coleman describes how the more requirements definitions are predictable
the more well-defined workflows are established. This particular aspect is
considered as a confounding factor\footnote{The identification of confounding
depends on the context in which the study is conducted and on the background
knowledge of the researchers \cite{Pearl2011}.}  in our study since we were not
able to ground this concept in the data.

But what characterizes Coleman's network is the \textit{cost of process} (core
category) and all the factors that in management contributed to the lack of
software process improvements (SPI). The \textit{cost of process} represents the
lack of formal and prescriptive workflows in development, mainly conducted by
means of verbal \textit{communication} limiting heavy \textit{documentation} and
\textit{bureaucracy}.

Coleman reports the practitioners' perception that \textit{documentation} alone
would not have ensured a complete shared understanding of project requirements.
Moreover, defined processes are perceived by managers as additional items with a
negative impact on the \textit{Creativity} and \textit{Flexibility} of the
development team. This is in accordance with our generated theory, which bases
the reasons of adopting evolutionary and low-precision engineering elements in
the flexibility and reactivity attributes of the development process in
startups.

In this regard we can paraphrase a comment recorded by Coleman during an
interview: \textit{``When we set up we had more supervisory and managerial roles
in that group than we have now and we had to scale that back which has made
things a lot more flexible. I do think you have to be nimble, quick and capable
being responsive in our position. That works well and I don't want to lose
it.''} \cite{Coleman2007}.

Verbal \textit{communication}, lack of heavy \textit{documentation} and
\textit{bureaucracy} can easily be associated to the \textit{accumulated
technical debt} category (see CAT5). According to Coleman's framework, they
describe how startups experience the lack of main engineering activities and
documentation. Additionally, the \textit{speed-up development} category
expresses the same concept of \textit{flexibility} and \textit{process erosion}
since they have direct relation to the subcategory of \textit{keep a simple and
informal workflow}, as discussed in CAT1.

As also reported in the GSM, the definition of a \textit{``minimum process''}
is not a matter of poor knowledge and training, but rather it is the necessity
of operating with solutions that let the company move faster.  \textit{``One-
size-fits-all''} solutions have always found difficulty in  penetrating small
software organizations \cite{Staples2007}. When startups began  establishing any
SPI process, they experienced \textit{process erosions}   \cite{Coleman2008}
%MUN:  %add ref (Coleman?). You should add, for each category you use from
%Coleman, add the reference to the paper. This makes it clear that the category
%comes from him (and is not one of our categories) %CG: fixed 
, which resolved to a barely sufficient workflow to satisfy the organizational  
business needs.

Then, software startups favor the use of agile principles in support of
\textit{creativity} and \textit{flexibility} instead of SPI methods, whereby
processes need to be predictable and repeatable. Nimble and ad-hoc solutions
prevent the use of heavy bureaucracy and formal communication strategies %MUN:
%seems to contradict each other: "ad-hoc solutions prevent the use of heavy
%bureaucracy AND fast communication strategies" I agree that ad-hoc solutions
%do not produce heavy bureaucracy, but they allow fast communication
%strategies... I may be wrong, but then you need to explain it better %CG:fixed,
%you are right :) ...   
, even though the accumulated tacit knowledge is hard to manage and to transfer 
to new hires as also discussed in category \textit{CAT6}.

Coleman describes how the \textit{``management approach''} is oriented towards
\textit{``embrace and empower''} solutions in contrast with \textit{``command
and control''} style, where there is an evidence for trust in the  development
staff to carry out tasks with less direct supervision, greater delegation of
responsibilities and a more generally consensual environment \cite{Coleman2008}.
Nevertheless, software development managers and founders have still impact on
management style and indirectly on the software development process. In the case
of early-stage startups, founders are mainly software development managers as
CEOs/CTOs and  technical practitioners at the same time. As Coleman identified
the influence of the founders' and managers' background on the software
development process, the GSM similarly identifies that the CEOs and CTOs
background shapes the high-level strategies adopted in developing the initial
product. Notwithstanding, team members remain self-organized, able to intervene
in all the aspects of the development process without any direct supervision, as
discussed in CAT4.

In conclusion, by the comparison of the GSM with Coleman's framework, we have
found that despite Coleman's study contributes to define which factors hinder
the introduction of SPI in startups  %MUN: add: "of SPI in startups"? CG: fixed
his results are coherent with the theory generated within our research, focusing
at explaining the overall strategies within software development in startups.
%MUN: add "our research, focused at explaining software development in 
%general in startups." ? CG:fixed 
In this regard, a wide variability of processes is a key factor for startups.   
In contrast with some SPI processes, such as \textit{Six Sigma} which objective 
is minimizing unpredictability in the definition of workflows \cite{Sixsigma}, 
startups seem to follow the \textit{Fischer's Fundamental Theorem of natural   
selection} \cite{Fisher}.
Moving towards flexible and variable processes (easy  to adapt to the uncertain
conditions), increase the odds of ``natural  selection'' only when some
``restrictive conditions'' are met. In other words, even though ignoring SPI
practices is advisable, under circumstances of high uncertainty and severe lack
of resources, %MUN: under which circumstances? This statement sounds too
%general, i.e. don't do SPI ever... %CG: fixed  
%And the following part,I did not fully understand. Try to clarify. CG:fixed   
not following SPI does not predict major adaptiveness of startups development 
towards the customer's needs. Nevertheless, to imply a major evolutionary 
product/market fit, certain conditions explained in GSM need to be considered.

%MUN: As I wrote already in the previous draft, you need to clarify at the
%beginning of this subsection which frameworks you are going to compare GSM to.
%Now, you say that Coleman is the only author who did research, comparable to
%yours, into startups. Now there comes Baskerville and Brooks. Write at the
%beginning of this section, shortly, which frameworks/authors you compare GSM
%to and why, then create subsections for each.

%CG: partially fixed. Agreed on the beginning of the section, not agreed on  the
%division in subsections. Let's discuss it by a skype call.

Also Baskerville \cite{Internet} refers to SPI as a development approach
typically effective only in large-scale, long-term development efforts with
stable and disciplined processes. Internet-speed software development (i.e. an
agile method oriented using daily builds aimed at developing a product with high
speed) differs from traditional software development. Baskerville, by applying a
grounded theory analysis in 10 companies, found that major causal factors that
influence development are:

\begin{compactenum}

\item A desperate rush-to-market. 
\item A new and unique software market environment. 
\item A lack of experience developing software under the conditions this  
environment imposed.

\end{compactenum}

As discussed in CAT7, 1) can be mapped to \textit{time shortage} and 3) to
\textit{limited access to expertise}, whilst  2) can refer to \textit{uncertain
conditions make long-term planning not viable}, explained in CAT2. Even though
with different research focus and context of study % - understand how and why
%Internet-speed software development (i.e. an agile method oriented using daily
%builds aimed at developing a product with high speed) differs from traditional
%software development - MUN: put this sentence where you introduce
%Baskerville's work. %CG: fixed 
, Baskerville revealed similar causal factors as the GSM.

Baskerville argues that the dawn of the Internet era has intensified software
development problems by emphasizing shorter cycle times, stating to:

\begin{compactitem} 
\item Speed-up development by releasing more often the software and  
``implanting'' customers in the development environment (similar
to the  \textit{find the product/market fit} concept presented in CAT2).
\item Make heavy use of simple tools and existing components (similar to the  
concepts expressed in CAT1). 
\item Invest time in facilitating development of scalable systems by the use of 
simple but stable architectural solutions (see subcategory \textit{using  
modularization and frameworks from day 1} discussed
in CAT5).
\item Tailor the development process daily according to the intense demands for 
speed. In this regard, he argues that the trend is to skip phases or tasks that 
might impede the ability to deliver software quickly even though producing lower 
quality software.
\end{compactitem}

With a wider research focus, Brooks \cite{BrooksJr1987} discussed what
challenges are involved in constructing software products. Brooks divides
difficulties in development into essence (inherent to the nature of the
software), and accidents (difficulties attending software production but which
are not inherent). In other words essence, concerns the hard part of building a
software through activities such as specification, design, testing. Accidents
refer to the labor of representing the software or testing its representation.

Brooks claims that the major effort applied by engineers in the last decades
was dedicated towards accidents problems, trying to exploit new strategies to
enhance software performance, reliability and simplicity of development, such as
the introduction of high-level languages for programming.  Despite the great
achievements in improving development performance, the \textit{essence} property
of the software remained unaltered.

Essence difficulties are inherent to: complexity, conformity, changeability and
invisibility. Complexity  mainly refers to interaction of software modules and
elements between each others in some non-linear fashion, which increases
accordingly to the project size. Conformity refers to the duty of software to
adapt to human institutions and systems. Changeability  refers to the constant
pressure of software products to accommodate culture, market, laws and hardware
transitions. Invisibility  refers to the difficulty of representing software in
space.

Here we present the basic attacks (or mitigation strategies) presented in 1986
by Brooks on the conceptual essence ascertained within the GSM model
description:

\begin{compactitem}

\item Buy versus build- the most radical possible solution for constructing
software is not to construct it at all, taking advantage of what others have
already implemented. It is the main strategy, which enables startups to
externalize complexity to third party solutions explained within CAT1,   \textit
{speed-up development}. 
\item Requirements refinement and rapid prototyping- avoid deciding precisely  
what to build but rather iteratively extract and refine the product requirements 
with customers and users. It represents the evolutionary development approach to 
maintain a fast release-cycles during development (CAT2). 
\item Incremental development- starting from simple solutions allows 
organizations to early prototype and control complexity overtime. It represents 
the main purpose of focusing on \textit{limited number of suitable 
functionalities} examined in \textit{product quality has low priority} category 
(CAT3). 
\item Great team- people are the center of a software project and it is  
profoundly important to empower and liberate their creative mind. It is the main 
objective of empowering developers' capabilities described in \textit{team is  
the catalyst of development} (CAT4).
\end{compactitem}

Brooks' forecast about software development strategies is revealed to be
extremely accurate, according to the state-of-practice in modern startups.
%------------------------------------------------------------------------------%
\subsection{Theoretical categories and existing literature}
\label{sect:theory:validation:sms}

In this subsection we report the evaluation of our theory in contrast with the
main contributions of the studies retrieved and discussed within the SMS
\cite{SMS}. For each retrieved article we present the main results, mapping them
to categories in accordance to the GSM model. Table \ref{tab:an:literature_comp}
shows the main contributions of the 37 identified studies, where an 'X' is
placed in correspondence to the theoretical category they considered. The table
is sorted by the number of GSM categories covered by the studies.

\begin{table*}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{SMS \cite{SMS} and GSM}
\label{tab:an:literature_comp}
\centering
\begin{tabular}{|l||c||c||c||c||c||c||c||c||c|}

\hline
Author (year) & CAT1 & CAT2 & CAT3 & CAT4 & CAT5 & CAT6 & CAT7 & Count & Ref. \\
\hline
Sutton (2000) & X & X & X & X & X & X & X & 7 & \cite{Sutton2000} \\
Kajko-Mattson (2008) & X & X & X & X & X & X & X & 7 &
\cite{Kajko-Mattsson2008} \\
Crowne (2002) & X & X & X & X & X & X & X & 7 & \cite{Crowne2002} \\
Coleman (2008) & X & X & X & X & X & X & X & 7 & \cite{Coleman2008a} \\
Coleman (2008) & X & X & X & X & X & X & X & 7 & \cite{Coleman2008} \\
Coleman (2007) & X & X & X & X & X & X & X & 7 & \cite{Coleman2007} \\
Carmel (1994) & X & X & X & X & X & X & X & 7 & \cite{Camel1994a} \\
Yoffie (1999) & X & X & X & X & X & X & & 6 & \cite{Yoffie1999} \\
Zettel (2001) & X & X & X & X & & & X & 5 & \cite{Zettel2001} \\
Jansen (2008) & X & & X & & X & X & X & 5 & \cite{Jansen2008} \\
Heitlager (2007) & & & X & X & X & X & X & 5 & \cite{Heitlager2007} \\
Deias (2002) & X & X & & X & X & X & & 5 & \cite{Deias} \\
Ambler (2002) & X & X & & & X & X & X & 5 & \cite{Ambler2002} \\
Wood (2005) & X & & X & X & & & X & 4 & \cite{Wood2005} \\
Tingling (2007) & & X & & X & X & X & & 4 & \cite{Tingling2007} \\
Taipale (2010) & X & X & & X & X & & & 4 & \cite{Taipale2010} \\
da Silva (2005) & X & X & & X & X & & & 4 & \cite{Silva2005} \\
Mirel (2000) & X & & X & X & X & & & 4 & \cite{Mirel2000} \\
Midler (2008) & X & X & & X & & & X & 4 & \cite{Midler2008} \\
Tanabian (2005) & X & & & X & & & X & 3 & \cite{Tanabian2005} \\
Stanfill (2007) & X & & & & X & & X & 3 & \cite{Stanfill2007} \\
Mater (2000) & X & X & & & & X & & 3 & \cite{Mater2000} \\
Kuvinka (2011) & X & X & & X & & & & 3 & \cite{Kuvinka2011} \\
Deakins (2005) & X & X & & X & & & & 3 & \cite{Deakins2005} \\
Yogendra (2002) & X & X & & & & & & 2 & \cite{Yogendra2002} \\
Wall (2001) & X & & & & & & X & 2 & \cite{Wall2001} \\
Su-Chuang (2007) & X & & & X & & & & 2 & \cite{Su-Chan2007} \\
Steenhuis (2008) & & & & X & & & X & 2 & \cite{Steenhuis2008} \\
Sau-ling Lai (2010) & & X & & & & & X & 2 & \cite{Lai2010} \\
Kakati (2003) & & X & & X & & & & 2 & \cite{Kakati2003} \\
Himola (2003) & X & & & & & & X & 2 & \cite{Hilmola2003} \\
H\"{a}sel (2010) & X & & & X & & & & 2 & \cite{Hasel2010} \\
Hanna (2010) & X & & & & & & X & 2 & \cite{Hanna2010} \\
Bean (2005) & X & & & & X & & & 2 & \cite{Bean2005} \\
Kim (2005) & & X & & & & & & 1 & \cite{Kim2005} \\
Fayad (1997) & & & & X & & & & 1 & \cite{Fayad1997} \\
Chorev (2006) & & & & X & & & & 1 & \cite{Chorev2006} \\
\hline
Count & 29 & 22 & 13 & 26 & 18 & 14 & 20 & &\\
%MUN: add a row which shows the sum of each column. This shows how much
%categories are covered overall %CG: fixed
\hline
\end{tabular}
\end{table*}

Only 7 studies out of 37, discuss aspects related to the overall software
development process in early stage startups. The remaining part of the studies
is only partially covering the phenomenon described by the GSM. Looking at the
papers' main contribution%MUN: which results? The papers' main contribution
%or...? %CG: fixed 
, we were able to extrapolate information which confirmed the GSM categories.

We can see how the majority of the retrieved studies (29) mention issues
related to \textit{speed-up development} (CAT1) which confirms the importance of
the core category. On the other hand, we can observe that only a limited number
of studies mentions results related to \textit{product quality has low priority}
(CAT3), \textit{accumulated technical debt} (CAT5), and \textit{initial grow
hinders performance} (CAT6), suggesting directions of possible future primary
studies.

These results confirm the relevance of development teams as widely discussed in
the SE literature, here mapping 26 studies to \textit{team is the catalyst of
development} (CAT4). The importance of people has been widely discussed in other
studies in SE (Cooper \cite{Cooper1986}, DeMarco \cite{peopleware-demarco1999},
Coleman \cite{Coleman2004}, Valtanen \cite{Valtanen2008},  Adolph
\cite{Adolph2011} and Cockburn \cite{people-first-order}), advocating for the
need of empowering people, which is the element with the greatest impact on
software development. Humans, being non-linear variables \cite{Human} 
%MUN: add ref %CG:fixed
, are unlikely to follow repeatable prescriptive methodologies. As
stated by  Highsmith \cite{Highsmith2000}, \textit{``A drawback of each and
every  methodology is to expect people to behave consistently over time, when is
clearly not like that''}. Especially in early-stage startups, where the whole
company overlaps with the development team, the GSM confirms the importance of
factors related to people.

%MUN: Think of structuring this long subsection. Have a subsubsection for each
%CAT, i.e. 5.2.1 Severe lack of resources (CAT7) I've added to each category you
%mention the corresponding CATx since that   identifier you use in table 2. I
%did not check if you discuss each and every  category. Do you? If yes, ok. If
%not, why not?  CG: partially fixed, the categories CAT3 and CAT6 are not well
%covered as  discussed above.

%------------------------------------------------------------------------------%
\subsubsection{Limited resources}

Starting from the category \textit{severe lack of resources} (CAT7), many
authors describe startups facing with inescapable constraints defined by
engineering/business concerns and constant time-pressure \cite{Camel1994a,
Cugola98softwareprocesses:}. Because of their young and inexperienced working
conditions, startups present limited resources in terms of management and
strategic alliances \cite{Sutton2000}.

Crowne \cite{Crowne2002} identifies four main stages of a startup's lifecycle:
startup, stabilization, growth and maturity. Inexperienced practitioners
operating in a chaotic environment characterizes the first phase, where there is
no strategic plan for developing the product. But, moving towards the
stabilization phase, the product starts to be unreliable and requirements
unmanageable. Consequently, when startups move to the growth phase, they
initiate structuring and controlling processes, which are incrementally
integrated in the development environment. The maturity phase represents the
moment when the company is ready to introduce process improvement initiatives.
Crowne's study profoundly corresponds to different conceptualizations in the
GSM. Starting from the minimal and low-precision engineering elements, startups
start to grow the accumulated technical debt which requires fast integration of
more structured and standard processes overtime. Issues related to low-precision
engineering practices during companies' growth  are discussed in an experience
paper by Ambler \cite{Ambler2002}. He reports on  two case studies of growing
Internet startups approaching an initial public offering (IPO). %MUN:explain
%acronym %CG: fixed 
Despite the two companies were facing business and team growth, attitudes  
against the introduction of processes to structure and model the system were  
evident. The need to re-engineer the codebase lead the researchers to suggest a 
tailored version of the Rational Unified Process (RUP)
%MUN: add reference %CG: not fixed, not sure if needed since Ambler has the
%detailed explaination of his process, similar to RUP... anyway a dedicated
%resource would be an IBM website
development framework. Nevertheless, even with the aid of CASE  
tools\footnote{CASE tools are a class of software that automates many of the  
activities involved in various engineering activities.
See  \url{http://www.unl.csi.cuny.edu/faqs/software-enginering/tools.html} .},
startups failed in practicing engineering activities in view of excessive lack
of  flexibility. Developers fast moved to simpler tools such as using a white-
board,  where they were able to provide input and insights, question
assumptions, share  past experiences, and even talk about potential
implementation strategies.  
%MUN: Candidate for removal since details can be found in Ambler anyway: "In  
%general the development remained highly iterative, where project teams started 
%to have a little modeling and testing only when necessary. They didn't take  
%the serial approach of completely developing and accepting the requirements  
%model, analysis model and so on. This let them react swiftly and adapt  
%rapidly according to changes into their highly competitive markets. The  
%researchers report \textit{``modeling was streamlined''}, that is modeling  
%what was needed and go straightforward to coding, avoiding analysis  
%paralysis." %CG: ok

%------------------------------------------------------------------------------%
\subsubsection{Minimal project management}

The use of whiteboards in working environments supports \textit{use of  well-
integrated and simple tools}. In both the startups studied by Ambler,
documentation was minimal, finding the ``sweet spot'' to have just barely enough
tools and artifacts to meet their needs. Generally, the use of \textit{keep a
simple and informal workflow} endorses the use of low-precision engineering
practices which lead to speed-up development initially, but with negative
effects on the accumulated technical debt of a minimal project management and
consequently on the final company growth.

Kajko-Mattsson et al. \cite{Kajko-Mattsson2008} investigated a Swedish software
startup company involved in mobile applications, reporting the following issues:
a lack of requirements gathering process; release cycles and their length were
not defined; the releases and their scope were not planned; lack of control over
the change requests; lack of process control in terms of  absence of any
documentation to track the status and progress of the process; defective
releases in terms of lack of testing; poor communication since the communication
inside the company was informal and not documented at all. In the same study
they designed a process improvement model, based on three main phases: Evaluate,
Plan and Change.

%MUN: the whole paragraph needs to be summarized to its essence. Details can be
%read in the referred paper.  
%CG: partially fixed, removed the detailed process description 
In the evaluation phase, the status of the development workflow
within the  company  is analyzed. The researchers documented a state of chaos
where any  initiatives of quantitative methodology assessment would have been a
waste of  time. Therefore, they applied a qualitative methodology to define a
model of the  release management process.  
%They moved towards the plan phase where the  %researchers decided to develop a 
%model to manage releases, introducing a process composed by: release scope 
%preparation, release planning, release development, system testing, 
%acceptance testing and release deployment.  
Despite the effort spent by the researchers to establish a development workflow, 
practices were not fully conducted even though some marginal benefits have been 
reported: planned major release from four to six months; planned minor releases 
from two to four weeks; encompassing urgent corrective releases from one to five 
days.  
Regarding the previously mentioned problems, only  partial improvements have 
been achieved since the major obstacle was to  motivate employees to change 
their habits. With perfect hindsight the researchers claim that understanding 
and adapting solutions by solving one problem at a time would have increased 
the benefits. Then, despite the reported results, starting the establishment of 
a simple and informal workflow aided the organization to move from the stage of 
knowing nothing to knowing at least something about their development process.

The integration of minimal project management during the development activities
can foster the control over development productivity and control the
\textit{accumulated technical debt} (CAT5), but still with particular attention
to the development workflow status within the company, designing specific and
tailored solutions.

%MUN: To which category does the following paragraph belong?  
%CG: they represent subcategories of speed-up development  
Eventually to externalize complexity of the developing product 
%MUN: solutions on what? CG: fixed 
, two other studies \cite{Wall2001, Bean2005} suggest to \textit{``reuse the  
wheel, instead of reinventing it''}, taking advantage of open source software  
whenever it is possible, building systems quickly and effectively as a result.  
Even though the principles in leading-edge development processes are the same
principles of yesteryear, the way in which the fundamentals are applied has
changed from prescriptive to agile solutions, with less concern on quality
aspects, also confirmed by Mater \cite{Mater2000}.

%------------------------------------------------------------------------------%
\subsubsection{Simple workflows} Prescribed development models, oriented towards
SPI, are basically ignored by  startups, as discussed in \cite{Zettel2001}.
%MUN: <-- rephrase, not clear to me %CG:fixed 
To overcome the need of structure and to accelerate the development process,  
Zettel developed a \textit{``LIghtweight Process for E-business software  
development''} (LIPE).
This lightweight development method integrates the XP  approach with ideas from
areas of software measurement for project control and  process improvement as a
compromise between an ad-hoc and a more rigorous  approach. Zettel argues that
LIPE provides scalability of the development  process with a certain degree of
flexibility, omitting parts of prescribed  practices. Nevertheless, we couldn't
identify any empirical evaluation of LIPE  that supports its suitability in a
startup context.

Deakins \cite{Deakins2005} designed a new spiral methodology, constructed
during a fieldwork with a developer team that had limited resources managing an
innovative product in volatile e-commerce environment. This methodology
represents a fast evolutionary approach with flexible processes.  It aims to
obtain customer's feedback as soon as possible for continuous  product
improvement.%MUN: %<-- rephrase, sentence not parsable CG:fixed 
Even in this case, the spiral methodology%MUN: which model? CG:fixed 
is compliant with the nature of a random, ad-hoc and  \textit{evolutionary  
development approach} (CAT2). Evolutionary development approaches are  
typically more suitable for web applications, as confirmed in  
\cite{Deshpande2001}. Moreover, they provide practices oriented towards rapid 
and iterative design with complete freedom to the developer team, enabled to 
operate anywhere within the organization. 
%MUN: <-- the whole paragraph is not that clear to me. Try to start  
%with the category (team is catalyst), then write about Deakins spiral model 
%CG: partially fixed, the focus here is on the evolutionary development approach

%------------------------------------------------------------------------------%
\subsubsection{Full-stack development team}

Carmel \cite{Camel1994a} reports on the need of flexible and rapid development
solutions to shorten time-to-market, but even more important, the need of  high-
skilled team developers, which has an impact on \textit{speed-up  development}
(CAT1). He suggests that entrepreneurs need to look for a well  formed, skilled
core development team and not just a set of product ideas and  features.
Moreover the team must be empowered with full-stack and  self-organization
settings, paired with the flexibility of minimal bureaucracy.  A studied company
stated: \textit{``we didn't need weeks or months of detailed  modeling and
documentation, but rather modeling the architecture a little and  then either
exploring strategies by providing users with code or simply starting  work on
the actual software itself''}.

%MUN: When you subdivide this sections, as proposed above, finish with a
%subsubsection "Summary", summarizing the supported you have found in
%literature for your theoretical categories. Nothing extensive, 3-4 sentences
%to conclude this part of the validation, similar as below but a bit more
%explanatory (what does "compliant with the behavioural model of the
%phenomenon" mean?).  CG: not fixed, when I wrote it I tried to have an overview
%of the main  contributions and make it fluently by papers description. Creating
%subsections  would mean to divide papers from different focuses. This is a
%trade-off where  we took a decision in the past. Take it in consideration, but
%we can discuss  it during the next meeting. MUN1: Why would creating
%subsubsections like 5.2.1 Severe lack of resources  (CAT7) remove the flow?
%This subdivision would give the reader clues what the  subsubsection is about,
%he can relate it to Table 2, and provides some  navigational support. Now there
%are nearly 2p of flowing text with now clue in  what order/priority the
%literature is covered. I think you don't even need to  rewrite much, just add
%the subsubsections and a proper summary. CG: partially fixed

%------------------------------------------------------------------------------%
\subsubsection{Summary} Summarizing the comparison, we have found that most of
the  contributions of the studies are compliant with the GSM model. Startups
focusing on simple workflows and full-stack engineers face the uncertainty and
immaturity of the company. Nevertheless few considerations are presented towards
the consequences of a minimal project management and the use of low engineering
activities as presented in the GSM model. However, the capability of the model
to accommodate the results identified by previous researches give us more
confidence in the generalizability of the GSM model.

%------------------------------------------------------------------------------%
\subsection{Confounding factors} \label{sect:an:comp:cat-vs-literature:conf}

The purpose of this subsection is to identify which confounding factors need to
be taken into consideration when evaluating the GSM. Confounding factors are
those variables which and have not been considered in the model and might
interfere with the theoretical framework positively or negatively
\cite{ColinRobson2009}. We report those factors that have been explicitly
identified by the SMS \cite{SMS} %MUN: do you mean the SMS? Then refer to it.
%CG: fixed  
%I remember you mentioned CF in the results (GSM), maybe also refer to the  
%respective sections CG: not parsed. 
%MUN1: me neither... forget what I wrote :-). CG: ok 
, related to: creativity and innovation, market requirements and application  
type, and developer experience.

%MUN: Again, structure this into subsubsections according to the above factors.
%I did it, but check if you agree... 
\subsubsection{Creativity and innovation}
Heitlager et al. \cite{Heitlager2007} contributed to the investigation of
creativity and innovation in startups.  
%MUN: the rest of this paragraph is too detailed. I'd remove this "The study  
%reports how development approaches oriented to the product in the early-stage  
%life-cycle, called \textit{fluid} have main impact to innovative solutions  
%initially. After a while the need of structuring and controlling the growing  
%company size (along with \textit{transition} and \textit{specific} phases)  
%arises the need of project management activities and software processes for  
%long-term scalability issues."  CG: ok. 
%MUN: Please reformulate the first sentence, I cannot parse it. CG: fixed 
The proposed model identifies the three lifecycle stages of a
startup:  \textit{fluid}, \textit{transition} and \textit{specific}. At each
stage, the  model presents different development approaches (product-oriented
and  process-oriented), according to the level of innovation achieved. The study
reports how product-oriented development, in contrast with forced and mature
process imposition, provides a degree of freedom to the development team, that
enhances the creativity of developers and augments the innovation capability of
the company in the early-stage.

In the interviews we conducted, and on which the GSM is grounded, the theme of
innovation only seldom emerged. However, if innovative and creative factors were
considered in this model, the integration of results would have been
straightforward, since it is compliant with the description of how the
\textit{accumulated technical debt} (CAT5) causes the need of establishing a
process in the long run (see how the \textit{initial growth hinders performance}
(CAT6)). %MUN: not clear to me what you mean by "evolution  
%of this model" and what "compliant with our assumption" means. Also, don't use 
%only CATx in this context, but cite the category name AND CATx. CG: fixed 
From  the \textit{fluid} phase to the \textit{specific} phase, the innovation  
confounding factor seems to behave initially the same as the \textit{speed-up  
development}, both highly focused towards a product-oriented development.  
%MUN: <-- please provide some motivation... not obvious. CG: fixed 
Nevertheless, we were unable to ground these evidences in the GT data even  
though those are factors that might be positively correlated
%MUN: "present strict correlation" %What does that mean, I don't understand.  
%CG: fixed. 
to the choice of an early product-development orientation and the need of  
low-precision engineering elements.

\subsubsection{Market requirements and application type}  
Market requirements \cite{Coleman2007} and application type \cite{Sutton2000,  
Coleman2008,Coleman2008a} are two further potential confounding factors to the  
GSM. Their main impact is related to the adoption of flexible and reactive  
solutions for the development process. In particular, they refer to the  
necessity of fulfillment of quality concerns that goes beyond scalability and UX 
defined by the GSM. Especially when requirements are rigidly imposed or the  
application domain is well-known, providing low-quality products to final users 
might determine the failure of a startup.

For example, safety-critical systems possess quality concerns that must be
considered. Even in e-commerce systems factors related to quality have  high-
impact effects on obtaining the first customer reference,  
%MUN: <-- cannot parse it. Remove the first part and write: "Even e-commerce  
%systems...". Then rephrase the second part of the sentence to make it clearer. 
%CG: fixed 
as discussed in \cite{Deakins2005,Kim2005,Silva2005}. Quality attributes
especially %MUN: what is "this"? CG: fixed 
characterize those applications, whose part of functionalities are already  
successfully implemented in other software systems or whose failure might  
inflict serious financial or social damages \cite{Bass2003}. Yet, limiting our  
context to early-stage software startups, which typically develop first  
cutting-edge products in beta versions, partially mitigates the high quality  
demands of users.

%MUN: The following 4 paragraphs explain how you use the Kano model to position
%startups w.r.t. customer expectations and quality concerns. It's ok, but it
%adds little value to the validation of GSM. I'd remove it. To understand how to
%obtain processes in accordance to a set of customer  needs, correlating the
%proper levels of customers' satisfaction, we  introduce the Kano model,
%described in \cite{Ullah2011} and extensively  studied in product development.
%As shown in the Kano model, users  preferences can be classified into four
%categories: \textit{basic},  \textit{excitement}, \textit{performance}  and
%\textit{indifference} areas.   The \textit{basic} area represents the space
%where customers are dissatisfied  if those \textit{must-be} functionalities are
%not met. This is the case of  similar functionalities already implemented in
%the current market with  well-known solutions. The \textit{performance} area
%represents the space where  customers' satisfaction is proportional to how many
%of functionalities are  implemented. In the \textit{indifference} area the
%customer does not really  care about the functionalities: the satisfaction
%remains invariant. This  represents the case where startups have not (yet)
%achieved the right  product/market fit. Finally, the \textit{excitement} area
%represents the space  where customers will be more satisfied if functionalities
%are implemented, but  not disappointed if they are not implemented. This is the
%case of extremely  innovative product development.   Well-known domains such as
%e-commerce applications or other \textit{``quality  critical products''} belong
%to the \textit{basic} area, where if a quality  concern such as security in
%bank transfer services is missing satisfaction of  customers will be profoundly
%affected. But as discussed, in early-stage  startups that typically develop
%first beta products, developers usually deal  with cutting-edge technologies.
%Those innovative solutions, if in the right  market, are likely to belong to
%the \textit{excitement} area, where  satisfaction can hardly be negatively
%affected by faulty products. In fact, in  the GSM model, minimal
%functionalities and less concern about quality aspects  have low-impact on
%customers' satisfaction, supporting startup's ability to  mainly focus on
%\textit{speed-up development}. Moreover operating in the  \textit{excitement
%area} is fully compliant with an evolutionary approach and  minimal
%functionalities since from the first product released to the market,  it is
%unlikely to lose users because   of high-quality demand.   Nevertheless, this
%aspect profoundly undermines the generalizability of the  theoretical model
%according to the theory, in particular \textit{CAT3}.  therefore, future
%studies within wider startups' context are advisable to  extrapolate insights
%from those situations where quality concerns cannot be  postponed or
%externalized. CG: ok

\subsubsection{Developer experience}  
%MUN: Ups, a confounding factor you did not mention initially. Add it...  
%CG: fixed 
Finally, also the experience of team-members is considered as an important  
confounding factor. Since only little evidence has been gathered during the  
interviews, ``lack of experience'' was not considered as a sub-category in  
\textit{severe lack of resources}.
Nevertheless, startups rely at the beginning  on clever, but inexperienced
developers \cite{Crowne2002}. Having team-members  with deep experience is a
``double-edge sword'' \cite{Yoffie1999}. On one hand,  experience might quickly
provide structure and maturity to the development  process, on the other it
might cause challenges in managing self-confident  overachievers that almost
inevitably clash. Consequently, team management  requires higher control and
coordination activities that inevitably hinder the  flexibility of the
development environment, that is essential in early-stage  startups.
Accordingly, also Coleman \cite{Coleman2007, Coleman2008a,Coleman2008} states
that the operating strategies towards  \textit{``minimum processes''} in
startups is not a matter of poor knowledge and  training, but rather it is the
necessity of operating with solutions that let  the company move faster.

%MUN: removed below paragraph since it does not add much Nevertheless, since the
%amount of the results did not allow ``the lack of  experience'' to be in the
%GSM model (in view of the fact that not enough data  were identified
%\footnote{A category is formed if the codes are representative  of more than
%the 70\ of the analyzed transcripts.}), this confounding factor  requires
%further investigation. CG: ok

%------------------------------------------------------------------------------%
\subsection{Static assessment of the GSM} \label{sect:theory:validation:stats}

%MUN: explain in the research method section (new section on how you performed
%validation), exactly how you did static validation and what statistical
%methods you used on what data, then refer to that section from here (no need
%to repeat again here). CG: fixed 
In this subsection we validate the most important causal relations presented in 
our model by attesting their correctness  with a quantification of empirical  
trends observed in the interview data.  
%MUN1: The following is part of the method and should be in 3.5: "We identify  
%trends and recurrent patterns by thoroughly analyzing and synthesizing the  
%complete set of data obtained from the case study on which the GSM is built  
%upon. We illustrate our observations aided by simple statistical methods,  
%providing researchers an additional tool, supported by further companies'  
%data, to validate our model." CG: not important, I would not add this.

We argued in section \ref{res:gsm:frmov}, that the most essential feature
startups aim to achieve is high speed. As illustrated in Figure \ref{fig:gsm},
the core category (\textit{speed-up development}) is influenced by: the degree
of quality concerns (contributes to the \textit{efficiency}); the undertaken
development approach (provides \textit{effectiveness}); and finally the
intrinsic characteristics of the team (guarantee \textit{performance}).

%MUN: for the paragraphs below, you need to explain what and why you do this.
%CG: fixed, explained in methodology  
%MUN:What do you mean by "we grounded in the interview transcripts the fact that 
%those relations were characterized in terms of performances since  
%time-to-market was the main concern." Not clear to me.  
%CG: fixed, performance for us is efficiency AND effectiveness 
Before starting the evaluation of the impact of the three categories on  
\textit{speed-up development}, we grounded in the interview transcripts the fact 
that those relations were characterized in terms of "effectiveness" and  
"efficiency" since time-to-market was the main concern (as confirmed in  
\cite{SMS}):

\begin{compactitem} \item The team enhances the overall \textit{performance} of
software  development. Recalling that startups deal with severe lack of
resources, the  team members are the company's main assets. They develop
software with no need  of formalities and structures, left to the team's own
capabilities to  efficiently develop the product. Moreover, the effectiveness is
granted by  continuous advice given by mentors and by a management style
oriented towards  self-organization, full-stack, and multi-role settings. The
importance of a  skilled team for development performance is expressed in many
studies such as  \cite{surviving-os-cockburn, Highsmith2000}, described in CAT4
and confirmed by  Sutton \cite{Sutton2000}. \item The evolutionary approach
enhances the \textit{effectiveness} in  implementing the correct
functionalities. This approach improves the company's  capabilities to adjust
the trajectory of product development (in agreement to  the concept of pivoting
discussed by Ries \cite{Ries2011}). \item Since product quality has low
priority, the development is focused only on  implementing a limited number of
suitable functionalities, enhancing its  \textit{efficiency}. An essential
prototype (MVP) does not require to comply  with heavy quality constraints,
enabling the team to quickly implement  functionalities, ready to be validated
by final users. \end{compactitem}

An optimal combination of these three categories enables startups to ship code
extremely quickly. However, the price that startups have to pay for achieving
such high speed is related to the extent of the \textit{accumulated technical
debt}, which will finally contribute to hinder performance in later stages.

Accordingly, by analyzing the data of the 13 companies, we formulated two
hypotheses which further describe the above illustrated relations: 
%MUN: You need to be more precise with terminology. Above you state that the  
%two items below are "trends which further empower relations", but then,  
%further below, you write "to assess these hypotheses". Observed trends !=  
%Hypotheses. This is also why you need to be extremely explicit and precise  
%when explaining what the a) aim, b) method, and c) outcome/result of this  
%section is. I'm aware that it is difficult to boil down what you wrote in the  
%thesis to this limited space, but I think it is doable (and relevant for  
%supporting validity of GSM).
%CG: fixed

\begin{compactenum} 
\item The more a capable \footnote{The team capability is
evaluated according  to the characteristics presented in CAT4.} team, with few
quality constraints,  uses an evolutionary approach, the faster the company
releases the product. 
\item The faster the company releases the product, the larger the degree of  
\textit{accumulated technical debt}\footnote{The debt here is considered  
regardless of context-specific debt mitigation factors and tactics.}. 
\end{compactenum}

To assess these hypotheses we defined three measures reflecting the theoretical
model:

\begin{compactitem}
\item \textit{Execution speed}: a metric that represents the development speed
of the startup during different phases of the first release, computed by means
of a weighted average speed for each phase. It was obtained by analysing
interview transcripts looking at subcategories of \textit{Speed-up development}.

\item \textit{Technical debt}: a metric that represents the extent to which
processes are controlled, structured, planned and documented by means of
engineering artifacts and practices. It has been computed by means of a weighted
average of the debt accumulated in each development phase, observing
subcategories of  \textit{accumulated technical debt} with consequences on the
startups growth.

\item \textit{Potential capability}: a metric that represents the degree to
which each company reflected the capability of reaction and flexibility to the
dynamic environment during the development process, given by the three
categories that mostly affect \textit{speed-up development}. 
\end{compactitem}

We quantified these measures for each company involved in our empirical study
by defining score metrics based on a set of rubrics and evaluating startups
accordingly, through an analysis of interview transcripts, codes and
questionnaire results. The complete statistical procedure and provided rubrics
to quantify the measures are illustrated and discussed in Appendix. 
%MUN: make sure it is clear that the PROCEDURE material is available in the  
%ONLINE appendix and refer to it. Still, you need to report the actual RESULTS  
%of you statistical analysis here...  
%CG: partially fixed. we always referred to appendix without references until  
%now.

%MUN1: I don't really understand why you add the second clause in H01  
%"AND with less quality constraints" --> just makes the hypothesis more  
%difficult to understand. CG: fixed. 
We conducted statistical tests  using the analysis of
variance to assess the  existence of relations between the dimensions. First we
defined two null  hypotheses:

\begin{itemize} 
\item[$H0_1$:] \textit{Startups do not release the product
faster  when a capable team adopts a more evolutionary approach AND with less
quality  constraints} 
\item[$H0_2$:] \textit{The execution speed does not
increase the amount  of accumulated technical debt}.
\end{itemize}

The formulation of the two hypotheses reflects the representation of the
categories related to \textit{speed-up development} presented in the GSM model
(see figure \ref{fig:gsm}). We tested $H0_1$ and $H0_2$  with an  one-tailed
test using Pearson's product moment correlation  coefficient\footnote{To perform
the Pearson's correlation we verified two  assumptions: a) data is normally
distributed; b) data is on interval scale.},  with positive association
analysis, fixing the level of confidence to 95\%. We reject both $H0_1$ and
$H0_2$, and conclude that in our sample: 
\begin{compactenum}
\item Higher values of \textit{Execution speed} are strongly associated  
(p-value: 0.002) with higher values for \textit{Technical debt}.

\item Higher values of \textit{Execution speed} are  strongly associated
(p-value: 0.005) with higher values for   \textit{Potential capability}.
\end{compactenum}

The validation has been performed on the most critical  causal relationships of
the model. According to the paradigm model those are the causal condition which
are linked to the core-category (see Figure \ref{fig:gsm}). There are four
high-level relationships  which have not been validated by quantitative
methods. However their correctness has been verified by a cross-analysis with
existing literature and by attesting the causal relations among codes and
categories, through axial and selective coding processes.

%------------------------------------------------------------------------------%
\section{Implications of the GSM} %MUN:Implications of the GSM. CG:fixed
\label{sect:theory:impl}

%MUN: There are two issues with this section: it is too long and it basically
%reiterates (in a concise manner) what you have said in the preceding 12 pages.
%The real value of this section (suggested by my proposed name above), would be
%if you can state in 0.5 (MAX 1 page), what are the implications of your
%results on THEORY (or state-of-art, research) and PRACTICE (state-of-practice,
%people who want to embark a startup). I'd propose you have a short intro here,
%then have two subsections: State-of-art, State-of-practice where you state the
%implications of GSM (or use a better subdivision if you come up with one, or
%use none if you don't exceed 0.5 pages). These implications need to logically
%follow from what you have said before (and the data you collected in the
%interviews and, when necessary, by reference to the SMS). Don't have a long
%bullet list, but structure your arguments in a logical flow. This is the
%section in which you have to show, based on your previous rigorous and well
%done analysis, the essence / gained knowledge, indicating what are the possible
%next steps for research in this area. You can be creative/visionary here as
%long as your arguments follow from the previous analysis. CG: not fixed, I have
%shortened the content of one page. But still we cannot  create the subsections
%you proposed since the implications are all derived  from the model. Then, we
%can still shorten it if you want. Let's discuss it  during the next skype call.

%MUN1: Ok, the subdivision is better than the plain bullet list and I'm fine
%with the subsection names. Still, I would rename the section: "Discussion" -->
%"Implications of the GSM". CG: fixed

In this section we present relevant implications which  emerge from the behavior
of early-stage startups, formally expressed in the GSM,  answering to our
research question (\textit{What is the current  state-of-practice related to
software development strategy in early-stage  startups?}) through its sub-
questions. Although the startups we inquired in the  case study were spread
across different nations and market sectors, we have  found several patterns
that emerged very clearly.

%------------------------------------------------------------------------------%
\subsection{Shortening time-to-market}

The most urgent priority of software development in startups is to shorten
time-to-market. When developing a new product, which has not been attempted
before, it is crucial for the company to release the product soon and observe
the users reactions. The company's chances to survive, are strictly related to
the ability to find the right product/market fit as quickly as possible.
Developing software without releasing it for a long period of time, coming up
with very complex solutions often is not a viable option. Business demands,
investors interests, market pressures and the general uncertain conditions force
startups to quickly iterate on the product by exploring and experimenting new
functionalities progressively rolled-out to early-adopters. 
%MUN1: Ok, I'm now "inventing" possible implications in the sections below...  
%this is your job and you are much more qualified to do that than me :-)

%MUN1:Do we (research) know enough about how to shorten time-to-market without
%accumulating technical debt? What are possible venues for research? CG: fixed
%see summary.

%------------------------------------------------------------------------------%
\subsection{Light-weight methodology} 
Startups do not apply any standard development methodology. When focused  on 
building the first version of the product, startups do not observe any  
specific and standard development methodologies or processes.  From an accurate 
and in-depth investigation into daily activities, it emerged that the number of 
the observed development \textit{practices} is extremely small and never  
systematic (only little \textit{pair programming} sessions and rarely  
\textit{test driven development}). Despite the wide number of publications,  
which discuss the adoption of lightweight methodologies, we have found that  
especially in this first stage not only methodologies but even individual  
practices are basically neglected.

The closest development approach undertaken by early-stage startups is  the
\textit{Lean startup methodology}.  A highly evolutionary development  approach,
centered around the quick production of a functioning prototype and  guided by
customer feedback is a central aspect of the \textit{Lean startup  methodology}
pioneered by Ries \cite{Ries2011}. However, startups do not  formally follow the
cycle \textit{build-measure-learn} as it is proposed by the  methodology, where
they should set-up an experiment to \textit{test the riskiest  element of the
business idea}, nor make explicit use of the suggested good  practices. Instead,
startups in the early stage prefer also to apply fast cycles of "build and fix"
when necessary, hacking the first functionalities to present to potential
customers and investors in order to obtain quick feedbacks. 
%MUN1: Why don't follow startups the build-measure-learn cycle proposed by  
%Ries? Well, naively I'd say that since he published in 2011, his ideas do not  
%have yet transferred to industry... Any other reason you can think of? Is it  
%still too heavyweight and in conflict with the highest priority of shorting  
%time-to-market? Is Lean still to Heavy for startups (in their early phases)? 
% CG:fixed. 
Moreover, when it comes to user feedback, startups do  not explicitly follow 
the step-by-step process of \textit{customer development} formally defined by 
Blank \cite{Blank2005}. Instead, startups absorb and  implement the high-level
principles that outlines the customer development methodology
%MUN1:<--  %which methodology? CG: fixed 
, reflected in the GSM by the theoretical category \textit{find the  
product/market fit quickly}.

%MUN1: paragraph below is candidate for removal The greatest part of engineering
%activities of startups are focused on  the implementation while only little
%attention is given to more conventional  activities (project management,
%requirement specifications, analysis,  architecture design, automatic testing,
%\ldots). Especially when bringing the  first product to the market, the
%attitude of startups towards traditional  engineering activities is strongly
%characterized by a general feeling of  unnecessary. Every document, artifact or
%process, which does not produce a  short-term tangible benefit to the software
%production is perceived as a waste  of time. CG: ok

%------------------------------------------------------------------------------%
\subsection{Focus on minimal set of functionalities} 
The first release of the product includes only a limited set of well  suitable 
functionalities focused on user experience. The beta version is usually  
focused on a minimal and essential set of functionalities providing a reasonable 
user experience (UX). The first functionalities are usually set by the 
founders and their experience in a particular business domain. As soon as 
possible they provide demos to demonstrate the idea to potential customers. 
%MUN1: This rises the question, how do startups prioritize requirements.  
%There's abundant amount of literature on how to select/prioritize reqs., but  
%do they work in startups? Are they also too heavyweight or could they be  
%adopted by startups in a modified form in order to increase the probabilities  
%of success in the market?. CG: fixed. 
As confirmed by practitioners, the user must be able to  accomplish a task on 
the product without major interruptions in the flow, which  would cause 
frustration before being able to capture his attention. To realize  UX in a 
short time frame, developers focus on a limited number of suitable  
functionalities, especially in the first version(s) of the product. Other  
product qualities are regarded as less important: for instance, obtaining a  
high-reliability is not a priority since early adopters of innovative product  
are keen to accept minor bugs in exchange of new functionalities\footnote{These 
results further highlight differences between software development in startups 
and traditional web engineering, where quality aspects are rated as very  
important, even more than time-to-market \cite{Offutt2002}.}. From a business  
perspective, one might argue that established companies might take advantage of 
the new idea and develop a better version of the product internally. Despite 
this might represent a risk, established companies usually cannot  afford to 
operate in a market of \textit{innovators} and \textit{early adopters}  group 
of users - described by the Roger's bell curve in \cite{Rogers} - because  of 
the little return of investment they would gain, as discussed by Christensen  in
\cite{Christensen2003}. Instead, established companies have more interests  in
acquiring a startup and all the knowledge it has gained during the
\textit{customer development} process while focusing effort on acquiring larger
shares of a ``mass market'' sector.  
%MUN1: candidate for shorting 
%Furthermore, thanks to cloud computing and more in general high availability 
%of third party components, aspects such as performance, security and 
%scalability can be achieved with a reasonable effort. Thus, given the 
%general lack of quality concerns, and supported by the fact that the optimal 
%strategy to improve the UX is by having the product used by real customers, 
%startups can effectively focus on achieving short time-to-market. CG: fixed

%------------------------------------------------------------------------------%
\subsection{Empowering the team members} 
The founding team is the most important determinant of speed. More than  
techniques and methodologies, the features of the team are the essential factors 
to achieve high-development speed. Discussed in the theoretical category the  
\textit{team is the catalyst of development}, this result has been confirmed in 
multiple occasions both in the interviews and in the questionnaires. What is  
truly important for the startup, among other factors, is the extremely  
high-colocation which enables continuous informal and colloquial coordination  
and discussions.  
%$MUN1: candidate for removal 
%The initial (small) development team usually spends most of the hours of the 
%week working closely to implement the product, often extending  
%the conventional office hours. CG: ok. 
The large amount of communication, spontaneously held in informal places, has  
been proven to be an extremely effective element of several successful software 
projects \cite{Highsmith2000}. Hence, the vastly \textit{informal} development 
environment of startups acts as a catalyst of speed. 
%MUN1: candidate for removal 
%Other important elements, which contribute in achieving such an initial  
%high-speed, are: minimal set of functionalities with low concerns on product  
%quality; evolutionary prototyping approach; smart use of advanced tools and  
%frameworks; low precision artifacts; and the close colocation and capability  
%of the team. CG: ok.

%------------------------------------------------------------------------------%
\subsection{Paying back the technical debt} 
The initial lack of structures and processes negatively affects the  
performance when the company starts to grow.
One of the primary objectives of  startups is to expand their business into fast
growing markets. If  successful, the company will face a growth in terms of
number of customers,  employees and product capabilities. Then, the increased
complexity arises the  necessity of controlling the initially chaotic software
development environment.  Under delicate circumstances, where an increased
market demand meets a poorly  engineered product and process, the development
team need to start returning  the accumulated technical debt. As reported by
practitioners, introducing more  structures, documentation and standard
procedures, cause an initial drop-down in  performance which severity depends on
the speed of growth, the amount of  technical debt and the capacity of the team
to absorb it.  
%MUN1: what does literature say about technical debt? Are there strategies that 
 %maybe applicable to startups and if not why? Is there more research needed?  
%Obviously, shortening time-to-market and preventing technical debt are two  
%conflicting goals, but is there a chance to find a sweet spot where  
%development speed can be balanced to achieve market success and still reducing 
%the performance bump that hampers growth? 
%CG: partially fixed below in summary.

\subsection{Summary} All of the above observations are extracted from the
analysis of the model,  which is systematically derived from empirical data of
the case study. Indeed,  with slightly different levels of adherence, the
implications we presented are  reflected in the behavior of the great majority
of companies in the sample. The  results of this analysis, which has been
focused on the software development  strategy from the idea conception to the
first release, indicate that startups  are far from adopting standard or ad-hoc
development methodologies, especially  in this phase. The typical tendency is to
focus on the team capability to  implement and quickly iterate on a prototype,
which is released very fast. Thus,  in a context where even for the most
\textit{lightweight} agile methodologies is  hard to penetrate, we believe that
it is premature to discuss issues related to  SPI. Researchers should rather
focus on the trade-off between development speed  and accumulated technical debt
\cite{Brown:2010:MTD:1882362.1882373}, which  appears to be the most important
determinant for the future of  software development in the company.

Future research can investigate the performance drop down caused by the
necessity of returning the accumulated technical debt while expanding the
company's operation and structuring mitigation strategies with four software
development objectives: integrating scalable solutions with fast iterations and
minimal set of functionalities (this allows startups to maintain effective
planning and realistic expectations); empowering team members enabling them to
operate horizontally in all the activities of the development environment
simultaneously; improve desirable workflow patterns through the initiation of a
minimal project management over time, as a natural result of emerging activities
of tracing project progress and task assignment mechanisms; and finally, only
when the chaos has been initially managed, planning long-term performance by
adoption of Agile and Lean development practices.

%------------------------------------------------------------------------------%
%\subsection{Future work} %label{futwork} 
%MUN: Don't have future work as a subsection in threats to validity. Either  
%merge it in "Implications of the GSM" OR into "Conclusions". CG: fixed

%MUN1: The paragraph below does say little about future work but rather
%presents the contributions of GSM. In "Future work" you should rather state
%what you consider needed to work on in relation with the research you have
%done. Granted, replicating your case study and validating/extending GSM is
%future work too, but it is not what *you* would do. Hence I'd suggest to
%remove this section. Check that you state the contributions of GSM in the
%conclusions.

%The GSM model enables practitioners in startup companies to speed-up
%development empowering the role of the team, focusing on minimal
%functionalities and adopting evolutionary development approaches. Researchers
%can use the GSM model to reproduce the study with other software startups, and
%focus on mitigation strategies to minimize the impact of the accumulated
%technical debt over the need of shortening time-to-market. Moreover, the
%mutual collaborative dialog between the research in the development strategies
%in software startups and technical debt can benefits in understanding how both
%influence the growth of a company after the release of the first product to the
%market. CG: ok

%------------------------------------------------------------------------------%
\section{Threats to Validity} \label{valt} 
%MUN: Also here, introduce some structure: first, explain shortly what types of 
%validity threats exist for the research methodology you applied. Then, discuss 
%each type individually. Think about which threat type is the most prominent  
%one for your research methodology and start with that one. I did not review  
%this section in detail now since it will change considerably after your  
%restructuring. %CG: fixed

%MUN1: After reading the discussion on validity threats, I think you have chosen
%the wrong framework (Wohlin) to discuss threats. His framework is  targeted at
%experiments and even though it has been applied in other contexts,  other
%taxonomies are a better fit (see Yin and also Runeson). You tried to  squeeze
%your threats and strategies into Wohlin's taxonomy which seems in some  cases
%quite artificial and forced. I think it is better if you adopt Yin's  taxonomy
%of validity threats. If you read Yin (p 33-38) you'll see that his  taxonomy is
%a much better fit and that you can easier place and motivate your  mitigation
%strategies. CG: to discuss since the ref of Yin and Runeson present very
%similar validity threats.
%------------------------------------------------------------------------------
In this section we discuss the validity of the overall methodology. 
%MUN1: threats  %are discussed w.r.t. the research methodology, not the 
%results... CG: fixed. 
We structure the discussion according to Wohlin's taxonomy \cite{Wohlin2000}.

%MUN1: Ok, what I meant was, not to explain the 4 categories in *general*, but
%to explain (shortly) their relevance w.r.t. GT. I guess my comment above was
%misleading... \begin{compactitem}

%\item External validity - which requires establishing the domain to which the
%study's findings can be generalized (partially discussed above).

%\item Internal validity - which is enhanced by establishing causal
%relationships whereby we attest that certain conditions lead to other
%conditions, as distinguished from spurious relationships\footnote{A
%\textit{spurious} relationship is a relationship in which two events have no
%direct causal connection, but rather due to either a coincidence or the
%presence of confounding factors.}. 
%\item Construct validity - which is enforced by clearly specifing operational  
%procedures, and providing an adequate description of  
%constructs\footnote{Constructs are complex ideas that we as humans form in  
%order to summarize observations about things that we cannot see directly. For  
%example a construct is \textit{``job satisfaction''}, which to be defined,  
%requires a higher level of abstraction than a concrete object(e.g. a book).}  
%that are used to generate the theory. 
%\item Conclusion validity - which represent the degree to which conclusions  
%are reasonable, according to the defined relationships in the data. It  
%involves ensuring adequate sampling procedures, appropriate analytical methods 
%and reliable measurement procedures. Reliability requires to demonstrate that  
%the operations of a study - such as data collection procedures - can be  
%repeated by other researchers, obtaining the same results. 
%\end{compactitem}

%------------------------------------------------------------------------------%
\subsection{External validity}

One threat to external validity is the selection of subjects interviewed  for
the study. This threat profoundly affects GT, since it is a qualitative
research method, which makes the use of semi-structured interviews, and centers
the research on respondent's opinions.  In order to mitigate this threat we
selected interviewees which  covered positions of CTOs' and CEOs'. Their
perspectives on what is taking  place within the organization, were the only
meaningful data taken in  consideration in the study.  In many cases there was
no supporting evidence to  verify the opinion expressed. The questionnaire
dedicated to ``data  triangulation'' was not of primary support because of the
limited time  practitioners were able to grant to it. However, researchers have
to accept the  veracity of what respondents reported during the interviews
\cite{Coleman2008a}.  %MUN1: candidate for removal 
%Notwithstanding, the issues surrounding semi-structured interviews are vital.  
%Within our study, although the reality might be different from what has been  
%described, the reported data are CTOs' and CEOs' perception, which is used to  
%base their decisions. CG: ok

The comparison of the GSM with similar frameworks helped in establishing the
domain to which the study's findings can be generalized. In particular the
previous framework developed by Coleman \cite{Coleman2008} has allowed  us to
find similarities, differences and broader reasoning related to  factors that
hinder maturing processes in startups. Moreover, we analyzed  Brook's framework
describing the \textit{basic attacks} to the challenges of  developing software
systems and other literature covered by the SMS on startups.
\cite{SMS}\footnote{Note that the generalizability of results is applied only to
those startups that operate in software-intensive activities.}.  
%MUN1: candidate for removal: detail on RM / seems out of place 
%In conclusion, regarding the validation of  
%the subcategories, a continuous re-evaluation of the interview transcripts has 
%been performed. CG: ok

%------------------------------------------------------------------------------%
\subsection{Internal validity} %Internal 
To enhance the internal validity, we created a three-dimensional research  
framework. By means of a grounded theory approach, formed by a  
\textit{systematic mapping study}, \textit{interviews}
and \textit{follow-up  questionnaires}, we searched for convergence among
different sources of  information to confirm the generated theory.

To validate the GSM we conducted a comparison of the emergent theory with
existing literature and previously developed models. Furthermore we
quantitatively evaluated three important relationships in the model. With the
theory validation we highlighted and examined similarities, contrasts and
explanations as discussed in \cite{Eisenhardt2007}. In this regard Eisenhardt
stated: ``Tying the emergent theory to existing literature enhances the internal
validity, generalizability, and theoretical level of the theory building from  a
case study research [\ldots] because the findings often rest on a very limited
number of cases.''

%MUN1: candidate for shortening 
%Little ambiguities about direction of causal influences have been reduced by  
%introducing arrows to indicate the direction of the causal relation in the GSM 
%model. Furthermore, by means of comparison between the GSM model and extant  
%studies in the broader fields, we revealed that the theory resembles in wider  
%startup development context. CG: ok
Nevertheless, we identified important confounding factors, related to
innovation, market requirements and developer experience, which might impact
the internal validity of the GSM.

%------------------------------------------------------------------------------%
\subsection{Construct validity} %Construct 
One threat to this study is a possible inadequate description of constructs. In 
order to diminish this risk, the entire study constructs have been adapted to  
the terminology utilized by practitioners and defined at an adequate level for  
each theoretical conceptualization. For instance, we defined \textit{Time  
shortage} in terms of \textit{Investor pressure, CEO/business  pressure, Demo 
presentations at events and internal final deadline} as used by  most of the 
interviewees in the study.
Moreover, during the coding of interview transcripts, we adopted explanatory
descriptive labels for theoretical categories, to capture the underlying
phenomenon without losing relevant details.

%MUN1: I get the threat, but I don't understand its mitigation: let the
%interviewees drive the study?! Try to reformulate... CG: fixed. 
The second important threat is caused by the fact that interviewees might  
already be aware of the possible emergent theories analyzed by researchers. To  
reduce the risk that some hypothesis could have been involved in the design of  
the research, we didn't disclose any goal and emergent results of our research 
to the interviewees.

Another threat associated with construct validity is evaluation  apprehension,
which describes people's rejection attitude towards the  assessment of their
behavior.  To address this risk we didn't attained to first answers from
practitioners, but  rather we asked for details of artifacts and engineering
strategies. %MUN1: <--  %I don't understand this... CG: fixed 
Moreover, we developed a rigorous data collection protocol to create case study 
databases and employ integration of multiple data collection methods.%MUN1: <-- 
 %Since you write "Moreover, ..." the reader gets the impression that this is a
%further mitigation strategy to evaluation apprehension. I don't see however
%how this mitigates this risk. Fix... CG: right, removed because misleading.
Reporting bias was mitigated by packaging all the needed  material for
conducting new researches in other contexts in order to criticize  our findings
(interview package with instructions has been made available
online\footnote{Available at  \url{https://github.com/adv0r/BTH-Interview-
Package}.}. %MUN1: --> What exactly  
%is reporting bias? Let's assume that it means that you only report results,  
%i.e. interviewee statements, that are favorable to your theory. How does  
%publishing the auxiliary study material help to mitigate this risk? Publishing 
%the raw data (what you not do) would only allow researchers to check the  
%conclusions you made. The published auxiliary material helps to replicate the  
%study, i.e. makes is possible to test your theory in a different context. 
%CG: by this, we meant providing all the material to replicate the same study 
%with other startups and compare the results. 
Moreover, two researchers not involved in the execution of the study
conducted  a peer-review analysis of the model's constructs.

To control distortion during analysis we made extensive use of memos. To
demonstrate that we were aware of personal biases, memos provided us another
important function in controlling the quality of data analysis. Through the use
of memos and comparative analysis we were able to check if data fitted into the
emerging theory and also countering subjectivity that ultimately enhanced the
likelihood of producing accurate research findings.

%------------------------------------------------------------------------------%
\subsection{Conclusion validity}

Grounded theory has been already applied by other researches in similar
contexts to attest relationships among conceptualizations of an examined
phenomenon (see \cite{Coleman2007, Basri, Coleman2008a}). Those relationships
between sampling and analysis should be verified in such a way that emerging
findings remain consistent as further data are collected. To attest this, we
explored and tested:

\begin{compactitem} 
  \item Each category and the strength of relations between them. 
  \item Hypotheses, derived from and related to the emergent theory. 
  \item Deviant cases to ensure robustness and general applicability. 
\end{compactitem}

In particular we were prepared to modify generated categories so that the new
data could be adapted into the emerging theory according to the concepts of
theoretical sampling and theoretical saturation.

According to the theoretical sampling concept, we adjusted our  research design
and the emergent theory until only marginal results were  generated. Moreover to
enhance reliability of the outcome conceptualizations and  relations, the
process of coding interviews has been systematically conducted  following a
detailed process. Then, we asked to other practitioners a  confirmation to
further validate the generated framework and enhance the  reliability of the
relations among the identified categories.

An important issue is related to the fact that the limited number of interviews
might not represent the complete scenarios in our context of study.
Nevertheless, this issue has been partially mitigated as result of the
theoretical saturation concept. This experience is attested by Martin's
statement: \textit{``By the time three or four sets of data have been analysed,
the majority of useful concepts will have been discovered.''}
\cite{giardinoEtAl}. To obtain major certainty of the obtained results, what we
applied was the final validation.

Despite we couldn't apply any power analysis procedures, in grounded theory the
sample size depended on when data saturation occurred. Ramer in \cite{Ramer},
comparing quantitative to qualitative researches, states: \textit{``reaching
data saturation, which involves obtaining data until no new information emerges,
is critical for obtaining applicability in qualitative research''}. To avoid
interrupting the research prematurely, after attesting that no more relevant
information could be gathered from executing additional interviews, we iterated
grounded theory study one more time, verifying that the explanatory power of the
core category was fulfilled.

The  approach of generating theory allowed us to check emerging categories  and
their properties by gathering new evidence within an iterative and  evolutionary
approach, in view of the wide context of the research area that was  under
investigation.

%------------------------------------------------------------------------------%

\section{Conclusion} \label{conc} 
Startups are able to produce cutting-edge software products with a wide impact  
on the market, significantly contributing to the global economy. Software  
development, especially in the early-stage, is at the core of the company's  
daily activities. Despite their severely high failure-rate, the SMS \cite{SMS}  
found that the quick proliferation of startups is not supported by an adequate  
and scientific body of knowledge. %MUN: <-- last sentence: isn't that a  
%conclusion from the SMS? CG: fixed 
To be able to intervene on the software development strategy with scientific and 
engineering approaches, the first step is to understand startups' behavior.  
Hence, in this research work, we provided an initial explanation of the  
underlying phenomenon by means of a grounded theory case study, focusing on  
early engineering activities, from the idea conception to the first open beta  
release of the product.

Through an exploratory research, conducted by a systematic approach, we
produced a theoretical model grounded into the hindsight knowledge collected
among practitioners with the aim of explaining how development strategies are
engineered startups. The explanatory capability and correctness of the model
have been validated by means of systematic comparisons with the state-of-the-art
results and empirical data. The systematic mapping study \cite{SMS},  with 37
studies extracted, revealed a  multi-faceted state-of-the-art inadequate to
support software development  activities in startup companies. 
%MUN: <-- you mix again SMS and GT here... fix it CG: not fixed, I would keep 
%it since important part of the initial GT methodology 
%MUN1: Ok, I understand that. But this paper is about the GT, ONLY the GT. The  
%SMS is published elsewhere. Maybe it is ok how you use the SMS now... let's  
%see what Tony says... CG: ok 
On the other hand, the case study conducted in 13 early-stage startups,  
provided a broad set of empirical evidences obtained by combining different  
research methodologies in a grounded theory approach.

The overall results of our research confirm that startups possess 
%MUN:  %why unique? CG: fixed. 
characteristics of uncertainty, lack of resources and time-pressure. These  
factors influence the software development to an extent that
transforms every  decision related to the development strategies into a
difficult trade-off for  the company. Moreover, although startups share
different characteristics with  other similar SE domains (e.g. market-driven
development, small companies and  web engineering), the unique combination of
coexisting factors poses a whole new  set of challenges which need to be
addressed by primary studies. Especially when  bringing the first product to
market, startups' most urgent priority is  releasing the product as quickly as
possible to verify the product/market fit  and to adjust the business and
product trajectory according to early feedback  and collected metrics. In this
stage, startups often discard any formal project  management, documentation,
analysis, planning, testing and other traditional  process activities.
Practitioners take advantage of an evolutionary prototyping  approach, using
well-integrated tools and externalizing complexity to third  party solutions.

However, the initial gain obtained in terms of flexibility and speed is
counterbalanced by the need of restructuring the product and controlling the
engineering activities when the company starts to grow. If successful, the
startup will face a growth in terms of number of customers, employees and
product functionalities that leads to the necessity to control the initially
chaotic software development environment. In this context the most significant
challenge for early-stage startups is finding a sweet spot between being fast
enough to enter the market early while controlling the amount of accumulated
technical debt.  %MUN: Not sure whether you have discussed/mentioned the meaning
%of chaotic/complex domain in the paper, I think not. Hence you need to
%rephrase that (but you can not introduce new concepts here in the conclusion).
%CG: fixed.  
Therefore, from a software development perspective, working with low-precision  
engineering activities represents the only viable strategy that startups can  
follow in the early stage of their lifecycle.

In this research we discussed a number of novel challenges for both
practitioners and researchers, while presenting a first set of concepts, terms
and activities which set the software engineering scene for the rapidly
increasing startup phenomenon. By making a comparison with Berry's definition of
SE \cite{Berry1992}, we would like to see the rise of a new discipline -
\textit{startup engineering} - which can be defined as  \textit{the use of
scientific, engineering, managerial and systematic approaches with the aim of
successfully developing software systems in startup companies}. %MUN: Nice ;-)

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

% use section* for acknowledgement
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi

The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{tse}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{John Doe}
Biography text here.
\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here.
\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}